### 2016-04-21 dissertation-export.tex
Start carving out unused dissertation material

\subsection{Introduction}\label{introduction-2}

Although imaginative and carefully designed, the Harvard course taught
that there is only one right way to approach the computer, a way that
emphasizes control through structure and planning. There are many
virtues to this computational approach (it certainly makes sense when
dividing the labor on a large programming project), but Lisa and Robin
have intellectual styles at war with it. Lisa says she has ``turned
herself into a different kind of person'' in order to perform, and Robin
says she has learned to ``fake it.'' Although both women are able to get
good grades in their programming course, they represent casualties of
this war. Both deny who they are in order to succeed. (Turkle \& Papert,
1991, p. 165)

Over 20 years ago Sherry Turkle and Seymour Papert called for
epistemological pluralism in computing education. Their argument --- in
my view ---had multiple parts:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Programming a computer involved a shift in how one thought about the
  nature knowledge from a propositional \emph{what is} to an imperative
  \emph{how to} (Abelson \& Sussman, 1996; Papert, 1980).
\item
  Historically, there was a tendency to treat computers and the act of
  programming them as an extension of the same Western logico-deductive
  knowledge tradition that fueled science.
\item
  Recent scholarship, often at the intersection of post-structuralism
  and social studies of science, had been troubling the idea that
  knowledge-work in science proceeded solely by way of logical
  deductions (Keller, 1983; Latour, 1987; Traweek, 1988)
\item
  Data from professionals suggested programming, too, didn't proceed
  solely by way of the structured-planning approach emergent from
  Western logico-deductive knowledge traditions. There were many ways of
  knowing and constructing knowledge in programming.
\item
  A viewpoint of \emph{epistemological pluralism} --- embracing multiple
  kinds and ways of knowing --- was informing the field of science
  studies but not, to a large degree, how education researchers thought
  about or taught computing.
\item
  The lack of support for diverse ways of knowing in computing
  classrooms demonstrably hurt students.
\end{enumerate}

To Turkle and Papert (1991), the result of this culturally-rooted
epistemological tension was a war. More specifically, it was a war where
students of computing were the casualties, and the aggressor was a
distributed failure to recognize, embrace, and support diverse ways of
knowing. While I think their strong-form characterization of a ``war''
no longer applies, their consideration of epistemological issues is
still relevant two decades later. An understanding that computing does
in fact involve diverse ways of knowing should continue to inform
research and discussions on learning

This paper tries to model what directs and sustains students'
in-the-moment activity when they program. Its focus is on early-stage
program design; analyzing how what students say, do, write, and gesture
\emph{even before they type code} can help us improve theories of
cognition and activity in learning programming. Using a ``revelatory
case study'' (Yin, 2009) of two students in an introductory programming
course, I argue the following:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Students' early-stage design activity reveals patterns outside the
  explanatory scope of misconception-based accounts of cognition.
\item
  For a subset of phenomena, we can recast students' productive
  capacities and their difficulties in terms of epistemological stances.
\item
  Such recastings are analytically powerful. Beyond Rebecca and Lionel,
  they could potentially explain the diversity of practices I saw other
  students take up.
\item
  As evidenced by work in science and math education, dynamic
  epistemological models can offer a lens for reforming assessment and
  instruction.
\end{enumerate}

\subsection{Literature Review}\label{literature-review}

I begin by exploring the historical legacy of Turkle and Papert's (1991)
remarks on epistemological pluralism in computing. I'll then describe
what I see as a well-intentioned obstacle: the research-based
preoccupation with students' misconceptions in computing education.
These two stra.

\subsubsection{Computing culture should support a diversity of ways of
knowing}\label{computing-culture-should-support-a-diversity-of-ways-of-knowing}

Turkle \& Papert (1991) describe a tension between individuals and a
kind of cultural collective. In this tension, individual students and
their personally-identified ways of knowing were in conflict with the
larger culture of computing and its manifested ways of knowing. Lisa,
for instance, was a Harvard student who identified as a poet:

Lisa experiences language as transparent, she knows where all the
elements are at every point in the development of her ideas. She wants
her relationship to computer language to be similarly transparent. When
she builds large programs she prefers to write her own smaller
``building block'' procedures even though she could use prepackaged ones
from a program library; she resents the latter's opacity. (Turkle \&
Papert, 1991, p. 164)

Turkle and Papert connect Lisa's programming practices to Lisa's
personal view of knowledge. She writes building block procedures because
of a continuity between herself as knower-of-her-own-poetry and that of
knower-of-her-own-programs. Across both contexts, Lisa \emph{resents
opacity} because it is an obstacle to her knowing. But, that resentment
slows her progress on projects in the class.

In part because of her commitments to transparency and the difficulty
they created, Lisa changed her approach to programming. She ultimately
had to, in her words, ``be a different kind of person with the machine''
(Turkle \& Papert, 1991, p. 164). Turkle and Papert explain:

She had been told that the ``right way'' to do things was to control a
program through planning and black­boxing, the technique that lets you
exploit opacity to plan something large without knowing in advance how
the details will be managed. Lisa recognized the value of these
techniques --- for someone else. She struggled against using them as the
starting points for her learning. Lisa ended up abandoning the fight,
doing things ``their way,'' and accepting the inevitable alienation from
her work. (Turkle \& Papert, 1991, p. 164)

The phrase ``inevitable alienation'' is particularly damning. One might
accuse Turkle and Papert (1991) of hyperbole, but research from other
disciplines suggests they're not overstating the case. Identity
alienation for epistemological reasons --- or some variation of them ---
is playing out in other subjects as well.

Connections and tensions between identities, learning, and personal ways
of knowing are not unique to computing education. Wortham (2006), for
example, detailed instances of social identification and academic
identity as \emph{jointly emerging} in a combined high school English
and History class. Nasir and colleagues explored how different learning
environments afford access to identity, domain practices, and
self-expression (Nasir \& Cooks, 2009; Nasir \& Hand, 2008). To single
out one study, Nasir \& Hand (2008) followed players ``from the
{[}basketball{]} court to the classroom.'' They found the court afforded
players a sense of role on the team, chances to be expressive through
play, and ``access to the domain {[}of basketball{]} as a whole'' (Nasir
\& Hand, 2006, p. 147). Comparatively, the mathematics classroom offered
little (if any) sense that students were on a team, a much narrower role
for them as having either right or wrong answers, and no strong sense of
self-expression or creativity through activity. These markedly different
contexts led not only to differences in kinds of activity but, Nasir and
Hand (2006) argue, divergent \emph{practice-linked identities} for the
players. In a sense, players had a different identity on the court than
they did in the mathematics classroom because who they were was strongly
coupled to a sense of their relationship with a practice.

Boaler's work on learning and identity in mathematics classrooms (Boaler
\& Greeno, 2000; Boaler, 1998, 2000, 2002) offers resonant findings. In
particular, Boaler (2000) strongly echoes the student voices in Turkle
and Papert (1991). Where Turkle and Papert's aggressor is the dominant
computing culture, Boaler describes a beast with similar effects in the
form of ``school mathematics'' --- the \emph{monotonous},
\emph{meaningless} (Boaler, 2002, pp. 383--384) imposition of routinized
mathematical procedures in ways that seem far removed from the real
world. Of school mathematics, Boaler writes:

School mathematics, for many of them, was \emph{of another world} and to
fully engage in that world, students needed to suspend their knowledge
of the real world, suppress their desire to interact with others, and
strive to reproduce standard procedures that held little meaning for
them. (Boaler, 2002, p. 392 emphasis in original)

Schoenfeld (1988, 1991) had already argued school mathematics could and
did depart substantially from the core of mathematical thinking. Boaler
established that such school mathematics could alienate learners
(Boaler, 1998, 2000) and influence their feelings about pursuing
mathematics after secondary school (Boaler \& Greeno, 2000).

Most recently, Danielak, Gupta, and Elby (in press) extended those
findings in engineering education. In their 3 ½ year case study of an
electrical engineering undergraduate named ``Michael,'' those authors
argued Michael's practice of sense-making in engineering coursework was
strongly coupled to his identity. Michael's passion for deeply
understanding concepts, much like Lisa's (Turkle \& Papert, 1991) desire
to understand every element of her program, did not mesh neatly with the
larger culture of his courses. But, because Michael's practice of
sense-making entwined with his identity, forces that pushed against
sense-making alienated him. That tension was particularly pointed when
Michael described his father's resistance to Michael's way of thinking:

{[}My dad said{]} ``you're just an undergraduate. Nobody expects
undergraduates to understand how anything works. That's why you go to
graduate school.'' I was like ``look, you know. I'm gonna be an unhappy
person if I have to....I have life goals other than to just get a good
grade on the exam. Other things are important to me.'' (Danielak et al.,
in press)

And, much like Turkle and Papert's (1991) Lisa, Michael learned to
curtail parts of his identity to get by in class:

I think the reason {[}pursuing deep learning{]} actually hasn't affected
my GPA is because I view learning as a hobby. So, as with any hobby, you
shouldn't let it interfere your GPA. But it \emph{is} one of my hobbies,
and I \emph{do} enjoy learning, I just---up to the point where I get my
grades done \{raises eyebrow\}. (Danielak et al., in press)

Ultimately, research suggests the identity-epistemology tensions Turkle
and Papert (1991) described in computing resonates with those found in
related disciplines. And, what such accounts --- particularly those of
Boaler and others (Boaler \& Greeno, 2000; Boaler, 1998, 2000, 2002;
Danielak et al., in press) --- have in common is a culturally-sanctioned
``right way'' (Turkle \& Papert, 1991, p. 164) of thinking or knowing
that alienates some students. Avowedly such ways of knowing have use and
value. Turkle \& Papert themselves acknowledge ``there are many
virtues'' to black-boxing and top-down design in programming (Turkle \&
Papert, 1991, p. 165). But,

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  When a preponderance of results from science studies suggests a
  plurality of ways of knowing among practicing scientists and
  programmers, and
\item
  When research from education shows that the inflexible imposition of
  one way of knowing above all else is alienating students, then
\item
  It's worth reconsidering whether a narrow sense of what counts as
  knowing might still be hindering improvements in computing education.
\end{enumerate}

In the next section, I turn to misconceptions research in computing
education. In keeping with observations 1--3, I argue research
approaches that inflexibly privilege canonical knowledge do so at the
expense of other productive knowledge and ways of knowing that students
have.

\subsubsection{Misconceptions research in computing education tends to
ignore students' productive
knowledge}\label{misconceptions-research-in-computing-education-tends-to-ignore-students-productive-knowledge}

In the past three decades, educational research has had a marked focus
on students' misconceptions in programming. It's a focus with a sensible
origin. Students get things wrong in programming --- often
systematically so --- and in ways that seem resistant to instruction.
The cause of those errors is theorized to be something cognitive,
whether it's a ``bug'' (Pea et al., 1987; Pea, 1986; VanLehn, 1990) a
``misconception'' (Bayman \& Mayer, 1983; Bonar \& Soloway, 1985;
Clancy, 2004; Gal-Ezer \& Zur, 2004; Herman et al., 2008; Kaczmarczyk et
al., 2010), a ``belief'' (Fleury, 1993) or a ``student-constructed
rule'' (Fleury, 1991, 2000). Instruction should try to identify,
address, and correct these misconceptions (Clancy, 2004) because they
can represent barriers to learning.

How that line of thinking and research becomes problematic is two-fold.
First, when taken in total the alleged brokenness of student knowledge
begins eclipsing all else in studying the cognition of learning to
program. In other words, most cognitively-focused educational research
in computer science treats students as having varied degrees of
deficiency with respect to canonical knowledge. Below is an unordered,
partial sampling of topics about which researchers have documented
students' misconceptions. Note across the list the variation in both the
grain sizes of students' misconceptions and the programming languages in
which they manifest:

\begin{itemize}
\item
  Objects in object-oriented programming (Holland, Griffiths, \&
  Woodman, 1997)
\item
  Algorithms and data structures (Danielsiek, Paul, \& Vahrenhold, 2012;
  Paul \& Vahrenhold, 2013)
\item
  Programming statements in BASIC (Bayman \& Mayer, 1983)
\item
  Programming in Java (Fleury, 2000)
\item
  Programming in Pascal (Fleury, 1993)
\item
  Parameter-passing (Fleury, 1991)
\item
  Arrays in Java (Kaczmarczyk et al., 2010)
\item
  Objects in Java (Kaczmarczyk et al., 2010)
\item
  Algorithms and computational complexity (Trakhtenbrot, 2013)
\item
  Boolean logic (Herman et al., 2008)
\item
  The efficiency of algorithms (Gal-Ezer \& Zur, 2004)
\item
  The Build-Heap algorithm (Seppälä, Malmi, \& Korhonen, 2006)
\item
  Hashtables (Patitsas, Craig, \& Easterbrook, 2013)
\item
  The correctness of programs (Kolikant \& Mussai, 2008)
\end{itemize}

Clancy (2004) provides a comprehensive overview of this line of
research, though in the past decade it has only grown. Indeed, roughly
half the articles above were published in the ten years since Clancy's
overview.

Identifying and removing barriers to student learning seems like a good
thing. So, it should follow that cataloging student misconceptions and
developing remedies for them should also be a good thing. But, the
logical implication isn't that clean. In some cases students display
productive, useful knowledge that's either ignored or outright
criticized by researchers. Aligning students toward canonical knowledge
makes sense, but doing so at the expense of---or in direct contradiction
to---useful ways of knowing seems undesirable at best. Next, I expand on
two examples from misconceptions research in programming. Specifically,
I show how and why I think a misconceptions focus in programming casts
aside students' useful intuitions and understandings.

The first example comes from Kaczmarczyk et al. (2010). Part of that
study involved giving students snippets of Java code and asking students
to diagram (or pseudo-code) how the information would be stored in
memory. Below I have reproduced the code for Problem 2 (Kaczmarczyk et
al., 2010, p. 110):

Cheese{[}{]} cheeses = new Cheese{[}4{]};

Meat{[}{]} meats = new Meat{[}2{]};

Turkey turkey;

Ham ham;

RoastBeef roastBeef;

boolean lettuce = true;

boolean tomato = true;

SauceType sauceType = new SauceType();

int numMeat;

int numCheese;

In diagramming this information, a student in the study makes a mistake:

Student3 makes incorrect assumptions about connections between variables
to the extent that the student makes a mistake concerning the types of
the variables. As a result, the student places Objects of different
types in an array whose type matches none of them: ``And so because
there's two arrays, cheese and meats, uh, all those turkey and ham and
roast beef are gonna be sorted into the meats array.'' (Kaczmarczyk et
al., 2010, p. 110)

The researchers are correct in the sense that \textbf{turkey} and
\textbf{ham} and \textbf{roastBeef} will \emph{not} be sorted into the
\textbf{meats} array. First, there is no code here that places
\textbf{turkey} and \textbf{ham} in the array; there is only code that
declares them as variables. Moreover, as written, an attempt to place
\textbf{turkey} and \textbf{ham} and \textbf{roastBeef} into the array
would fail. Because of type restrictions in Java, only objects of class
Meat (or objects that inherit from class Meat) can go in the array.
\textbf{turkey} and \textbf{ham} and \textbf{roastBeef} are, perhaps
confusingly, references to object instances of classes \textbf{Turkey}
and \textbf{Ham} and \textbf{RoastBeef}, so in the current snippet they
cannot enter an array of type \textbf{Meat} because (1) they don't yet
exist as objects and (2) even if they did exist, their types don't match
the array's type. The authors call this misconception \emph{semantics to
semantics}, which occurs ``when the student inappropriately
assume{[}s{]} details about the relationship and operation of code
samples, although such information was neither given nor implied''
(Kaczmarczyk et al., 2010, p. 110).

Again, the researchers are right that the student is failing to describe
the code in a way consistent with canon.\footnote{In this case, canon is
  the specifications and operations of the Java language and its
  compilers.} But, in their non-canonical thinking Student3 evidences
potentially productive insights about design. Precisely \emph{because}
there is no code stating that \textbf{turkey} and \textbf{roastBeef} and
ham are sorted into the array, the student is \emph{inferring} that to
be true. And, while that behavior is not what's happening, it
\emph{would} be sensible to design a program where specific instances of
classes \textbf{Turkey} and \textbf{Ham} and \textbf{RoastBeef} could go
into an array of type \textbf{Meat}. To do so, a designer could define
\textbf{Turkey} and \textbf{Ham} and \textbf{RoastBeef} as inheriting
from \textbf{Meat}.

Student3 has an idea about a relationship between entities where that
relationship is not specified in the code. Kaczmarczyk et al.
(Kaczmarczyk et al., 2010) focus only on the downside of the idea: the
student fails to display a proper understanding of how arrays work in
Java. Moreover, the student might be prone to similar mistakes of
inferring information that does not actually exist in code. But, there
is also an upside of this idea. Because Student3 is thinking about
real-world propositions like turkey and ham being kinds of meat, they
might be prepared to appreciate and discuss an object-oriented way to
put turkey in a \textbf{Meat} array. But, that possibility is
speculative conjecture. We can't know for certain whether Student3 could
be tipped into a productive object-oriented design activity around the
meats example because that question was not a focus of the research.

My second example of research that criticizes students' non-canonical
understandings comes from Bonar and Soloway's (1983) study of Pascal
programmers, part of which is discussed in Pea (1986). A student in
Bonar \& Soloway's study was asked to ``Write a program which reads in
ten integers and prints the average of those integers'' (Bonar \&
Soloway, 1983, p. 12). In pseudo-code, she wrote:

Repeat

(1) Read a number (Num)

(1a) Count := Count + 1

(2) Add the number to Sum

(2a) Sum := Sum + Num

(3) until Count :=10

(4) Average := Sum div Num

(5) writeln (`average = `,Average)

The interviewer then asked whether (1a) and (2a) were ``the same kinds
of statements.'' That interchange is reproduced here:

Interviewer: Steps 1a and 2a: are those the same kinds of statements?

Subject: How's that, are they the same \emph{kind}. Ahhh, ummm, not
exactly, because with this {[}1a{]} you are adding - you initialize it
at zero and you're adding one to it {[}points to the right side of 1a{]}
which is just a constant kind of thing.

Interviewer: Yes

Subject: {[}points to 2a{]} Sum, initialized, to, uhh Sum to Sum plus
Num, ahh - thats {[}points to left side of 2a{]} storing two values in
one, two variables {[}points to Sum and Num on the right side of 2a{]}.
That's {[}now points to 1a{]} a counter, that's what keeps the whole
loop under control. Whereas, this thing {[}points to 2a{]} was probably
the most interesting thing\ldots{}about Pascal when I hit it. That you
could have the same, you sorta have the same thing here {[}points to
1a{]}, it was interesting that you cold have, you could save space by
having the Sum re-storing information on the left with two different
things there {[}points to right side of 2a{]}, so I didn't need to have
two. No, they're different to me.

Interviewer: So - in summary, how do you think of 1a?

Subject: I think of this {[}point to 1a{]} as just a constant, something
that keeps the loop under control. And this {[}points to 2a{]} has
something to do with something that you are gonna, that stores more
kinds of information that you are going to take out of the loop with
you. (Bonar \& Soloway, 1983, p. 12)

Pea's (1986) interpretation? ``Here, again, we see the student believing
that the programming language knows more about her intentions than it
possibly can'' (p. 32).

As before, this student has an idea about relationships in code. Pea
(1986) see the downside of her idea: believing PASCAL can understand
shades of programmer intent when, in fact, it cannot. And again, that
downside is real. It could cause trouble for this programmer later on if
she expects PASCAL to interpret her intent and it cannot.

In defense of the student, the question --- as asked --- is vague. Are
those statements the same \emph{to whom and in what way?} Pea (1986)
treats the data as though she meant ``the same to PASCAL.'' Indeed,
maybe she did, in which case his interpretation has traction. But,
another interpretation is that she meant to herself, or to someone else
reading the code. Those statements might not be the same \emph{to her}
because she treats (1a) as having a function of controlling iteration
while (2a)'s job is to combine two numbers into a new sum.

These two purposes, which for the sake of description I'll call
\emph{keeping control} and \emph{totaling up} are, in a sense,
different. The PASCAL compiler (and runtime) does not differentiate
them, but humans can. And, humans may well \emph{want} to differentiate
them. diSessa (1986) describes exactly this kind of differentiation as a
consequence of separating the structural understanding of a programming
language from a functional understanding of a language. As an example,
he discusses the structure/function difference with respect to
variables:

The structural aspects of a variable in a computer language are given
primarily by the rules for setting their values and for getting access
to their values. These rules apply in all contexts. In contrast, a
variable's functions might vary. Sometimes they might be described as
``a flag'' or more generally, as ``a communications device.'' At other
times a variable might function as ``a counter,'' ``data,'' or
``input.'' (diSessa, 1986, p. 202)

The student in Bonar \& Soloway's (1983) study did not show evidence of
understanding the structural similarities between (1a) and (2a) in her
pseudo-code. And, those authors as well as Pea (1986) justly insist that
similarity is important for students to understand. From a conceptual
standpoint, seeing the structural similarity constitutes a part of
``knowing'' PASCAL. But, even if knowing PASCAL were not the goal,
seeing the similarity helps one to take the perspective of a computing
agent that has no means for discerning programmer intent. Such
perspective-taking may help students avoid mistakes that arise from
over-assuming what a computer ``understands.''

The student did show evidence of understanding a functional difference
between (1a) and (2a), but Pea (1986) does not remark on that kind of
understanding at all.\footnote{Also glossed over is, to me, another
  important difference: a programmer might not know in advance which
  numbers are being passed in to the sum statement. So, in advance the
  programmer can say nothing about how the value of \textbf{Sum} will
  change as the loop iterates. In contrast, the programmer knows exactly
  how the value of Count will change with each loop iteration.} Again, I
claim this oversight is part of a subtle but observable trend in
programming misconceptions literature. While, or perhaps because
research has been so preoccupied warring with students' problematic
knowledge, it has sometimes failed to recover the productive knowledge
(or resources for building it) students have. In this example, the
student already has a grasp that syntactically similar statements could
serve different conceptual purposes. Hypothetically, the student might
use that information in design by calling up the ``◻=︎ + 1'' syntactical
template when the situation seems to demand \emph{keeping control},
while calling up ``◻= ︎ + number'' when \emph{totaling up} is the goal.
And, the idea that structurally identical symbol templates can serve
different functional and conceptual purposes fits precisely in line with
Sherin's (2001) theory of symbolic forms. For example, the symbolic
forms \emph{parts-of-a-whole} and \emph{base+change} have different
conceptual schemata. Parts-of-a-whole refers to the contributions of
component entities while base±change describes a kind of accumulation.
Specifically, the terms in base±change ``play different roles'' (Sherin,
2001, p. 534) But, the two distinct conceptual schemata share what I
would argue is the same concrete symbol template of how to write an
expression: ◻=︎ + ◻.

The problem, for learning to program, comes in needing to fluidly
interpret and write code in languages that may demand incommensurable,
or at least distinct, conceptual schemata. As I show later in Table 4,
three current programming languages make remarkably different use of the
plus sign (+) as an operator. Crucially, some of the entailing ways to
make sense of how + works in those languages don't exist in Sherin's
(2001) catalog of conceptual schemata. In other words, I would argue
there are conceptual ways a programmer may need to think about
interpreting or writing a ◻=︎ +◻ symbol template that Sherin doesn't
enumerate.\footnote{One of the most obvious is the movement from seeing
  ``◻=︎ +◻'' as a statement of equality to seeing it as the storage of a
  sum to a variable.} To follow that implication, the canonical body of
knowledge about which programmers must reason is itself fractured,
because different languages design their operations around different
symbolic and conceptual metaphors.

To return to misconceptions, what drives research on students'
misconceptions is largely a need to get students to program computers
and reason about computation in ways that are canonically correct. And,
those are assuredly worthwhile goals. But, as I've argued, we can
already identify cases where a narrow misconceptions focus is silent
about or dismissive of students' useful intuitions. We can also identify
the further problems whereby unilateral emphasis on one language's canon
opposes the vocabulary of symbolic forms (and diverse conceptual
schemata) expert programmers ultimately need (Sherin, 2001) Taken in
total, that silence, dismissiveness, and narrow view of refining
knowledge perpetuates a deficit-focused discourse about student's
knowledge in computing.\footnote{Worse still, in my view, is that there
  is a pattern of researchers failing to discuss or acknowledge
  students' productive knowledge \emph{even when their own data supports
  it}.} Such a perspective also fails to address what aspects of
students reasoning might get broken by attempts to ``fix'' such
``misconceptions''.

It's important to note that misconceptions research in computing didn't
always treat students' non-canonical knowledge this way. I begin the
next section by backtracing to some of the earliest work on students'
cognitive ``bugs''. There, we find researchers talking more explicitly
about what's useful in students' non-canonical knowledge---a stance
largely absent in modern computing misconceptions research.

\subsubsection{Not all cognitive programming bugs imply a problem with
the
student}\label{not-all-cognitive-programming-bugs-imply-a-problem-with-the-student}

What's curious about Pea's (1986) comments on Bonar \& Soloway (1983) is
that Pea drew different conclusions from the data than Bonar and Soloway
did. Pea (1986) emphasized that when the student thought two
semantically-equivalent assignment statements in PASCAL were different,
it was a problem of egocentrism: ``students assume that there is
\emph{more} of their meaning for what they want to accomplish in the
program than is actually present in the code they have written'' (p.
30). In other words, Pea treated the data as a fairly clear example of a
class of bugs. Specifically, he saw a bug class in which students simply
assumed the interpreter or runtime could infer the code author's shades
of intent.

Bonar and Soloway (1983), by contrast, were less quick to make
inferences either about the nature of the bug or about the intervention
it entailed. Rather than jump to definitive conclusions, they were
circumspect:

It is not clear exactly how to react to the bugs we have uncovered in
novice understanding of programming. In some cases it may be appropriate
to design new languages or constructs. Often, better instruction would
take care of the problem. The intent of our studies is to better
understand the source of the mismatches and misconceptions that cause
novice bugs. Only once a bug is uncovered and understood are we ready to
create a remedy for that bug. (Bonar \& Soloway, 1983, p. 12)

That the authors would even consider developing new constructs or
languages is a noteworthy distinction. Rather than assume in toto that
students with ``bugs'' had wrong knowledge, the authors instead suppose
cognitive bugs have plausible origins worth designing around. Moreover,
they treat students' divergence from canon as \emph{an opportunity for
research to learn from students}. Precisely because students saw
functional differences (cf., diSessa, 1986) in semantically-equivalent
PASCAL statements, Bonar and Soloway (1983) reflect that perhaps
programming languages should be more expressive:

We find it quite interesting that novices seem to understand the role or
strategy of statements more clearly than the standard semantics. Such
roles discussed here include ``counter variable,'' ``running total
variable,'' ``running total loop,'' and ``first, then rest loop''. (See
Soloway et al {[}1982b{]} for a detailed discussion of novice looping
strategies.) Much work in programming languages is concerned with
allowing a programmer to more accurately express his or her intentions
in the program. Perhaps we can learn something from novices here - our
programming systems should support recording the roles the programmer
intends for various statements and variables. (Bonar \& Soloway, 1983,
p. 12)

Again, what's noteworthy here is that rather than treating students'
non-canonical views as a burden for instruction, Bonar and Soloway
instead see them as an opportunity for programming language designers to
make languages better.

That insight---that instruction and design can meet novices where they
are---carries through to their final remarks about studying and
analyzing novice programming knowledge:

The experience and understanding of a novice are available for analysis.
In particular, our results suggest that the knowledge people bring from
natural language has a key effect on their early programming efforts.
Our work suggests that we need serious study of the knowledge novices
bring to a computing system. For most computerized tasks there is some
model that a novice will use in his or her first attempts. We need to
understand when it is appropriate to appeal to this model, and, when
necessary, how to move a novice to some more appropriate model. (Bonar
\& Soloway, 1983, p. 13)

I assume Bonar and Soloway structured their final line deliberately. If
so, their phrasing has three consequences:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Appealing to novice's existing models gets precedence. That is,
  understanding how to leverage novices' existing knowledge comes first
  in research.
\item
  Changing the models novices have comes next, and ``when necessary.''
\item
  They speak of ``how to move a novice to some more appropriate model,''
  which does not necessarily entail the removal or destruction of
  novice's existing models.
\end{enumerate}

Taken together, these points convey a sense of how Bonar and Soloway
view learning and instruction in computing. Instruction explicitly
includes appeals to prior models and knowledge students might already
have. Learning, meanwhile, involves the movement \emph{when necessary}
to more appropriate models of computation. As I explain in the next
section, such a view exactly aligns with a particular branch of
constructivism, where cognition is viewed as the complex activation of
manifold resources for thinking and knowing.

\subsubsection{Examples motivate the need for contextual-sensitivity in
modeling programming
cognition}\label{examples-motivate-the-need-for-contextual-sensitivity-in-modeling-programming-cognition}

Let's begin with two motivating examples. My aim with these examples is
to show how a practicing programmer might employ specific, distinct
conceptual models to reason locally about a piece of code. First,
consider variants of the PASCAL statements from Bonar and Soloway
(1983), where an assignment statement worked to increment a value and
store the result back to that value. In Table 2 below, I imagine five
different ways of writing a programming statement. Each of the five ways
is semantically equivalent to the others (or near enough for explanatory
purposes). And, in each example I add a layer of specificity to the
syntax. I also propose a corresponding interpretation of how I might
apply interpret and think about a given statement.\footnote{I assume for
  simplicity's sake that all variables and functions in my code samples
  evaluate to floating-point numbers that can be combined under
  addition. I also assume, crucially, that all statements are in the
  same hypothetical language and that the variable names connote nothing
  to the computational interpreter.}

\protect\hypertarget{ux5fToc252445956}{}{}Table -- Five different
interpretations of semantically equivalent programming statements in the
same language

\begin{longtable}[]{@{}lll@{}}
\toprule
\begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
Statement
\strut\end{minipage} &
\begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
Programming Syntax
\strut\end{minipage} &
\begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
How I Might Think About It
\strut\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
1
\strut\end{minipage} &
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\texttt{x\ =\ x\ +d}
\strut\end{minipage} &
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
An incremented or accumulator
\strut\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
2
\strut\end{minipage} &
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\texttt{x\ =\ x\ +\ update}
\strut\end{minipage} &
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
Updating the value of \texttt{x}
\strut\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
3
\strut\end{minipage} &
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\texttt{x\_position\ =\ x\_position\ +\ update}
\strut\end{minipage} &
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
Updating the x-position with uniform velocity
\strut\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
4
\strut\end{minipage} &
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\texttt{x\_position\ =\ x\_position\ +\ update(t)}
\strut\end{minipage} &
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
Acceleration in the x-direction
\strut\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
5
\strut\end{minipage} &
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\texttt{x\_position\ =\ x\_position\ +\ update(...)\_}
\strut\end{minipage} &
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
Generalized x-position update, where the update function might be
anything from stochastic noise to low-level hardware functions that get
keyboard or mouse input
\strut\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

I stress that the third column is about \emph{how I might think about}
each programming statement\emph{.} By no means am I making normative
claims about how one \emph{ought} to think about it, or whether the
statement actually does what its names might suggest it does. Rather,
``How I might think about it'' reflects the kind of local meaning or
interpretation I might attach to such a statement when I work with it,
given my understanding of its role and context.

Consider specifically statements 1 and 4. Statement 1 could very well be
an incrementer in some kind of iterative code. If I had to debug code
that employs statement 1, what I might do is exploit my knowledge of
what d is (does it hold a constant value? Is it 1?) and try inserting
intermediate print statements into the iterative code. Doing so, I can
inspect textually how values change with each iteration. In statement
4's case, it could very well be that the position-update code animates
images on a screen. If there's a problem with that code, one of the
easiest ways I might notice is that the acceleration seems off in the
graphics. Consequently, my debugging might call upon knowledge I have
from kinematics. I might try inserting code that draws a dot at the
object's on-screen position with each iteration, leaving a trail I can
visually inspect. I could then look at the path of the object's
trajectory and the spacing patterns between successive dots as a
first-pass test of whether my code achieves the motion I want.

To be clear, it's not just that graphics might improve my efficiency in
debugging a statement. Rather, applying an interpretive frame that
treats an assignment statement as a kinematic position update lets me
\emph{use conceptual knowledge from physics} to diagnose and fix
problems in my code. If I view the assignment statement as saying
something about the motion of an object (cf., Hammer, 1994, p. 165), a
field of knowledge and concomitant techniques from physics becomes
available to me to think with. But, if I focus only on the
semantic-level equivalence of statements 1 through 5, there would be no
obvious reason for me to access what I know about physics in order to
reason about the code.

My second example concerns statements that are syntactically-similar,
rather than semantically-similar. I proposed that the statements in
Table 2 all came from the same language. But, another phenomenon comes
into play when different languages use the same symbology for
conceptually different operations. Table 3 shows examples of what look
like semantically-equivalent operations, but in fact are not.

\protect\hypertarget{ux5fToc252445957}{}{}Table -- Three
syntactically-similar statements with very different semantics

\begin{longtable}[]{@{}llll@{}}
\toprule
\textbf{Statement } & Programming Syntax & Language & How I think about
it\tabularnewline
\midrule
\endhead
\textbf{1} & i = i + 1 & C & \textbf{Increments} i by 1\tabularnewline
\textbf{2} & w = w + ``ly'' & JavaScript & \textbf{Appends} ``ly'' to
the string w\tabularnewline
\textbf{3} & p = p + geom\_point() & R (ggplot2) & \textbf{Composes} a
layer of points onto a plot p\tabularnewline
\bottomrule
\end{longtable}

The catch here is that the plus operator takes on different roles in
different languages because of how those languages define its use.
Statement 1 increments a number in C; Statement 2 appends the letters
``ly'' to a string; Statement 3 adds a layer of points to a statistical
graphics plot. These different kinds of operations become even more
apparent and consequential when, for example, such statements are
repeated. In the R/ggplot2 code below,\footnote{In R, the assignment
  operator can be written as a directional arrow. The symbol
  \textless{}- (``less-than, hyphen'') indicates the value on the right
  side of the symbol is being assigned to the variable on the left side
  of the symbol. As used, the symbol itself is semantically equivalent
  to having written an equals sign (``=``).} I'm using multiple
reassignment statements to compose a statistical graphics plot. The
layered creation of a plot invites a very different kind of conceptual
interpretation than, say, repeatedly accumulating numbers into a running
sum:

p \textless{}- InitializeGgplot\_1w()

p \textless{}- p + GrandMeanLine(owp)

p \textless{}- p + GrandMeanPoint(owp)

p \textless{}- p + ScaleX\_1w(owp)

p \textless{}- p + ScaleY\_1w(owp)

p \textless{}- p + JitteredScoresByGroupContrast(owp, jj)

One obvious reason for thinking about this code with a different
interpretive frame is that numeric addition is commutative; composing a
plot is not necessarily commutative.\footnote{To convince yourself that
  plot composition isn't commutative, imagine a scale function that
  squares up the aspect ratio of a plot and another scale function that
  sets the aspect ratio to 1.5:1. Applying the square function last
  produces a square plot; applying it first produces a rectangular plot.
  String concatenation is also not necessarily commutative. ``cat'' +
  ``dog'' evaluates to ``catdog,'' while ``dog'' + ``cat'' evaluates to
  ``dogcat.''} So, despite the syntactic similarity, reassignments that
compose a plot using reassignment (as above) does not obey the same
rules as reassignments for a running total. But, even within this code
block, statements that look alike perform operations of a different
nature. While some expressions (e.g, ``+ GrandMeanLine(owp)'') add a
visual layer to a plot, others modify features of the plot (e.g, ``+
Scale\_X(owp)'', which adjusts the scales on the plot's x-axis to fit
the numeric range of the data).

Given these motivating examples, it seems sensible to think there's
utility in a programmer having different conceptual metaphors available
to think about and work with code. Example 1 shows that choosing to
apply knowledge from physics to a piece of code can change the cognitive
nature of debugging. There, debugging a position-update statement
becomes, in part, reasoning kinematically about the properties of motion
trails.\footnote{For comparison, consider the argument that ringing an
  aircraft speedometer with physical markers changes the nature of
  cognition when pilots work to land a plane (Hutchins, 1995b).} Example
2 shows that across languages, programmers might have to deploy
different conceptual metaphors to reason about statements in a
locally-consistent way. Knowing that plots in ggplot2 can be composed
layer-by-layer with reassignment is crucial if you're trying to write or
understand code that creates statistical graphics. But, I would argue
that thinking about ◻ = ◻ + ◻ as ``compose new layer onto plot'' can and
does appeal to different kinds of knowledge when compared to thinking
about ◻ = ◻ + ◻ as ``include this addend in the sum,'' which itself can
and does appeal to different kinds of concepts when compared to thinking
about ◻ = ◻ + ◻ as ``increment the counter.''

Stepping back, I can build the following argument

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Programming can be helped by applying conceptual models to code,
  particularly when relevant domain-knowledge structures can
  advantageously transform a problem (example 1)
\item
  But, conceptual models don't work all the time for all statements.
  Because languages are designed differently, the same syntax can
  actually correspond to very different operations in code (example 2).
  And, that's true both within and across languages.
\item
  Consequently, it makes less sense to treat conceptual models as right
  or wrong, and more sense to treat them as differentially advantageous
  for thinking about what a piece of code does. (Thinking a ``+''
  implies numerical addition isn't \emph{globally wrong} in JavaScript,
  but it won't explain why 1 + ``1'' yields ``11'' as a result.)
\item
  It seems plausible that successful programmers, when reasoning about
  or writing code, are able to dynamically access or deploy conceptual
  models that are advantageous given the context (language, syntax,
  surrounding code). Certainly such a supposition is in line with work
  suggesting students have resources for thinking conceptually and
  dynamically about how mathematics models real-world situations (Izsák,
  2004; Sherin, 2001)
\item
  To model how programmers think with conceptual models, a suitable
  framework should be able to account for the dynamic, context-sensitive
  deployment of conceptual knowledge.
\item
  To model how programmers develop expertise, a suitable framework
  should be able to describe higher-order phenomena. Such phenomena
  include explaining how programmers come to have conceptual models or
  generate new ones, why they decide to deploy them, and how programmers
  consider which conceptual model (i.e., which way to think about code)
  is appropriate.
\end{enumerate}

Taken together, these assertions propose criteria for how we might
strive to model cognition in programming. Our modeling frameworks should
be context-dependent, dynamic, and capable of explaining where
conceptual models come from. They should also be able to account for
phenomena that are not themselves conceptual, including what directs the
use of certain kinds of conceptual knowledge. In the learning sciences,
such models already exist and have proven useful and productive for
thinking about thinking.

\subsubsection{Manifold models of cognition explain context-dependence
and the growth of
expertise}\label{manifold-models-of-cognition-explain-context-dependence-and-the-growth-of-expertise}

In 1993, a pair of articles in the learning sciences staked a strong
claim for viewing knowledge as a network of pieces, isolated enough to
be locally triggered but trainable enough to fire in larger concerted
patterns (diSessa, 1993; Smith, diSessa, \& Roschelle, 1993). Informed
in part by agent-based accounts of cognition (Minsky, 1986) and complex
systems models (diSessa, 2002), the central tenets of an ``in-pieces''
approach hold that knowledge is emergent from interacting primitives,
rather than unitary and monolithic. An example from Smith, diSessa, and
Roschelle helps illustrate the point.

The authors show that we might think of a rubber band as a different
conceptual entity depending on context. In several different
situations---wrapped around a newspaper, pulled taut as a string, spun
to store energy in a toy plane propeller---we intuitively think about
the rubber band's physical behavior differently: one as a negligible
part of the newspaper's point mass, another as a transverse pendulum
(likely) obeying Hooke's Law, and the third as a torsional spring. Those
differences in intuitive thinking reflect the contextual dependency of
what we know about the physical world:

In each of the rubberband examples, various pieces of intuitive physical
knowledge describe the mechanism at work: the rubber band binds the
newspaper, grips the jar lid, and acts a source of springiness for the
bobbing object. Although a mapping cannot be made from the rubberband to
scientific entities, it is quite easy to map these qualitatively
distinct physical processes to scientific entities and laws. For
example, instances of binding almost always map to a practically rigid
body. Likewise, gripping maps to friction forces, and springiness maps
to Hooke's law. This suggests that applicability can depend directly on
our intuitive knowledge---knowledge that exists prior to any formal
scientific training (Smith et al., 1993, p. 144).

The in-pieces approach to modeling cognition has been used, among other
things, to explain how experts reason about fractions and decimals
(Smith et al., 1993), how students reason about forces in physics
(diSessa \& Sherin, 1998; diSessa, 1993; Hammer, 1996; Sherin, 2001),
how students construct and evaluate algebraic representations of
physical situations (Izsák, 2004), and how knowledge transfers across
contexts (Hammer et al., 2005; Wagner, 2006). Because its starting
assumption is that knowledge is fragmented, knowledge-in-pieces can
account for wide variations of how people---particularly novices---use
knowledge on a moment-to-moment basis. In other words, because it
assumes knowledge is local, it can still explain the kinds of
globally-inconsistent ways people might reason about physical situations
(diSessa, 1993). As a framework, an in-pieces approach ultimately argues
that models of concept replacement and good/bad criteria for knowledge
should be supplanted by a learning model of alignment/refinement of
prior knowledge and the consideration of knowledge as
productive/unproductive.

Hammer and colleagues have worked to extend the in-pieces approach to
explain how students' epistemological activity---how they orient toward
knowledge and knowing in a context (Hammer et al., 2005; Hammer \& Elby,
2002, 2003). Specifically, those authors use two core theoretical
constructs to explain students' stances toward knowledge and knowing:

\begin{itemize}
\item
  \emph{Epistemological Resources} are the epistemological equivalent of
  diSessa's phenomenological primitives (p-prims). Resources, the
  authors propose, are the atomic units involved in how people cognize
  about the source of knowledge, the nature of knowledge, and
  epistemological activities (Hammer \& Elby, 2003; Louca, Elby, Hammer,
  \& Kagey, 2004).
\item
  \emph{Epistemological Frames} are the emergent result of subsets of
  resources acting in concert. Drawing from both Goffman's (1974)
  sociological notion of frame as structures of expectations and
  subsequent work on framing in discourse (Tannen, 1993),
  epistemological frames are a participant's local answer to the
  question ``what is it {[}specifically, what knowledge activity{]}
  that's going on here'' (Goffman, 1974, p. 8).
\end{itemize}

Resources can frames can interact in activity settings to produce
larger-scale patterns called ``epistemological coherences'' (Rosenberg,
Hammer, \& Phelan, 2006) where evidence from data suggests that a
network of discrete cognitive units can nonetheless give rise to stable
cognition.

A useful conceptual metaphor for resources and framing is to think about
a lecture hall with different sets of lights: a spotlight in the back to
highlight a lecturer, chalkboard lights, house lights, and a projector.
Each light has an individual brightness, but the lights are only
controllable at the per-bulb level. In that sense, they are atomic. But,
light patterns interact with one another to create a field of lighting
for the room. Thus, lights are a bit like resources, and different light
configurations can correspond to different, sociologically-stable uses
of the room.

\protect\hypertarget{ux5fToc252445958}{}{}Table -- Using room lighting
configurations to think about epistemological framing

\begin{longtable}[]{@{}lllll@{}}
\toprule
\textbf{What is it that's going on?} & Houselights & Spotlight &
Chalkboard & Projector\tabularnewline
\midrule
\endhead
\textbf{On-stage monologue} & Low & ON & Low & OFF\tabularnewline
\textbf{Presenting slides} & Low & ON & Low & ON\tabularnewline
\textbf{Working through equations} & Low & ON & High &
OFF\tabularnewline
\textbf{Students discussing with each other} & High & OFF & Low &
OFF\tabularnewline
\textbf{Watching a movie} & Low & OFF & Low & ON\tabularnewline
\textbf{Cleaning the room after a movie} & High & OFF & High &
OFF\tabularnewline
\bottomrule
\end{longtable}

To convince yourself of the sociological stability of these
configurations, imagine you were watching a movie when suddenly the
house and chalkboard lights came up to full intensity. It would jar you
out of the experience. You might wonder whether something was wrong: is
there an emergency? Should you evacuate?

The metaphor isn't perfect. Ontologically, for example, resource and
framing theory propose these constructs exist not in the world but in
the minds of individuals. But, the metaphor is quite useful for
understanding how individual elements---in this case lights---can work
in concert to create and sustain a stable frame. And, crucially for a
cognitive system, the metaphor lets us account for variation in
activity. The same lecture hall can become a window onto one person's
thoughts (monologue), a site for instruction (working through equations
at the board), a shared space for collaboration (student discussion), or
a place in need of repair (cleaning up) simply by varying the
intensities of lighting banks in particular ways.

An example helps ground this in-pieces approach to epistemology. In
Russ, Coffey, Hammer, and Hutchison (2008), the authors describe the
situation where an elementary student reasons about why an empty juice
box collapses when you suck on the straw. One student gives what the
authors deem to be an excellent mechanistic account of why the juice box
collapses. But, as the teacher seems to steer the discussion toward
vocabulary---in this case, ``pressure''---the student clearly pulls back
from her mechanistic reasoning, seems much more diffident, and claims
that pressure is hard to explain. That example highlights the disconnect
between doing science as knowing vocabulary and doing science as
reasoning mechanistically. Moreover, it strikingly highlights that a
student who by all accounts produced an excellent explanation of how
pressure works was left nonetheless with the impression that pressure
was hard to explain.\footnote{In fairness to the teacher, I'm trying to
  focus on the result of the interaction and far less on the intent the
  teacher might have had. The student still left with a sense that
  science might be about vocabulary, even if that's not the view the
  teacher would espouse or was trying to enact in the moment.}

\subsection{Methods and Theoretical
Commitments}\label{methods-and-theoretical-commitments}

\subsubsection{Student population, course background, and selection
}\label{student-population-course-background-and-selection}

This paper focuses on Electrical Engineering (EE) students from Flagship
State, a large public research institution in the mid-Atlantic. The
students I studied were taking Intermediate Programming, the second
semester of an introductory programming course taught in C. The course
was exclusively for EE majors and taught by EE faculty, and its
enrollees were typically first-year or second-year EE majors. Its model
was two 75-minute lectures and a 1-hour teaching assistant-led
discussion section for students each week. Course topics included:

\begin{itemize}
\item
  Strings
\item
  Pointers
\item
  Dynamic Memory Allocation
\item
  Testing
\item
  Debugging
\item
  Hash tables
\item
  Trees
\item
  Linked Lists
\item
  Abstract Data Types
\item
  Functional Decomposition
\end{itemize}

Students in the course had varying degrees of experience with
programming. Partly, that variation is because students with Advanced
Placement (AP) Computer Science credit could place out of Basic
Programming, the first semester C course. So, some students in
Intermediate Programming came from technical magnet schools where they
may have already had one or more years of programming in multiple
languages, while other students may have been first-time programmers
with only one semester of programming experience: Basic Programming.

I studied the same course, Intermediate Programming, during the fall
2011 semester and the spring 2012 semester. During fall 2011 I
ethnographically observed over 50\% of the course lectures, and during
spring 2012 I sat in on several discussion sections. At the beginning of
each semester I solicited student participants for semi-structured
clinical interviews. Of the students who responded to the solicitation,
I scheduled interviews opportunistically with students given my
resources as the sole researcher on the project. In total, I worked with
a cohort of 6 students in fall 2011 and 4 students in spring 2012.

My specific interest was in how students design computer programs. By
that, I mean I wanted to know not just how students program, but whether
and how they structured programs, how they did (or did not) try to
incorporate modularity into their programs, and how the final form of a
program reflected what they had learned about how to manage complexity
through software. In the sections that follow I outline the palette of
methods I used to pursue those questions.\footnote{The language in
  sections 3.3.2 and 3.3.3 is taken with slight modification, from
  ``Studying students' early-stage software design practices,'' a paper
  I authored, along with William Doane, and submitted to the 2014
  International Conference of the Learning Sciences (ICLS).}

\subsubsection{Studying design as a complex phenomenon of disciplinary
practice}\label{studying-design-as-a-complex-phenomenon-of-disciplinary-practice}

Practicing professionals make complex use of talk, gesture, and
representational artifacts in physics (Gupta, Hammer, \& Redish, 2010;
Kaiser, 2005; Ochs, Gonzales, \& Jacoby, 1996); chemistry (Stieff,
2007); field biology (Hall, Stevens, \& Torralba, 2002); civil
engineering (Stevens \& Hall, 1998); structural engineering (Gainsburg,
2006); mechanical and electrical engineering (Bucciarelli, 1994;
Henderson, 1999); and architectural design (Hall et al., 2002; Hall,
1999). Given that:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Science and engineering education research has made progress by
  looking for continuities in how novice learners develop disciplinary
  practices (Gupta et al., 2010; Hall \& Stevens, 1995; Smith et al.,
  1993; Stevens \& Hall, 1998), and
\item
  Emerging research on software engineering reveals that early-stage
  software design involves complex inscriptional, discursively, and
  epistemic practices,
\end{enumerate}

it seems striking that there is no contemporary body of research,
comparable to studies of expert practice, that looks at students'
software design practices. In other words, we have every reason to
believe that expertise in software design involves complex practices,
but little (if any) research on what productive capacities students have
for them. Finding those productive design capacities requires a shift
from questions such as \emph{how can we assess and mitigate students'
difficulty in programming?} toward questions such as \emph{how do
students learn and display evidence of design thinking in programming?}

Rephrasing research questions asked of professional software engineers
(Petre et al., 2010, p. 533) and instead treating \emph{students} as
designers, I ask:

\begin{itemize}
\item
  What do students actually do during early stage software design work?
\item
  What does students' exploratory design thinking look like?
\item
  How do students communicate?
\item
  What sorts of drawings do students create when they design software?
\item
  What kinds of strategies do students apply in exploring the vast space
  of possible software designs?
\end{itemize}

\subsubsection{Deploying methods to capture the complexity of
early-stage design
work}\label{deploying-methods-to-capture-the-complexity-of-early-stage-design-work}

The methods below form the core of my developing program to study
students as software designers. None of the methods below are new; all
have been used in prior educational research. What \emph{is} new, I
believe, is the opportunity to combine them all under the umbrella of
understanding what happens when students design software. For each of
the four students in my spring 2012 cohort I captured multiple streams
of data.

\begin{itemize}
\item
  I collected their \textbf{code history}. By this, I mean I preserved
  frequent snapshots in time of what students' code looks like. Research
  has already shown code snapshotting to be a useful method for
  understanding large-scale patterns of student error (Jadud, 2006;
  Rodrigo et al., 2009; Spacco, Hovemeyer, et al., 2006). And, the
  resolution of snapshots is extremely fine: Spacco et al. are able to
  capture the contents of a file each and every time a student saves it
  to disk. But, that research takes an aggregate view: it identifies
  large-scale error patterns at the expense of detailed naturalistic
  understandings of why students make those errors. Moreover, it's
  primarily used to identify what \emph{mistakes} students make, which
  is distinctly different from a research orientation that considers the
  negative \emph{and} positive consequences of students' design choices.
  A currently untapped advantage of collecting code history data, then,
  is that while we historically use it to aggregate \emph{across
  programmers} it nevertheless also gives us fine-grained individual or
  team-based histories of how designs evolve.
\item
  I conducted \textbf{clinical interviews} with them. Clinical
  interviews have proven historically useful in understanding the
  substance of students' knowledge and the nature of conceptual change
  (diSessa \& Sherin, 1998; Duckworth, 2006; Ginsburg \& Opper, 1988;
  Ginsburg, 1997). Crucially for Computer Science Education (CSEd)
  research, clinical interviews can tell us about students'
  epistemologies---how they view knowledge and knowing in a discipline
  (Hofer \& Pintrich, 1997)---which can affect how they approach and
  adopt that discipline's practices (Hammer, 1989, 1994; Lising \& Elby,
  2005). Moreover, because my interviews were videotaped and analyzed
  from perspectives of interaction analysis (Goodwin, 2000; Jordan \&
  Henderson, 1995), they offer rich evidence of the substance of
  students' design practices
\item
  I analyzed their in-interview \textbf{inscriptions} when they designed
  --- what they wrote, how they wrote it, and how those writings got
  used. Evidence from both science studies (Hall et al., 2002;
  Henderson, 1999; Hutchins, 1995a; Kaiser, 2005; Latour, 1990; Ochs et
  al., 1996) and educational research (Hall \& Stevens, 1995; Lehrer,
  Schauble, Carpenter, \& Penner, 2000; Stevens \& Hall, 1998)
  highlights the centrality of inscriptions to disciplinary practice in
  science and mathematics. Ethnomethdological data from studies of
  professional software engineers supports the same result:
  inscriptional practice is central to how professional engineers design
  software (Rooksby \& Ikeya, 2012; Rooksby, 2010). And, since part of
  the inscriptional environment when designing software is the computer
  itself, I captured and analyzed what happened on-screen as students
  design programs.
\item
  I paid close attention to students' in-interview \textbf{gestures}.
  Perspectives of gestural analysis hold that gestures can not only
  support or extend thinking, they can also communicate entirely
  different information than what's being said (Goldin-Meadow, 2003).
  Moreover, perspectives from cognitive anthropology and embodied
  cognition studies argue that bodily motion \emph{is itself} cognition
  (Hall \& Nemirovsky, 2012; Hutchins, 1995a; Nemirovsky, Rasmussen,
  Sweeney, \& Wawro, 2012). For example, when students describe a part
  of code by moving their hand across an imaginary row of items and
  tapping each item, their bodies convey information we can interpret
  about how they understand iteration.
\end{itemize}

Figure 3-1 presents a visual overview of some of these methods and modes
of analysis. In particular, the first (top-left) panel depicts how we
deploy these data collection methods during a clinical interview:

\begin{itemize}
\item
  a voice recorder captures speech (often a redundancy in case another
  recording device fails)
\item
  a LiveScribe Pulse pen digitizes written inscriptions, allowing us to
  play back what was written in time
\item
  a videocamera records data for knowledge analysis, interaction
  analysis, and gestural analysis,
\item
  an in-interview computer tracks code history and its screen-capturing
  software records real-time activity.
\end{itemize}

\includegraphics[width=6.25556in,height=4.20000in]{media/image4.jpeg}

\protect\hypertarget{ux5fToc252445967}{}{}Figure ‑ -- An comic-based
overview of my methods for capturing students' design practices

\subsubsection{Developing a revelatory case of students' early-stage
design
work}\label{developing-a-revelatory-case-of-students-early-stage-design-work}

In the empirical work that follows I explore what productive knowledge
and resources students have for structuring programs. I begin with
Lionel, a male student in Intermediate Programming, fall 2011. First, I
propose the idea that Lionel connects ``designerly'' (Archer, 1979)
stances across contexts. I analyze his retrospective account of
modifying a bike and compare it to his retrospective accounts of how he
programs in Intermediate Programming. Then I detail Lionel's multi-stage
cascading approach to structuring a program solution in an interview.
The argument throughout is that from bike design to code design Lionel
evinces knowledge and stances that productively help him solve problems.

Next I transition to Rebecca, first explaining the epistemological
difficulty she felt when coding, then showing how she circumvented that
block by structuring a program piece using gesture and talk. I argue
that what Rebecca and Lionel have in common are resources for handling
design obstacles as they emerge. Where the two students differ is in
their in-the-moment judgments about what sorts of knowledge and
activities are appropriate to work through a design problem.

I argue that taken together, Rebecca and Lionel constitute what Yin
(2009) calls a ``revelatory case,'' by which I mean my investigation
arises from ``an opportunity to observe and analyze a phenomenon
previously inaccessible to social science inquiry.'' In my case, I argue
the phenomenon I'm observing is what engineering students do in the
early stages of design work on a program. Its inaccessibility is
demonstrated by the almost complete paucity of studies that draw on
real-time (e.g., video records) of how students begin work on a complex
program. Contrary to being ``representative'' or ``typical,'' my
revelatory case does not aim to generalize. Rather, its value is in
phenomenological richness. As Yin (2009) writes, and I think others
including Erickson (1986) would agree, such case studies are worth doing
``because the descriptive information alone will be revelatory.''

\subsection{\texorpdfstring{\protect\hypertarget{ux5fToc247188553}{}{\protect\hypertarget{ux5fToc252445926}{}{}}Lionel's
approach to design and to
programming}{Lionel's approach to design and to programming}}\label{lionels-approach-to-design-and-to-programming}

\subsubsection{While modifying a bike Lionel decomposed tasks, explained
solutions to himself, and created intermediate design
representations}\label{while-modifying-a-bike-lionel-decomposed-tasks-explained-solutions-to-himself-and-created-intermediate-design-representations}

Prior to entering my study, Lionel had already earned a Bachelor's
degree in Business Information Technology at a different university. We
began the interview by discussing why he re-matriculated to earn an
electrical engineering degree.

Interviewer: How did you, um end up coming out of your business program
and your business experience and deciding that electrical engineering
was what you wanted?

Lionel: Um, well, since I was Business Information Technology, basically
all my job was doing was writing SQL code and, you know, short
statements like that. I also worked at a help desk for---I guess I
worked at Department of Homeland Security. And, pretty much besides the
SQL statements I worked with this one program and whenever somebody had
a problem with it they'd come to me. And basically that---that's what my
whole degree was---was going down that track and then that's really not
what I wanted to do: just sitting at a desk, you know people coming to
me being like ``oh, you gotta do this for me.'' So, I was like, alright.
Or, I have to fix a problem for them, pretty much. And that's really
what I didn't want to do. I wanted to do something more hands-on where I
could, you know, I'd be given a project where I'd do something on my
own, create something on my own. And, um, also my Dad and my uncle were
both electrical engineers. And my brother became a computer scientist.
Cu---so I've always been around, you know, that type of environment,
where my dad's always building something and, you know, I always see it
and I'm like, you know, that's very interesting. I'd like to do
something like that too. So it's really just uh, a mixture of bad
experience with my previous job and being around company that likes to
do that sort of stuff. (Interview 1 of 1, October 17, 2011)

Lionel apprenticed building things with his father first, taking on
small, highly-directed roles in projects.

Lionel: Be-before I started my degree, you know, I really didn't know
how to do any of this stuff. So, I would just kinda help my dad out a
little bit. He'd be like ``Oh, here,'' you know, ``to, to do this you
have to do exactly this.'' You know, ``solder this here,'' or, you know,
``drill this,'' blah blah blah. He'd tell me exactly what to do. So, I
just kind of followed along. (Interview 1 of 1, October 17, 2011)

Presumably in those projects Lionel was at a periphery and direction
came from his father. More recently, though, Lionel had struck out on
projects of his own. As Lionel said, ``now that uh, my dad kinda showed
me how to do certain typical types of things, I've tried to become
\textbar{}\textbar{}more adventurous\textbar{}\textbar{} \textbar{}\{air
quotes\}\textbar{} /Sure/ and do things on my own'' (Interview 1 of 1,
October 17, 2011).

One of Lionel's recent adventurous projects was adding a motor to a
beach cruiser bicycle. The gist of the project was simple: using a
low-priced motor he found online, Lionel realized he could mount the
motor to an inexpensive bicycle and make a motorized bike. With one
exception, Lionel was solely responsible for the project from start to
finish.\footnote{Lionel told me that the only part his father helped him
  with was showing him how to use a Dremel power tool to smooth, sand,
  and cut edges (Interview 1 of 1, October 17, 2011).} He bought the
parts, he mounted the motor, and he maintained it as a working vehicle.
But, in practice the project was not straightforward. In the next
subsections I show at length how Lionel addressed the emergent
challenges of modifying\footnote{Often called ``modding'', for short.}
his bike with a motor. My analysis focuses on three features of Lionel's
story that I argue carry over to his programming approach:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  When faced with ambiguity, Lionel worked to decompose large tasks into
  smaller subtasks he could understand.
\item
  Throughout the task, Lionel was metacognitive about the state of his
  design. His constant refrain was ``\emph{how} am I gonna make this
  work?''
\item
  Before committing work and materials to a solution, Lionel created
  intermediate designs whose suitability he could evaluate.
\end{enumerate}

I argue in later sections that these three features of Lionel's
approach---decomposing subtasks, explaining procedures to himself, and
explicitly planning prior to committing time and resources to a final
product---have strong resonances with the approach Lionel took to while
programming.

\paragraph{Lionel had to decompose Step 1, which simply said ``Install
the
Engine''}\label{lionel-had-to-decompose-step-1-which-simply-said-install-the-engine}

Lionel bought an inexpensive motor online with the plan of attaching it
to a cheap bike. As he discovered, though, frugality came with a hidden
cost: poor instructions. Attaching the motor to the bike was not at all
a straightforward process.

Lionel: it was funny. When I, when I got the motor /Mmmhmm/ Um, oh,
it---like I said it was like cheaply made and whatnot and it came with
instructions, but um, no exaggeration at all, step 1 said ``install the
engine''. And that was it. \{Laughs\}

Interviewer: Huh

Lionel: And then step 2 was, um, I guess it was somethin like ``put the
gas tank on.'' It really didn't tell me anything. It was kinda like
``OK, those are the obvious steps, but *how* do I do that? So I kinda
had to figure stuff out on my own. /So/ yup

Interviewer: Did it---how did it end up going? Was it a lot of trial and
error? Like, did they even have pictures of how it was supposed to look
when it /There, there/ was installed?

Lionel: There are maybe two or three pictures, but they were black and
white pictures printed out on like a, I guess like a crappy printer. So
they were kinda hard to see. Um, so I guess my first step was to go look
online at better quality pictures. Unfortunately nobody really---I was
hopin to find you know, like a YouTube video: this is how I did it.

Interviewer: Mmmhmm.

Lionel: I couldn't really find somethin like that, so I just looked at
pictures. (Interview 1 of 1, October 17, 2011).

Lionel recognized that having an attached motor is an obvious goal
state, but that goal state was poorly-specified by ``crappy pictures''
that were hard to interpret. Moreover, the states in between start and
goal were undocumented and not obvious. Lionel's subsequent search for
materials that showed a step-by-step progression came up empty. Faced
with a poorly-specified goal-state and no intermediate guidance,
Lionel's response was to force himself to think carefully about how he
would proceed.

\paragraph{In making modifications Lionel frequently asked himself ``how
am I gonna make this
work?''}\label{in-making-modifications-lionel-frequently-asked-himself-how-am-i-gonna-make-this-work}

Faced with the challenge of interpolating between ``crappy pictures,''
Lionel thought hard about what to do.

Lionel: I spent, I guess the, I guess the first five hours I just spent
kinda staring at it like, ``hmm, *what* am I gonna do?'' Like, ``*how*
am I gonna make this work?'' (Interview 1 of 1, October 17, 2011)

Ultimately, he was struggling with more than just poor instructions.
Even if he did figure out how the motor was supposed to be mounted, he
still faced another problem: the motor was designed for a mountain bike.

Lionel: I had to make a few modifications on my own. Because I guess
it's built for uh, a mountain bike. And the tubes on mountain bikes are,
y'know like *that* thin \{makes circle with right thumb and index
finger\}, but on my beach cruiser there's one tube \{widens thumb and
forefinger\} that's pretty thick where I had to mount the engine.

Lionel: And since the engine mount was built for a motorized bike it was
really small. So I kinda had to like devise my own plan on *how* to,
y'know, add a bigger mount.

Interviewer: Mmmhmm

Lionel: So, that was, that was my first problem. That was, y'know the
first step was to put the motor on I'm thinkin, well, how am I gonna put
the motor on if, y'know the, the specs don't even align. So, I guess
spent y'know five hours just thinkin', like, ``what am I gonna do?''

Lionel: Finally I came up with a plan that seems to be working. So. And
then, I just kinda went from there. I'd put it together, look at the
picture and say ``OK, this piece looks like that piece, and y'know I
guess it looks like it attaches that way, so I'm gonna try that.'' And
just put it together and, y'know if it fit, it fit. If it didn't, then
I'd, y'know I'd be like ``OK, maybe that was wrong. Lemme, lemme work on
somethin' else on it.'' So. (Interview 1 of 1, October 17, 2011)

Asking himself \emph{how} to do something helped bring sub-problems into
focus. Within a larger mod task (viz., adding a motor to a bike), Lionel
in effect entered a nested task: modding a mountain bike engine mount to
fit a beach cruiser. Modding the mount itself was entirely undocumented,
so Lionel's solution in part was to look for affordances in the physical
structure of parts to see what he could attach. Crucially, the
environment afforded Lionel rapid, iterative feedback. It wasn't hard to
tell when a design choice worked because ``if it fit, it fit.''

\paragraph{Lionel used intermediate representations to consider
alternative
designs}\label{lionel-used-intermediate-representations-to-consider-alternative-designs}

Lionel grew increasingly frustrated as he saw how difficult it would be
to mount the motor. Crucially, part of his tactic around that
frustration was to resolve to keep trying and thinking of possible
solutions.

Lionel: So, I guess I had to give myself, like a few hours just to cool
down and then actually think ``OK, well maybe, you know I've already
bought it, so, I'm gonna continue on with it. So I had to think, you
know, OK, what's my next step. *How* could I make this work? You know,
I---I thought of a few different ideas. You know, I'd think of one idea
and think ``OK, would this work? It might, but let me think of another
idea.'' So I'd come up with another idea and say, ``would this work?''
You know, so once I got a few ideas I would just kind of, in my head
picture it. Say, ``OK, this is how it would mount, or this is how it
would mount. Which one would be safer? Or will one of these not even
work?''

Interviewer: Right.

Lionel: So, I guess after, you know, awhile of deciding which one was
better, you know, which one would be easier and safe at the same time
/mmhmm/ so then, that's when I made my decision, and just kinda, went
for it to see if it would work. And luckily it did \{laughs\}.
(Interview 1 of 1, October 17, 2011)

Before committing to a design, Lionel forced himself to evaluate
alternatives first. Naively, he could have gone with the first idea that
came to him. Instead, even if he thought an idea might work he pushed
himself to consider others first. Then, with a collection of possible
designs he evaluated what he saw as their suitability in his head. He
applied criteria as he evaluated, trying to balance the ease with which
a mounting scheme might work with a concern for the safety of the
user---in this case, himself. Again, at least in Lionel's telling, those
steps happened \emph{before} committing to a final implementation. In
other words, Lionel generated and evaluated multiple configurations with
an eye toward both the builder of and user of those configurations
before settling on a path to take.

In sum, I note three key features of Lionel's approach to the design
challenges of modifying his bike. First, he repeatedly asked himself
``how am I gonna do this,'' forcing himself to think in terms of a plan
or procedure to fill in the gaps between the instructional pictures.
Second, he proposed ideas, tried to evaluate their feasibility, and
continued pressing for alternate solutions (``would this work? It might,
but let me think of another solution''). Third, so much of his planning
took place before he ever mounted the motor. By his account, he fiddled
with pieces, envisioned multiple ways the engine could mount, and
evaluated their feasibility before ever taking concrete steps to mount
the motor.

\subsubsection{Lionel's approach to programming resonates strongly with
the design stance he held when working on his
bike}\label{lionels-approach-to-programming-resonates-strongly-with-the-design-stance-he-held-when-working-on-his-bike}

\paragraph{Lionel thinks about steps at a
chalkboard}\label{lionel-thinks-about-steps-at-a-chalkboard}

Later in the interview I asked Lionel to compare his experiences in
Basic Programming with those he was having in Intermediate Programming.
His response revealed a great deal about how he starts projects in
Intermediate Programming: he starts at a chalkboard.

Lionel: Well, {[}Basic Programming{]} I already learned the basics of
the class so I know the basics of where to start my code. And then,
obviously the projects in this class {[}Intermediate Programming{]} are
harder than last class, but I have the basics of, ``OK, well, I know how
a program runs. So how can I translate that into my program, for, for
{[}Intermediate Programming{]}?'' And, yeah, so I guess I'll sit down,
like, yeah, and I'll have like a chalkboard and I'll write things out,
and say OK well first, you know, \{sweeps hands{]} not even think about
the code. Just, in \{air quotes\} English words, so you know, my
project's gonna do this. You know, first I need to make this
calculation, then I'm gonna print this. And then I'm gonna make this
calculation. Et cetera. And I'll sit down and write that out, in the
order that it needs to be done. And then from there I'll go back and
think, ``OK, what's the code, to, you know, make this calculation or
print that?''

Interviewer: So, when you start a project, you don't start at a
computer, it sounds like you kinda start at your chalkboard.

Lionel: Right. Exactly. Yeah, cuz I've tried to start a project at a
computer, and you know I'll write a---I'll start writing a few lines of
code, but then I just really zone in on what I'm doing and kinda lose
sight of the whole picture.

Lionel: So then I'll write this code for, you know, this one little
section of the project, and then after I've done that I'll look back and
say ``Oh, yeah, I need to write the whole project,'' but then this
little code doesn't actually fit into my project, so it's kinda, you
know I guess I get tunnel vision when I start at a computer. (Interview
1 of 1, October 17, 2011)

For Lionel, the chalkboard is a pivotal object in his design activity.
It holds the top-down ``English words'' description of his program. In
so doing, it serves as the earliest durable representation of his
program's intended hierarchy and relationships. That function is
important, because without it Lionel is apt to ``lose sight of the whole
picture'' of what his code is supposed to do. Moreover, as the next
section reveals, the board continues to help Lionel orient his activity
as he digs into the specifics of how his program will work.

\paragraph{Lionel copies pseudo-code into the computer even though it
won't
compile}\label{lionel-copies-pseudo-code-into-the-computer-even-though-it-wont-compile}

As Lionel described the role of the chalkboard, I initially thought
there was a distinct separation between the chalkboard and the computer
as activity objects. The chalkboard, I assumed, was where Lionel
sketched out design plans while the computer was where he wrote code.
When I probed Lionel based on my hunch, I discovered Lionel's activity
wasn't cleanly separable into design on the board / code on the
computer.

Interviewer: Did you, when you start thinking about the code, do you end
up usually filling it in on your board, or will you actually go straight
over to the computer and do it?

Lionel: Um, on my board, I'll do, I guess like
\textbar{}\textbar{}pseudo\textbar{}\textbar{} \textbar{}\{air
quotes\}\textbar{} code /mmhmm/ uh, you know, first on the board it
would only say ``this function calls this function, this function calls
this function.''

Interviewer: Mmmhmm

Lionel: So then on the board I'd write down, ``alright, well this
function'' um, I dunno, for instance I'll, I'll say you know, I'm workin
with a function that checks to make sure that it's on the---that I'm
moving *on* the board between spaces 1 and 24.\footnote{At that point in
  the semester, Lionel's class was working through a Backgammon project,
  so the spaces and moves he's referring to pertain to programming a
  basic Backgammon game.} So then, you know I'll write down like,
pseudocode, write ``OK, well I'm gonna start. I'm gonna have a for-loop
that will go from 1 to 24, and then here I'll, you know I'll have an
if() statement that says if it's between you know, 1 and 24 it's valid,
if not it's invalid.'' So, that'd be my pseudocode, and then when I
actually---I'd, I'd try to write that out for the entire program, or as
much as I could.

Interviewer: Filling out the tree, you mean.

Lionel: Right, fillin out the tree. And then I'd go to my computer, so
that at least I already had a baseline of what to work with. And then,
on my computer I'd, you know I'd write out the pseudocode, I'd ac---I'd
literally write out the pseudocode, even though obviously it wouldn't
compile and actually work.

Interviewer: Mmmhmm

Lionel: So, then, from the pseudocode, as, as I read down on my computer
what the pseudocode said, I'd actually try to write the actual code that
would work. (Interview 1 of 1, October 17, 2011)

Lionel blurred my initial chalkboard / computer distinction in two ways.
First, his overall ``English words'' plan of functions calling functions
evolved on the chalkboard to include computational idioms, including
if-statements and for-loops. But, then, according to Lionel, he would
copy that pseudo-code as-was into the computer, knowing full well it
wouldn't compile. These observations---that Lionel writes computational
pseudo-code on the board, then copies that same code into the
computer---may seem trivial, but I contend they aren't.

Because of the choices Lionel makes, neither chalkboard nor computer has
a distinct, privileged kind of syntax. Rather, ``English words'' and
pseudo-code can peacefully co-exist on the chalkboard, and non-working
pseudo-code gets deliberately typed into the computer before it's
expanded out into ``actual code.'' Put another way, Lionel is capable of
distinguishing between ``English,'' ``pseudo-code,'' and ``actual
code,'' but all three forms of expression are a part of his design
process. It's perfectly acceptable for him to have something other than
``actual code'' exist in intermediate representations of his design.

\subsubsection{Gestural, inscriptional, and verbal in-interview evidence
reveals Lionel's resources for structuring a
program}\label{gestural-inscriptional-and-verbal-in-interview-evidence-reveals-lionels-resources-for-structuring-a-program}

To probe Lionel's retrospective account of how he programs, I gave him
an in-interview task to work on. In this section, I argue Lionel's
development process on that task moved rather linearly through what I
call a ``representational cascade.''\footnote{As I'll explain later, I
  think it would be a mistake to assume that all development activity
  proceeds linearly and top-down in the manner Lionel's does. Not only
  do I think Lionel's clean linearity is rare, I also think it's not
  normative. So, my point here isn't to stress that Lionel moved
  linearly, monotonically through the cascade. Rather, my point is to
  show that he moved between different representational levels \emph{at
  all}.} My schematic for that cascade --- along with the features and
affordances in each layer of the cascade --- is depicted in Figure 3.
What I show here is that actually, the final code Lionel produces is in
fact one representation of a procedure that actually existed in
different forms and modalities before it became the final code in Figure
3.

In the analysis that follows, I explore two of those modalities---verbal
description and written pseudo-code---and explain why each is crucial to
understanding the final produced code. Ultimately, I argue there is a
need to explain \emph{how} Lionel creates and moves fluidly between
these markedly different cascade layers. Specifically, I attempt to
answer the question of \emph{Why does he see the upper layers (verbal
code, hand-written pseudocode, and handwritten source code) as
legitimate activity?} My candidate answer is that Lionel's approach to
programming is in part stabilized by epistemological resources that
allow him to treat the upper layers of the cascade as valid, productive
knowledge expressions in programming. Such an answer is strongly in line
with the findings from sections 3.4.1, where we see Lionel deliberately
making use of other intermediate design representations.

\includegraphics[width=5.74444in,height=4.31111in]{media/image5.jpeg}

\protect\hypertarget{ux5fToc252445968}{}{}Figure ‑ -- An overview of
Lionel's representational cascade and the features and affordances of
each representational layer.

\paragraph{The range-finding prompt and Lionel's final source
code}\label{the-range-finding-prompt-and-lionels-final-source-code}

About 35 minutes into our interview, I gave Lionel the programming task
reproduced below:

Suppose you want to write a program that finds the range of a set of
numbers. So, the program would take 10 numbers as input, then it would
compute the range as the difference between the largest number and the
smallest number.

How would you write such a program? (Interview prompt, October 10, 2011)

The prompt was deliberately vague about certain constraints. It did not
prescribe \emph{how} the program should get input (Should the programmer
expect the user might type it in, or feed in a formatted file?). It also
did not prescribe how the program should provide the result (should the
programmer print the result on screen? Store it to a file? Return the
result to be used by another function?) The reasons for vagueness were
in part because I wanted to see what assumptions students might make
during their design process. Some solutions, I reasoned, might decouple
the procedures of getting input, computing range, and returning output.
Others might simply hard-code a procedure for getting keyboard-based
input. I wanted to prescribe no specific approach up front if I didn't
have to. I have reconstructed his original source code from
screen-capture footage of his in-interview programming:

\includegraphics[width=5.74444in,height=7.11111in]{media/image6.emf}

\protect\hypertarget{ux5fToc252445969}{}{}Figure ‑ -- Lionel's final
source code for the range-finding prompt

Let's explore some top-level features of Lionel's code. First, Lionel
chose to have a user enter text directly using the keyboard (lines
9--12), and his program displays the range result as printed text (lines
28--29). His strategy for finding the maximum and minimum is to
implement a procedure that iterates through the array of numbers and
compares the current number to the globally stored max and
min.\footnote{I think of this as a ``king of the court'' procedure. The
  current value stays on as an extreme if and only if it beats the
  current ``king'' for that extreme. Otherwise, the challenger is
  replaced and the contest resumes with a new challenger until there are
  no more challengers.} If the current number (array{[}i{]}) ``beats''
the global min (by being lower than min), its value replaces that of the
global min. Similarly, if the current number exceeds the global max, its
value replaces that of max. Before the procedure starts, Lionel assigns
the first number of the array as the value for both extrema.

\paragraph{Lionel's verbal description of the
program}\label{lionels-verbal-description-of-the-program}

Lionel chose to start the task on pencil and paper. After scribbling a
line about scanning input into an array (which I discuss in a later
section), Lionel stopped and began explaining out loud to me.

Lionel: so I'm thinkin, my general concept is I'm going to parse through
this array, check each number /Mmmhmm/ and I'll have two variables, min
and max, and, first I'll set both min and max to the first variable in
array, er, yeah first number in array /Mmmhmm/ and, I'll parse through
the rest of the array, checking to see if,

Lionel: Umm, you know if---I'll s---I'll start with max, for example.
/Mmmhmm/ And I'll check to see if the number in that section of the
array is bigger than max. /uh-huh/ And if it is, I'll set max to that
new number. /OK/ And then I'll do the same thing for minimum. And then
at the very end, I'll have my minimum and max and then I'll just
subtract `em.

Interviewer: OK

Lionel: So that's my main concept. (Interview, October 17, 2011)

This verbal description is in some ways bound by the least stringent
rules of expression. If we're hard-nosed about syntax, Lionel's speech
stops and starts (line 5 into line 6), which leans on the unspoken
understanding that ideas in talk are under negotiation. Stopping
mid-sentence and leaving that sentence unfinished is a valid and
acceptable move in conversational space. Moreover, there is a strong
mapping between Lionel's temporal language and the intended
flow-of-control for the program. ``First I'll set,'' ``then I'll do,''
``and then at the very end'' all describe when instructions will be run.

Finally, Lionel's instructions begin at the high level, and in some
cases implementation details are completely left out. For example,
Lionel does not specify precisely how he'll ``parse through the array.''
At this stage of the cascade, it seems sufficient for him to explain
\emph{that} he plans to parse through, but now how. What Lionel
\emph{does} do is gesture while describing ``parsing''. This gesturing,
I argue, creates a concrete representation of iteration and
incrementation --- the gestural marker of a loop.

\includegraphics[width=5.74444in,height=3.23333in]{media/image7.png}

\protect\hypertarget{ux5fToc252445970}{}{}Figure ‑ -- Lionel makes a
cycloid/helix gesture as he says ``parse through this array''

Contrast how Lionel specifies ``parsing'' through the array with how he
describes ``checking'' its contents. First, Lionel explains that he'll
``check each number'' (line 2), decides that he has underspecified how
``checking'' happens (line 5--6), and finally makes ``checking''
relatively more explicit (lines 7--8). Once checking for the maximum has
been made explicit (lines 7--8), Lionel says ``and then I'll do the same
thing for minimum'' (line 9). Line 9 thus becomes the verbal equivalent
of a repeated, parameterized procedure. It is, in a sense, the
talk-space seed of a function. Conversationally, both participants
understand that Lionel saying, ``then I'll do the same thing for
minimum'' (line 9) refers to a procedure defined elsewhere in the
conversation, not a new set of instructions.

\paragraph{Lionel's pseudo-code}\label{lionels-pseudo-code}

Immediately following the verbal description above, Lionel begins
creating what he calls his ``pseudo-code,'' narrating it as he writes.
There are four key features in Lionel's talk (below) that I discuss in
my analysis. Below is an overview of each feature and what importance I
think it has:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Lionel iterates over his initial verbal description by fleshing out
  some implementation specifics --- structures that further specify
  \emph{how} something in a program is to be done.\footnote{It's a
    limitation of my writing style that in this document I use
    ``iteration'' both to describe aspects of Lionel's design process
    and as a formal term of art in computer science. In both cases,
    though, the conceptual meaning is the same: the incremental,
    rule-based set of transitions to the state of an object.} This
  iteration is important because it suggests that his subgoals --- and
  criteria for satisficing them --- change from one micro-coherence to
  the next.
\item
  Reconciling what Lionel \emph{narrates} with what Lionel \emph{writes}
  reveals peculiar mismatches. The mismatches suggests a productive
  capacity: that Lionel can \emph{background} certain details within a
  micro-coherence by omitting them from the written code while still
  maintaining an understanding of how his program will work. Put another
  way: the inscriptional object in a microcoherence captures what Lionel
  thinks is important within that microcoherence, but the inscription
  does not necessarily represent the totality of his conceptual
  understanding.
\item
  At one point in his code, Lionel writes ``same for min,'' which we can
  reasonably infer means he intends to repeat a procedure for
  calculating the minimum. That line encapsulates the precursors of
  functional abstraction: Lionel not only recognizes procedures that
  share the same structure, he can background the specifics (point 2) of
  code he repeats.
\end{enumerate}

Lionel: So it'll be scanf, um, and I'll have my for-loop, and I'll just
have my int-i, and I'll go from zero uh, zero to, four actually cuz
it'll be zero, one, two, three, four, yeah, zero to four. /Mmmhmm/

Lionel: Umm, and I'll have, oh, up top here, to set the, right---right
after the scanf /Mmmhmm/, I'll just say min and max equal the
first---mm, hiccup, sorry

Interviewer: It's OK.

Lionel: \{laughs\} the first number in the array. So I'll say min and
max equal array of zero. And that's just setting them so that I have
something to compare to, I guess. /uh-huh/ I think that will work. Yeah.

Lionel: So, so then, as I parse through my array of five variables, I'll
say ``if min,'' er no, I'll say ``if array of i, you know of, of that,
if, uh, the number in the array that I'm looking at is greater than max,
then max equals that number, array of i'' /Mmmhmm/ and then I'll do the
same thing, same for min. And then at the very bottom, I'll have, uh,
range equals max minus min.

Lionel: So that's my pseudo-code, just so that *I* understand what's
goin' on in my head.

\includegraphics[width=5.75000in,height=3.00208in]{media/image8.emf}

First, Lionel's talk iterates over his initial verbal description by
fleshing out \emph{implementation details} --- programming structures
that further specify \emph{how} something is to be done. Compare here
how Lionel instantiates ideas of iteration in Table 1 below.

\protect\hypertarget{ux5fToc252445959}{}{}Table -- Comparing Lionel's
words and actions across different stages of ``programming'' over the
same putative section of his program. I have bolded some
language-specific syntax.

\begin{longtable}[]{@{}lll@{}}
\toprule
\textbf{Design Coherence} & What he says & What he does\tabularnewline
\midrule
\endhead
\textbf{Verbal Description} & ``I'm going to parse through this array,
check each number'' &
\includegraphics[width=2.22222in,height=1.48889in]{media/image9.jpeg}\tabularnewline
\textbf{Narrated pseudocode} & ``I'll have my \textbf{for-loop}, and
I'll just have my \textbf{int-i}, and I'll go from zero uh, zero to,
four actually cuz it'll be zero, one, two, three, four, yeah, zero to
four. &
\includegraphics[width=2.23333in,height=0.35556in]{media/image8.emf}\tabularnewline
\bottomrule
\end{longtable}

In his narrated pseudo-code, he introduces the C-specific syntax of
``for-loop'' and ``int-i,'' which handle iteration and indexing,
respectively. Moreover, he counts in order to specify the index bounds
of the loop (zero to four).

Second, it's actually somewhat complicated to address the question ``how
do Lionel's inscriptions align with what he says?'' He \emph{says}
``I'll have a for-loop'', but the visual evidence shows that he hasn't
bounded his for-loop body in curly-braces.\footnote{As a language, C
  typically uses curly braces to bound instructions that span multiple
  lines. So, a typical for-loop would be written like this, where the
  loop preamble is bounded by parenthesis and the loop body is bounded
  by curly braces (bolded):

  for(int i=0; i \textless{} maximum\_value; i++) \textbf{\{}

  // Do something

  \textbf{\}}} If he had done so, his for-loop preamble would end with
an opening curly brace; it doesn't (Table 1 graphic above). Also, for
matching, his for-loop should have concluded with a closing curly brace;
it doesn't. Nevertheless, his verbal evidence --- particularly him
saying ``int-i'' --- suggests that he understands details that variables
need to be declared. Together, these mismatches suggest that features
(viz., denoting iteration and setting its bounds) are important enough
to be included in this representation. Other specifics, including proper
declaration of variables, are things Lionel understands but does not
insist on writing.\footnote{One might argue that Lionel is simply being
  sloppy and forgetting to include details, even though he said them.
  The point I'd make here is that even if he is being sloppy, I would
  think it unfair to say, for example, that Lionel has conceptual issues
  with declaring variables. His talk (and many later iterations of his
  design) suggests he understands that declarations are necessary for
  working with variables in C. Moreover, even if he is being sloppy on
  the particular point of variable declarations, the overall character
  of this pseudo-code is still starkly different from that of the full,
  working source code solution he'll ultimately produce. So, even if we
  propose that Lionel has some unintended tolerance --- such as
  sometimes overlooking missing variable declarations --- for assessing
  whether his pseudo-code satisfices, there is still strong evidence
  that his criteria for evaluating his work in pseudo-code stage are
  markedly different from the criteria he'll apply to later iterations
  of design.

  no one would confuse the pseudo-code skeleton here for Lionel's
  fleshed-out, fully-working code later. So, even if Lionel \emph{meant}
  to put them in and forgot, he still judged this pseudocode as
  satisficing his subgoal} Together, and combined with later evidence of
Lionel's successful working program, these observations suggest that
ignoring the details is actually productive for Lionel.

Third, Lionel's talk reveals productive capacities for dealing with
design work. For example, he describes the guts of his program through
brief perspective-taking (lines 14 -- 18 above). As a reminder, Lionel's
procedure is what I refer to as a ``king of the court'' approach: if the
number currently-being-inspected exceeds the reigning champ, the champ
is dethroned and replaced by the number currently-being-inspected. But,
the way Lionel articulates that process actually uses multiple senses of
``I,'' where it's at times actually not clear whether ``I'' means ``I
the programmer'' or ``I the program.''\footnote{To keep your head from
  exploding, I remind you that the lower-case version of ``i'' in
  Lionel's talk actually refers to the subscript for indexing an array.}
For example, the I in ``I'll say `if the array of i'' could very well be
the programmer, where by ``I'll say'' he means ``I the programmer will
write a conditional test.'' But, that phrase is immediately followed by
``if, uh, the number in the array that I'm looking at is greater than
max'', where I believe we may be seeing Lionel taking the perspective of
the program itself.\footnote{My point here is that technically, Lionel
  can never inspect an array element himself; he can only write code
  that does that. You might disagree, in which case we get to have a
  really fun discussion over beer about how we think programming is a
  case of ``distributing cognition'' (Hall, Wright, \& Wieckert, 2007).}
If anything, the seemingly untroubled, interchangeable use of ``I'' here
might suggest that to do the work of programming one can occupy a hybrid
space of both being the program and being the programmer at the same
time.

Lionel concludes this procedure with the invocation ``same for min.''
Picking up the thread of hybridity, this phrase could mean, ``I the
programmer will write a similar procedure for min'' or ``I the program
will perform similar operations to determine min.'' That Lionel leaves
his pseudocode there (``so that's my pseudocode'', line 20) suggests
that whatever perspective it might be written from, this designed
artifact satisfices his pseudo-code production subgoal. And, that
satisfaction is remarkable because if ``same for min'' is an appropriate
place-holder, it suggests Lionel has ways of seeing that procedure as
\emph{similar enough} in structure to the procedure already laid out as
to not require further specification. This is an important point: if a
precursor to functional abstraction is recognizing repeated code, Lionel
has actually recognized repeated code \emph{before ever writing it}. As
we'll see later, he ends up writing this procedure out explicitly, which
means that whatever answer we might come up with as to why Lionel
doesn't abstract this to a function \emph{can't} be that he doesn't see
the repeated computation; to the contrary, repeating core parts of the
max computation was built into his design from the start.

\subsubsection{\texorpdfstring{\protect\hypertarget{ux5fToc247188557}{}{\protect\hypertarget{ux5fToc252445930}{}{}}We
can trace continuities in Lionel across different kinds of design
activities}{We can trace continuities in Lionel across different kinds of design activities}}\label{we-can-trace-continuities-in-lionel-across-different-kinds-of-design-activities}

We can trace patterns in Lionel's design activity by asking questions
about the intermediate design artifacts he creates. Specifically, we can
pay attention to \emph{what} representations he creates, \emph{where}
they live, and \emph{how} he uses them. Table 3 outlines those
relationships for the three contexts discussed: Lionel's bike modding
project, Lionel's retrospective account of his programming practices,
and Lionel's in-interview work on the range-finding task. Table 3 helps
support three key points:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  There is a specific consistency to Lionel's design activity across
  contexts. All activities---given the limits of Lionel's
  self-reports---\emph{involved the generation and use of intermediate
  design representations}. Those representations may at times have lived
  in Lionel's head, on paper, or in electronic format.
\item
  A thorough account of Lionel's in-interview programming activity
  cannot and should not ignore those intermediate representations and
  how he used them.
\item
  Because intermediate representations are a feature of all three design
  contexts, it seems likely that what stabilizes their creation and use
  is tied to something deeper or more global than just conceptual
  knowledge in a domain such as computer programming.
\end{enumerate}

\protect\hypertarget{ux5fToc252445960}{}{}Table -- Comparing Lionel's
design activity across contexts

\begin{longtable}[]{@{}llll@{}}
\toprule
\textbf{Context / Question} & Modding the bike & Retrospective
programming account & In-interview task\tabularnewline
\midrule
\endhead
\textbf{\emph{What} intermediate representations did he create?} &
Possible configurations for mounting the engine & Relationships of
functions to each other; pseudo-code for how functions would work & A
verbal description; narrated pseudo-code; hand-written pseudo-code;
hand-written source code\tabularnewline
\textbf{\emph{Where} do the intermediate representations live?} & In
Lionel's head & On a chalkboard and, later, in the text editor & In
Lionel's talk and gestures; on paper; in the text editor\tabularnewline
\textbf{\emph{How} did he use those intermediate representations?} & To
compare mounting configurations for ease and safety & To ``see the whole
picture'' of a project & To ``understand what's goin on in my head''
when structuring his code\tabularnewline
\bottomrule
\end{longtable}

\subsection{Rebecca's approach to
programming}\label{rebeccas-approach-to-programming}

In this section, I analyze data from a different student --- Rebecca. I
offer this data as an example of what developing design expertise might
look like. If Lionel's approach to programming uses high-level work
(e.g., his pseudo-code) to cascade down the specifics of syntax, it
seems reasonable to ask whether other students use pseudo-code in the
same way. And, if they do not, it seems worth asking how uses of
pseudo-code differ and what that difference might tell us about
students' approaches to programming.

In Rebecca's case, the data I'll present begins with her stuck on the
third of four projects in the course: the ``iTunes'' project.\footnote{In
  this project, students were given a series of text files ---
  representing the user's music collection --- that contained formatted
  albums and song titles. The task was to create a program to read in
  the music information, create a database representation of it, and
  allow the user to transact with that database. Typical transactions
  might be asking for a listing of all the albums in a user's library or
  creating a playlist by allowing the user to choose specific songs.} My
analysis will expand on the following points:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  To understand Rebecca's struggle on the iTunes project and her
  approach to programming during that struggle, we need to understand
  some of her prior experiences in programming. In our fourth interview
  Rebecca recalls a pivotal project experience from her Basic
  Programming course. At the cost of hours of work, half her codebase,
  and a good grade, Rebecca discovered that her sense of how her program
  should work did not match what was expressible in C. I analyze
  Rebecca's recollection because it sits at the intersection of
  self-efficacy and disciplinary practice. Rebecca lost confidence and
  felt ``demoralized'' because her attempts at expressing a procedure
  did not match the constructs and underlying workings of the C
  language. Moreover, the fallout from that experience persisted into
  Intermediate Programming. By our fourth interview she had considered
  dropping Intermediate Programming entirely because of her low grades.
  ``My *biggest* problem'' in programming, she said, ``is I think I have
  the logic, my logic just doesn't transfer to code. Or I don't know
  *how* to transfer it correctly'' (Interview 4 of 5, April 6, 2012).
\item
  Rebecca has the capacity to make sense of program designs. As she
  tries to understand why the instructor would prescribe an array of
  pointers to store track titles in the music server, Rebecca makes
  progress by reasoning about what she knows about pointers and arrays.
  Like Lionel, Rebecca can use gestures, talk, and inscriptions (viz.,
  pseudocode) to express how she might intend for her program to work.
  Rebecca also has a grasp of the conceptual issues --- from a
  computational and programming perspective --- relevant to the part of
  her design she's struggling with. But, Because Rebecca's approach to
  programming so strongly orients toward producing valid, working
  syntax, it may be limiting Rebecca from seeing other ways in which she
  can make progress on her design. In other words, Rebecca \emph{can} do
  things like write pseudo-code to plan, but she doesn't tend toward
  those behaviors when she's stuck. I'll offer an explanation based on
  epistemological dynamics to explain why that's the case.
\end{enumerate}

In what follows, I first present point 1 to give readers a sense of how
Rebecca's prior experiences may inform her approach to programming. I
then outline the particular part of the iTunes project where she feels
``stuck,'' discussing point 2. Before I jump into that analysis, I
\emph{strongly} urge readers who are not familiar with pointers or
dynamic memory allocation to refresh themselves on the subject.

\subsubsection{\texorpdfstring{\protect\hypertarget{ux5fToc247188558}{}{\protect\hypertarget{ux5fToc252445932}{}{}}Rebecca's
struggles on the iTunes project make sense in light of some prior
programming experiences she
had}{Rebecca's struggles on the iTunes project make sense in light of some prior programming experiences she had}}\label{rebeccas-struggles-on-the-itunes-project-make-sense-in-light-of-some-prior-programming-experiences-she-had}

At the time of this study Rebecca was a first-year electrical
engineering major who had gone to high school in rural Maryland. Her
university courses in Basic and Intermediate Programming were the first
experiences she had doing any programming at all, let alone in the C
language. Though she felt academically at ease in high school, a tough
project in her first-semester Basic Programming course shook her
confidence in programming. In the data presentation and analysis that
follows, there are four sub-points I'd like to make about Rebecca's
development in programming.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Prior to the semester I worked with her, Rebecca had difficult
  experience trying to understand arrays on a Basic Programming course
  project.
\item
  That difficult experience had two pieces to it that I believe relate
  to personal epistemology:

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \item
    Rebecca came to feel that her idea for a solution wasn't something
    expressible in C.
  \item
    When she got help from a friend who was ``very smart in
    programming,'' she felt like she typed the code he told her to into
    her project without really coming to understand it.
  \end{enumerate}
\item
  Reflecting on it in our first interview, Rebecca pinpoints the
  experience of that array project as demoralizing and having a lasting
  effect on her confidence in programming.
\item
  Epistemological issues continue to factor into Rebecca's confidence in
  programming. By our fourth interview, Rebecca was getting low grades
  in Intermediate Programming. She had strongly considered dropping the
  course. Trying to articulate her ``biggest problem,'' Rebecca
  explained that she felt like she understood the theory behind the code
  examples in class but had difficulty transferring her logic into code.
\end{enumerate}

\paragraph{Rebecca had a difficult experience on a Basic Programming
arrays
project}\label{rebecca-had-a-difficult-experience-on-a-basic-programming-arrays-project}

In our first interview, I asked Rebecca about how her experiences in
Basic Programming compared to those she was having in Intermediate
Programming. She explained that in that first semester, she ``kinda got
lost on one part'' --- arrays --- ``that was key to the rest of the
year.''

Rebecca: {[}I{]}n {[}Beginning Programming{]} we had three projects, and
the first one, uh, we did---I did discuss it with other students, and we
all worked together kind of. But that one I felt much more like I had a
grasp on everything I was doing. Whereas the second project, I would
stare at the code and be like ``ohmygosh, I don't even understand what,
like, it's asking me to do.'' Like, or how to make it do what it needed
to do, which was to like store in numbers in arrays, change them, and
add `em together and stuff. (Interview 1 of 5, February 10, 2012)

I asked Rebecca if she felt she had a better grasp by the time she
turned in the project.

Rebecca: No. Cuz, there was a friend on my floor, who---very smart at
programming, he's had like years of experience. And he gave me a lot of
pointers and tips doing it, so I felt like I was just using him a lot
instead of actually learning it, which was a downfall for me.

Interviewer: Huh. So, how, I mean. I feel like sometimes you could---you
might ask somebody for help, and then they show you something and then
you might be like oh, you've learned it now /yes/, but it sounds like
you're saying /no/

Rebecca: Just, I don't, like, I don't blame him at all because he helped
me, but, it was kind of more of a, he showed me how to do it, I was like
\{mimes exaggerated typing, in an almost singsong voice\} *OK, sure,
I'll type that code*, but I don't really, I never really, I don't know I
never made the connection. (Interview 1 of 5, February 10, 2012).

\paragraph{Rebecca's ``terrible'' array experience has epistemological
components}\label{rebeccas-terrible-array-experience-has-epistemological-components}

In our second interview a week later when we discussed debugging,
Rebecca returned to that experience in her Basic Programming class:

Rebecca: My array project, my---the second project we did last semester,
that I had a big error. I had to delete like half the code I worked on.
Like, just cuz, like I had said last week, I didn't do very well on
arrays. So.

Interviewer: Oh. Was it cuz of one thing, or was it like a repeated---

Rebecca: It was kind like a general, like, what I was trying to do
didn't work with C.

Interviewer: Do you remember what it was you were trying to do?

Rebecca: I don't know exactly, but uh, I just remember, like, whatever I
was tryin to do, cuz I had my friend help me, he's like ``you can't do
this this way,'' like, it just, C doesn't recognize whatever I was
trying to do.

Interviewer: Huh. So that was like, it sounds like it's almost kind of a
case where your, your plain English of what should happen /yeah/ can't
actually translate---it's not even that you're not sure how it's---

Rebecca: It's like, it was impossible.

Interviewer: It was never---

Rebecca: It was never allowed.

Interviewer: There is no word for that /yeah/ in this language.

Rebecca: I \{laughs\}, that was back then, so. That was always fun.

(Interview 2 of 5, February 17, 2012)

Epistemologically, two patterns stand out in Rebecca's recollections.
First, as Rebecca remembers it, one of her core problems was that her
visions for how her program should accomplish things were ``impossible''
and ``never allowed'' in C. Second, in order to get past that obstacle,
she ultimately typed code a friend told her to without feeling like she
understood it. As this analysis unfolds, I think these two points give
us a crucial starting point to understand Rebecca's practice. The
experience underscored a gulf between ideas Rebecca wanted to invoke and
the syntax she needed to realize them in C. And, her way of dealing with
that gulf was to make an end-run around it by taking and using code she
didn't fully understand.\footnote{I say this without any intended
  judgment (or pejorative connotations) about what Rebecca did. Given
  the assessment structure of the course, where 90\% of a student's
  grade came entirely from whether the program worked to spec, she may
  have felt that getting a good grade by using code she didn't
  understand was preferable to getting a bad grade with code that didn't
  work.}

It's worth noting that whether her friend's code worked to
spec\footnote{Here, I'm using ``worked to spec'' as a shorthand for
  ``worked in such a way that it passed all of the input/output tests
  the instructor would run to determine 90\% of a student's grade.''} is
immaterial to the two epistemological issues I raise. First, if the code
she copied worked, Rebecca would get a good grade, but she wouldn't
understand how her project functioned. And, if it didn't work, she would
get a bad grade \emph{and} she wouldn't understand how her project
functioned. Ultimately, no matter how the code performed, Rebecca would
likely be in a tough spot if she ever had to maintain, revisit, or
refactor the code she copied. Second, no matter how her copied-code
project functioned, it represented \emph{someone else's} articulation of
how a procedure should go. That is, the copied code was an artifact of
someone else bridging the gulf between high-level idea and articulation
in C. Rebecca still found difficulty traversing that gulf.

\paragraph{Rebecca's array project experience had a lasting effect on
her confidence in
programming}\label{rebeccas-array-project-experience-had-a-lasting-effect-on-her-confidence-in-programming}

As that interview drew to a close, I wrapped it up with what had become
standard protocol: asking whether there was anything else she wanted to
talk about.

Rebecca: I think I'm good, I---just

Interviewer: Anything else on your mind?

Rebecca: Sorry I'm not the greatest programmer \{laughs\}

Interviewer: No, hey hey hey. That's \{shakes head\}, first of all, I
mean, um, how come you're apologizing?

Rebecca: I dunno. It's a study, and I'm probably not the best test
subject. (Interview 2 of 5, February 17, 2012)

Rebecca's spontaneous apology spawned a discussion about why she feels
diffident in programming. For her, that diffidence traced its way back
Basic Programming. Feeling lost and stuck on that array project was more
than just a tough experience; it was a turning point:

Rebecca: When I got lost I think is when I started losing that
confidence thing. And then I never really got it back, so.

Interviewer: When was it you started to get lost?

Rebecca: Halfway through last semester, like with the arrays in the
second project. So\ldots{}. I guess it was like, demoralizing, kind of,
and, like, like I don't want to sound cocky or arrogant or anything, but
in high school I did very well in academics. And, so, I never really,
like anything that I had to work for, I worked for for a short period of
time, got it, and like I'd understand it, and if I didn't, like it was
really easy for someone usually to explain it to me. And, I felt like I
didn't really get that after I lost everything on that second project, I
was like, I wasn't able to build myself back up out of that, which
probably stems to why, not as confident in the class. (Interview 2 of 5,
February 17, 2012)

As I interpret how this array project experience may be playing a part
in Rebecca's later approaches to programming, I stress that my analysis
is about Rebecca's perceptions and recollections.\footnote{Due to
  deliberate study limitations I don't have access to the actual code
  from that project. To be as conservative as possible in our data
  collection and respect the privacy of participants, we asked
  participants to grant us access only to folders that would contain
  their Intermediate Programming projects. Outside code/data, including
  code from prior semesters, was excluded from the scope of our code
  snapshot data collection.} It's possible that her ideas for how her
project should work wouldn't have been implementable in any language,
let alone in C. And, while I think that possibility is unlikely, I have
to concede that I only know what Rebecca recalled; I wasn't present for
the experiences she's describing. Nevertheless, the question for my
research is not ``can we know what really happened in that experience?''
Rather, my question is, ``how might Rebecca's \emph{image} of that
experience --- with its specific epistemological facets --- be playing a
part in her approach to programming now?''

Part of the answer to that question may be in her quote above. She talks
about losing her confidence and ``never really {[}getting{]} it back.''
Later, she says ``I felt like I didn't really get that after I lost
everything on that second project.'' Because of the peculiarities of
speech, I had a difficult time parsing Rebecca's statement about losing
everything on that second project. It wasn't clear to me whether there
was a pause between ``get that'' and after, but the meaning of ``that,''
and in turn the meaning of the sentences, depends on the antecedent of
``that.'' So what she said could have two consequentially different
meanings:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  ``I feel like I didn't really get \textbf{that}. After I lost
  everything I lost everything on that second project, I was like, I
  wasn't able to build myself back up out of that.'' In this version,
  ``that'' may mean the topic---arrays---or the project as a whole. I
  might then interpret Rebecca's speech as saying this project was a
  turning point because of the particular topic. Her being ``unable to
  build myself back up out of that'' reflects a kind of debt that might
  have piled up when later assessments depended cumulatively on that
  early topic. Her failure to grasp arrays was thus cumulatively
  punishing because arrays formed a core part of many projects that
  followed.
\item
  ``I feel like I didn't really get \textbf{that} after I lost
  everything on the second project. I was like, I wasn't able to build
  myself back up out of that.'' In this version, ``that'' may mean her
  ready insight into topics or her facility in understanding when
  friends explain them to her. The distinction I draw in this second
  interpretation is ``that'' refers to a kind of continuity in her
  identity---someone for whom school always came easy. The array project
  disturbed that continuity because it was an instance in which things
  no longer came easy to her. And, the experience may have been so
  troubling that she was never able to re-establish being able to easily
  grasp things in programming. The result, then, is that she couldn't
  build herself---in the sense of her continuous identity as one who
  quickly grasps school topics in general and programming topics in
  particular---back up after the array project.
\end{enumerate}

Distinguishing between these possible meanings of Rebecca's speech might
seem pedantic. But, I think the bifurcation is important for theorizing
about learning and instruction in computing. If Rebecca's difficulty was
about the topic of arrays, a plausible narrative is that the difficulty
of that concept---and the assessments that continued to depend on
it---caused a run of bad grades, and the bad grades shook Rebecca's
confidence. But, if Rebecca's difficulty comes from troubling her sense
that she has a ready understanding of programming, the attack on her
confidence is more complex. It's not just that she doesn't get arrays,
it's that she feels she doesn't readily understand other new code she's
seeing in the course---even code that doesn't strongly rely on arrays.
In this view, the confidence Rebecca speaks of isn't just a product of
getting good grades; it's part of a feedback loop in which grades,
confidence, and a feeling of ready insight into programming are all
components.

That potential coupling of grades, confidence, and feelings about ready
insight into code matters because it further pushes the discussion
toward epistemological considerations. It is why, for example, I was
paying such careful attention to Rebecca's recollection of copying down
code her classmate told her. Instead of helping her confidence in
programming by potentially boosting her grade, the copying may have
diminished her confidence because she couldn't understand why someone
else's code worked.

\paragraph{Epistemological issues continued to interact with Rebecca's
confidence in Intermediate
Programming}\label{epistemological-issues-continued-to-interact-with-rebeccas-confidence-in-intermediate-programming}

Rebecca started our fourth interview with a warning to me that she might
``rant'' about programming.

Interviewer: What would you rant about? Go ahead.

Rebecca: Ohh god. Um, I am not doing well \{laughs\} in the class right
now.

Interviewer: OK

Rebecca: And it's not good, so. Yeah, like, we just had our last project
due, which I did a lot better on than the first project

Interviewer: Uh-huh

Rebecca: Which is making me really happy. Um, but, everything about
programming just upsets me, because there's the two guys, like I had
mentioned before who, uh, help me

Interviewer: Mmmhmm

Rebecca: Uh, one of them, he---like I worked every day from the day we
got this project on this project, and my project still didn't turn out
perfect. He wrote his twice in one day because he lost the whole thing.
And, his was perfect. And, I dunno---it just irks me *so* bad that that,
it can be that much of a disparity between people and programming.
(Interview 4 of 5, April 6, 2012)

In her telling, Rebecca wasn't the only person upset by the relative
ease with which this student programmed.

Rebecca: we didn't say it to him, but some of us were like, it's, we
just found it like, so, like aggravating. Like, cuz there's me and
another girl who, on my floor who, she's uh, never had programming
experience either---

Interviewer: Uh-huh

Rebecca: And we're just like it gets so aggravating that it comes easy,
like to him, and that he does it well, but, that's just---he ju---he is
very good at school. He's very smart. So. (Interview 4 of 5, April 6,
2012)

Note here that Rebecca is connecting the idea that programming came easy
to this student to the notion that he was ``very good at school.'' In
this moment, the statement seems to suggest Rebecca has a \emph{fixed
mindset} \{Dweck and colleagues\} about programming expertise. While it
would be naïve to say Rebecca believed --- globally --- that programming
came easily to those who were ``good at school,'' it's nonetheless true
she saw a discrepancy between how easy programming was for her and her
friend and how easy it was for this ``very smart'' classmate.

As we kept talking about this discrepancy, Rebecca offered further
evidence that struggling with programming troubled her academic
identity. In high school, she ``didn't have to study a lot'' because ``a
lot of things came relatively easy'' (Interview 4 of 5, April 6, 2012).
But, her experiences in Intermediate Programming pushed her into a new
position. As she said,

now I'm on the other end and I'm realizing how aggravating it
is\ldots{}. And, I don'---I guess it's just a shock, cuz I do not like
not being good at things \{laughs\}. So...'' (Interview 4 of 5, April 6,
2012).

Again, Rebecca's comments stress the frustration of feeling like
programming comes easily to some people but not to her.\footnote{Or, at
  least, at that moment, in that semester, it did not seem to be coming
  easily to her.} And, they are charged with emotion. She speaks of
being ``on the other end,'' connoting an idea of a structure whose poles
have people who ``get'' programming on one end and those who don't on
the other. It's ``aggravating'' to be on that end, and a ``shock'' to be
there for the first time.

I asked Rebecca to reflect on whether she could ever remember seeing
classmates in the position she was now, where something seemed to come
easily to Rebecca but not to them.

Rebecca: Um. I mean, there was a---there's a friend of mine, who, from
my high school, who, I, I mean just always got math. And, uh, she was
struggling in math a lot.

Interviewer: Mmmhmm

Rebecca: But her attitude was, uh, ``I don't need math. I'm gonna be an
English major.''

Interviewer: OK

Rebecca: So, she didn't even care about it

Interviewer: Oh, OK

Rebecca: Whereas, like, I was like, I'm an electrical engineer major, I
actually need this. So. Like that's the only time I think I've ever felt
like someone's had that feeling that I probably did.

Interviewer: Right.

Rebecca: And, I mean I'm sure there are people that I---uh, that I
didn't know, or didn't realize they had that feeling. But, it's just,
interesting to be on the other end. \{6 second pause\}

Rebecca: Umm. Just. Ugh. I've like, debated withdrawing from the class.
And, like, I've talked to three different advisors or whatever, and
they're like ``don't do it. It's not---you can still bring your---you
can bring your grade up. And even if you don't you have freshman
forgiveness,\footnote{In the interview, Rebecca explained that Freshman
  Forgiveness was a policy extended to students during their first 24
  credit-hours at the university. If a student gets a grade she's
  unhappy with, she can retake that course and her second final grade
  will replace the first.} so.''

Rebecca, unlike her friend, couldn't afford \emph{not} to care about
Intermediate Programming; it was a required course for her electrical
engineering major. But, the combination of feelings she had---consisting
partly of a feeling that programming did not come easily to her---pushed
her to think about withdrawing from the course anyway. And, Rebecca was
not in the habit of withdrawing from classes. Intermediate Programming
was the first and only class where she was experiencing this level of
difficulty (Interview 4 of 5, April 6, 2012).

It would be tempting to think Rebecca struggled in Intermediate
Programming because she didn't work hard enough. But, by her own
assertions she \emph{did} put a considerable amount of effort on
projects. On Project 2, for instance, she said ``I worked every day from
the day we got this project on this project, and my project still didn't
turn out perfect'' (Interview 4 of 5, April 6, 2012). Figure 5 below
shows Rebecca's activity over time from March 5, the day Project 2 was
handed out, to March 28, the day it was due:

\includegraphics[width=5.73333in,height=3.98889in]{media/image10.emf}

\protect\hypertarget{ux5fToc252445971}{}{}Figure ‑ -- Rebecca's
compilation activity over time for Project 2. Each compile a student
initiates creates a commit, so long as there has been a change to the
underlying code. The height of each bar maps to the number of commits
recorded that day. The red line charts \emph{cumulative} commits over
time. The red line is steepest in periods of frequent activity and
shallowest in periods with little or no compile activity.

From the data we can tell Rebecca was compiling Project 2 code almost
every day between March 19 and March 26. Moreover, roughly two thirds of
compilations for the entire project happened in the last 3 days leading
up to the deadline. So, at the very least our data shows she was
actively compiling her code for 8 out the last 10 days.

Rebecca registered no commits between March 12 and March 19. There were
also other days for which no commits were recorded. One possible cause
for those zero values is our data collection system was faulty. A second
possible explanation is that Rebecca was not working. But, a third
explanation is that Rebecca was working but not compiling. This third
explanation actually seems most plausible. It fits the pattern of
Rebecca's work on project 1, where she wrote dozens of lines of code
before ever compiling (Interview 2 of 5, February 17, 2012). It also
accounts for Rebecca's lack of compilations despite her claim that she
worked on Project 2 every day.

So, we can be fairly certain Rebecca was spending time on
projects.\footnote{The specific nature of \emph{what} Rebecca's work
  looked like on Project 2 is the subject of the first study of this
  dissertation, so I don't expand upon it substantially in this study.}
Another possibility for why Rebecca was struggling might be that she
wasn't studying. To address that conjecture, we have to step back to
when I asked Rebecca whether other classes --- classes other than
Intermediate Programming --- had ever pushed her close to withdrawing
from them. She said no.

Rebecca: the thing is, I *don't* know how to like, study for
programming. Cuz like, I look at code and stuff, and I try, and I work
every day on it, but like, even when---I dunno, it just, does not come
naturally.

Interviewer: So, what is it like when you study it? What do you usually
try to do?

Rebecca: Uhh, a lot of times, like, I'll read through the notes we took
in class, like, cuz, what we're doing now---we have a project, uh, we're
doing linked lists, and like, malloc'ing data, which is, uh, *you*
storing it in, as you go. And, like, I get the theory and everything
behind it, and, like it makes sense what it does, but, I just like try
and look at his code examples and stuff, and try and replicate it, but I
don't know what's happening right now, but it's not doing it correctly.
I just, like, my *biggest* problem, I realized, in programming, is I
think I have the logic, my logic just doesn't transfer to code. Or I
don't know *how* to transfer it correctly.

Interviewer: So, um. when you're spending time studying, um, do you
spend most of it on your computer trying to actually program or---

Rebecca: A lot of it is on my computer and a lot of it is also trying to
just look at his code and notes in class. Uh, that's probably about
split, 50/50. Uh, because, sometimes I'll just end up staring at my
computer screen, like, ``what do I even type in to try?'' So, I try and
go look at his examples, so.

(Interview 4 of 5, April 6, 2012)

There are two striking features of this exchange.\footnote{There are
  actually three striking features. The one I left out above is that
  Rebecca describes malloc'ing as \emph{you} storing in data. I think a
  great deal of Rebecca's difficulty with malloc specifically hinges on
  what she means by ``you'' storing it in.} The first is that Rebecca
felt she didn't know how to study for programming. The second striking
feature of this exchange is how Rebecca elaborated on what it might mean
for programming to come naturally to her.

Rebecca feels programming doesn't come naturally to her because she
routinely finds herself unable to transfer her ``logic'' to code. As a
follow-up to Rebecca's answer, I asked what ``logic'' meant to her in
this case.

Rebecca: Like, like how to get the---like, if I need to get a certain
data, like I have a process of how I want to get it, I ju---and I'm
like, ``Oh, you would just do this,'' but can the *computer* do that?
Can I---and it's my---me being able to tell the computer to do that is
where I get lost. (Interview 4 of 5, April 6, 2012)

The tension Rebecca is describing is one I see as strongly
epistemological. First, it's epistemological in the sense that it
reflects a struggle to articulate knowledge of \emph{how to do
something}. Papert (1980) and Abelson \& Sussman (1996) would describe
such a struggle as a pertaining to \emph{procedural epistemology}: the
struggle concerns a particular kind of knowledge work in which the goal
is not to describe what is, but rather how to. The second sense in which
it's an epistemological struggle is closer to how Hammer, Elby, and
colleagues use the word epistemological: what does students' activity
reveal about how they orient toward knowledge and knowing in a
discipline (Hammer et al., 2005; Hammer \& Elby, 2002, 2003; Scherr \&
Hammer, 2009). In this second sense, we can observe that Rebecca defines
her difficulty in terms of the creation of proper code. We can also pose
the question ``if Rebecca struggles to transfer logic to code, what
directs her activity as she \emph{tries} to transfer logic to code?''

Later in the interview, she compared the iTunes project --- on which she
was stuck --- to Project 2.\footnote{Project 2 did not require students
  to use malloc() to dynamically allocate memory. Project 3 did.} In so
doing, she offered another telling idea about the relationship between
her logic and ``syntax'':

Rebecca: Like, cuz, I guess I feel like I did a lot better on the last
project because it was so similar to stuff that I'm used to /Mmmhmm/ um,
whereas all this is brand new and I still don't have any of the syntax
down yet. /Mmmhmm/ Like, if-statements and while-loops, they make sense
to me, whereas I'm still like, trying to grasp at this stuff.

By ``this stuff,'' the context of the conversation suggests Rebecca was
referring to dynamic memory allocation and malloc, which we had been
discussing.

It seems sensible to think if-statements and while-loops might ``make
sense'' to Rebecca because their syntactical form aligns strongly with
their computational operation.\footnote{To borrow terminology from
  Sherin (2001) and rebut Saussure (1986) at the same time, the symbol
  template for an if-statement is strongly connected to its conceptual
  schema because we so often use everyday language in a manner
  concordant with an if-statement's branched control flow. In this
  instance, the relationship between signifier and signified is the
  \emph{opposite} of arbitrary. The similarity between the syntax for an
  if-statement and an understanding of what that statement does is by
  design, because C itself was a designed language.} By that, I mean in
principle the designers of C could have chosen \emph{any} symbol to
represent branched control flow, but they chose a word --- if --- that
invites connections to its meaning in everyday speech. If-statements in
C split control flow based on a Boolean expression's value; while-loops
allow continued iteration until a Boolean expression turns up false.
Both structures have everyday analogues with similar conceptual
properties. A parent might say, ``if you're staying out past 9, let me
know,'' or, ``while we're gone, can you make sure Sadie doesn't eat
anything from the garbage?'' If a child will be home by 8, there is no
need to let that parent know. Similarly, it's reasonable to expect that
once the parents return, the responsibility of keeping Sadie out of the
garbage no longer falls solely to the child.\footnote{Some parents might
  object that family members should \emph{always} be on the lookout for
  Sadie rooting through the trash, that these parents were just
  emphasizing the point. And, reasonable children might in general try
  to heed that. But, what makes computers computers is their lack of
  qualms about abandoning any and all responsibility to keep Sadie from
  the trash once the parents return.} The point of this observation is
that Rebecca may feel certain kinds of programming structures ``make
sense'' not simply because they were covered in introductory
programming, but because something about their syntactical form offers
an affordance for thinking about how they work.

\subsubsection{Rebecca has productive capacities for making progress on
software design
work}\label{rebecca-has-productive-capacities-for-making-progress-on-software-design-work}

In our fourth of five interviews, Rebecca was having trouble making
progress on Project 3, which the instructor described as an ``online
music server.''\footnote{Students in my study colloquially referred to
  this project as the ``iTunes project,'' so ``the iTunes project'' and
  ``Project 3'' all refer to this project.} ``Like, I am so lost, like I
don't even know where to start,'' she said. We discussed why she was
stuck. Then, with my help, we began to work through what Rebecca
identified as a core cause of her confusion. In brief, the assignment
contained a diagram showing the data scheme students were required to
use to store data about the music (Figure 8 below). Specifically, the
scheme required an array of pointers --- each entry of which pointed to
an array of characters --- as a mechanism for storing data about song
titles for the music server. Rebecca's work to resolve her confusion
offers evidence for her productive capacities in designing programs. In
this section, I argue:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Rebecca made sense of the array-of-pointers design\footnote{By
    ``design scheme,'' here, I mean the instructor-provided diagram
    indicating how students were to structure their data for Project 3.}
  using talk and gesture. She was able to explain both why an array of
  pointers \emph{should} exist in C as well as suggest candidate syntax
  for what might create such an array.
\item
  When I proposed we pretend her candidate syntax worked, Rebecca
  articulated through gesture, talk, and written pseudo-code how part of
  her design would incorporate that syntax.
\item
  After the interview, Rebecca resumed work on the project and
  incorporated a version of the design she developed in the interview
  into her code.
\end{enumerate}

\paragraph{Rebecca made sense of the array-of-pointers design using talk
and
gesture}\label{rebecca-made-sense-of-the-array-of-pointers-design-using-talk-and-gesture}

The figure below shows the instructor-provided diagram for how students
should structure their data.

\includegraphics[width=5.36667in,height=4.10000in]{media/image11.emf}

\protect\hypertarget{ux5fToc252445972}{}{}Figure ‑ -- The instructor
required students to use this data arrangement for storing information
in the music server. The scheme uses an array of pointers to represent
the track names of an album.

Rebecca would have seen this diagram before because the project
assignment was given out 11 days prior to our interview. But, the data I
present is the first time I had any access to how she was thinking about
it.

At the time of the interview Rebecca felt stuck on handling issues with
\emph{music albums} --- the top-left and top-right structures depicted
on the diagram. The assignment required that ``all of the data
structures described {[}in this diagram{]}'', including music albums and
all of the data they referenced, ``must be allocated dynamically''
(Instructor-provided Project 3 description). So, Rebecca and I started
discussing what parts of that scheme she understood. Table 3 below
presents what she said and what she did while beginning to explain her
understanding.\footnote{}

\protect\hypertarget{ux5fToc252445961}{}{}Table - Rebecca describes an
array of structures using gestures (Interview 4 of 5, April 6, 2012)

\begin{longtable}[]{@{}ll@{}}
\toprule
What she says & What she does\tabularnewline
\midrule
\endhead
Like, the---the thing that makes sense to me that I, uh, got right now,
is that, to malloc the array, to \{bounds a wide space with her hands\}
make the array a certain length, do that. &
\includegraphics[width=2.03333in,height=2.72222in]{media/image12.png}\tabularnewline
And then I know that each of the \{chops out three invisible boxes in
the air\} I guess nu---characters---I---uh, I'll call `em characters. &
\includegraphics[width=1.37778in,height=3.66667in]{media/image13.emf}\tabularnewline
Like, each of the spots in the array holds a structure and each of the
structures holds three things. & (gestures obscured by the computer
)\tabularnewline
\bottomrule
\end{longtable}

Rebecca's gestures create a kind of virtual object that looks much like
the music albums array in the diagram. As she chops out spaces for each
of the ``characters,'' those spaces are in series and fit --- more or
less --- within the larger area she bounded with her when she said ``to
malloc the array.'' The spatial subdivision is key because it offers
supporting evidence that in some way, Rebecca might have been thinking
about the album structures themselves as being \emph{contained} in or by
the array. And, that spatial containment metaphor resurfaced later when
we discussed the array of pointers, which I'll get to in a moment.

As Rebecca explained her understanding to me, I took to writing down
what I understood her to be saying on paper:

Interviewer: OK. So you have, so here's what you just said. You said
\{writes\} ``malloc the array'' um, ``each element has a structure,''
and then ``each structure has'' you said---

Rebecca: Has three parts

Interviewer: OK. Has three parts. OK.

Rebecca: Yeah. So, and the first part I can do `cuz it's an integer
\{laughs\}. The first part is just the number of tracks on the album.
So---

Interviewer: Oh, oh. So, of the three parts {[}{[}I'm sorry

Rebecca: Yes{]}{]}

Interviewer: I {[}{[}\{untillegible\}

Rebecca: Oh yeah, sorry{]}{]}

Interviewer: So you're saying, um, and then---and then this is number of
tracks, which is an integer, right? /yeah/ OK (Interview 4 of 5, April
6, 2012)

Figure 9 below shows what I wrote as we worked:

\includegraphics[width=5.62222in,height=1.48889in]{media/image14.emf}

\protect\hypertarget{ux5fToc252445973}{}{}Figure ‑ -- I write down what
I understand as Rebecca explains the overall data structure of albums to
me

It wasn't the first time --- either in my series of interviews with
Rebecca or in that interview in particular --- that I tried to write
down what she said. But, I highlight it here because it became part of
my activity as we worked together in this episode. Rebecca would
articulate something, often through talk and gesture, and I would try to
write down and offer back to her what I understood her to be explaining.

As we continued, we landed on trying to understand the ``tracks''
portion of the album data structure. Storing tracks was where Rebecca
felt she was getting ``lost'':

\protect\hypertarget{ux5fToc252445962}{}{}Table -- Rebecca says gets
lost on the array of pointers pointing to track names (Interview 4 of 5,
April 6, 2012)

\begin{longtable}[]{@{}ll@{}}
\toprule
What she says & What she does\tabularnewline
\midrule
\endhead
Rebecca: And, the next one is where I get lost, on the tracks part.
Because it's a pointer \{2 sec pause\} to \{hands make a vertical
cylinder in front of her; note that the array is represented on the
assignment as a narrow, vertical rectangle\}--- &
\includegraphics[width=3.40000in,height=2.62222in]{media/image15.emf}\tabularnewline
it looks like another array that's
\textbar{}\textbar{}pointing\textbar{}\textbar{} \textbar{}\{left hand
crosses from her left to right, at chest height, index finger extended
in the direction of motion; this is the same direction (from Rebecca's
perspective) in which the pointer array on paper points to each track
name\}\textbar{} to the names. &
\includegraphics[width=1.98889in,height=3.38889in]{media/image16.emf}\tabularnewline
\bottomrule
\end{longtable}

Rebecca's gestures for the array of pointers again spatially mimic the
depiction of the array on the page: a vertical cylinder. The reason I
note the orientation of her array gestures is that for practical
purposes arrays as represented by bits in C have no orientation in space
that the programmer could know, much less care about. An array cannot be
parallel to the ground, nor can it be oriented toward the
ceiling.\footnote{The computational hardware representing the array does
  have an orientation in space, but that's clearly not what is at issue
  here. The whole point of the array is to be an abstraction away from
  the soldered transistors, capacitive plates, flipped switches, vacuum
  tubes, tinker toys, or other physical means that store the state of
  the array.} But, the assignment \emph{depicted} the album array as
horizontal, and so too did Rebecca's gestures for it. The assignment
\emph{depicted} the pointer array as vertical, and so too did Rebecca's
gestures for it. In other words, the observation that Rebecca's gestures
align with the assignments pictures might suggest that she is doing more
than thinking about arrays in the abstract. Rather, some elements of the
representational features of the page have made their way into her
activity and, arguably, her cognition.\footnote{One could argue that
  gesture is possibly just undirected flailing, and that the correlation
  between Rebecca's gestures and the page is weak at best --- the sort
  of thing you'd expect to happen some percentage of the time anyway
  under the assumption that the pictures on the page have nothing to do
  with how she's thinking about it. I concede that it's \emph{possible}
  the correlation isn't structurally meaningful. But, I think the
  correlation is meaningful given how many other times her gestures
  align with canonical written representations of computational
  structures and processes.}

Rebecca continued on to try thinking through why an array of pointers
was part of the design.

Rebecca: But, I've always been confused as to why you---I guess, I
always---I was always like, when we learned pointers, I was like ``why
do you need pointers when you could just call it the name? Why do you
need two names for it?'' But, I think, what---at least what I'm seeing
here maybe is like, this is just another array /Mmmhmm/ because you
can't put more than one character in each element of the array, so that
is just an array of pointers that point to strings. And the strings are
the names.

Rebecca: So, that would make sense. So you would---the tracks would
point to the array, and so you could do the pointer of that \{1.5 sec
pause\} The pointer to the array, an array---the array of like 1 would
be a pointer to track 1's name.

Interviewer: Mmmhmm

Rebecca: So. OK, so---see that part makes sense now. I just---no idea
how I would ever access that. Or, store it, I guess, is the better word.
(Interview 4 of 5, April 6, 2012)

What ``confused'' Rebecca has an understandable origin. Most of the work
she would have been doing until that point would have used named
variables. For example:

int age;

float weight;

char *name;

If age, weight, and name are already-named pieces of data, why would
something ever need to refer to them indirectly? It would be as if we
had to call me ``The second-born son of the second-born daughter of
Henry'' for some reason. I already have a name. Why do I need another
name that uses the names and relationships of my forebears to refer to
me when historically my given name has perfectly fine for all
referential purposes?

Rebecca's way of finding sense in pointers was to observe that you can't
put names --- each of which is defined as an array of characters ---
into an array because ``you can't put more than one character in an
array.'' To verify I was understanding her idea correctly, I tried to
restate it:

Interviewer: OK, so you were saying the reason you---when you're tryin
to convince yourself like---you originally you were thinking ``why don't
you just put all the names {[}{[}here

Rebecca: Yeah{]}{]}

Interviewer: And you were {[}{[}saying the reason is

Rebecca: saying, no{]}{]}

Interviewer: because==

Rebecca: ==That makes sense now. Yeah, you can't put more than one
character in an element. And, it's a whole name of a song, so you would
need to point---that element would just \{gesture obscured by my
laptop\} be a pointer to a name. OK. (Interview 4 of 5, April 6, 2012)

The latching and overlapping turns of talk above offer strong evidence
that Rebecca's explanation aligned with my interpretation. And again,
notions of \emph{containment} come into play. Rebecca says, ``you can't
\emph{put} more than one character \emph{in} an element'' in part
because you're talking about ``a \emph{whole name} of a song'' (my
emphasis added here). If Rebecca has a capacity for thinking about the
canonical concept of type-restrictions on arrays, it's manifesting here
as a part-physical metaphor, evoked through both talk (the quote above)
and gesture (Table 4), in the service of making sense of a design
decision.

As she continued, Rebecca discerned that because the input data
contained the number of tracks on each album, she could use that number
to define how long the pointer array needed to be. But, something was
still troubling her.

Rebecca: So. \{4 sec pause\} That would \{6 sec pause\}

Interviewer: What're you thinkin?

Rebecca: Just that, um, I'm just tryna think, cuz like that makes sense,
but I don't know how to get to that. Cuz, how do you make a pointer
point to a random array that you just made? Uh, I guess \{4 sec pause\}

Rebecca: Uh, p---the pointer syntax I need---I would need to go back and
look at them because now you have an array, each array---my issue---cuz
like now that makes sense. Like, it's just how do I get to that point of
making that work?

Interviewer: {[}{[}So

Rebecca: And this is what{]}{]} I mean by my logic /OK/ I get the logic
and the theory behind it but I don't know how to actually put it into C.

Interviewer: OK, so in other words, what makes
\textbar{}\textbar{}sense\textbar{}\textbar{} \textbar{}\{air
quotes\}\textbar{} to you know /Mmmhmm/ that I guess didn't before is
that *that* \{gestures to the instructor's structural diagram on
screen\} is a---that's an understandable arrangement /yes/ for the data

Rebecca: Yes.

Interviewer: But then, how do I {[}{[}make it

Rebecca: make it{]}{]} Yeah \{nods\}. (Interview 4 of 5, April 6, 2012)

When Rebecca said ``I don't know how to get to that,'' I take ``how to
get to that'' to mean \emph{how to write the code for that procedure}.
We had just agreed that design-wise, she could use a number provided in
the input data to define the length of the pointer of array. But, she
was stuck because of her difficulty translating the ``logic'' we
developed into code.

From the outside, what Rebecca and I had done together was something I
think crucially important to the design of a program. We worked to
understand how someone else chose to structure their data.\footnote{That
  the ``someone else'' was in this case the instructor is important for
  larger power dynamic considerations.} We tried to decide for ourselves
whether and how those structures made sense. Along the way we created
ephemeral objects of that sense-making in the form of talk and gesture
as well as the durable artifacts of my inscriptions. But, none of that
work resulted in C code.

For Rebecca, the lack of certainty about \emph{how to make} our ideas in
C code was a marked concern. And, again, that concern makes sense. We
had just convinced ourselves that we could understand \emph{why} an
array-of-pointers. We even figured out \emph{how} we could use the input
data to provide the length of the arrays we would need to make. But we
had not laid out how to declare the array \emph{and} preserve a
reference to it (``how do you make a pointer point to a random array
that you just made?). Without that piece --- a piece Rebecca was unsure
of and felt she would need to look up --- our solution clearly wouldn't
work.

Absent that piece, Rebecca tried thinking of possible ways to fill in
the missing syntax.

Rebecca: Cuz, I'm like, like I'm trying to think right now of things
that I would try /Mmmhmm/ but I don't know what I would try first
because, all I'm thinking right now is that I know tracks is going to be
a pointer. /Mmmhmm/ So you make that a pointer.

Interviewer: OK.

Rebecca: But I'm not sure where it would point to, because you don't
have that array made yet. So I \{3 sec pause\} so you'd have to, I guess
you could make the array, and then make it point to the array, but \{3
sec pause\} Cuz you have to make an array of pointers.

Interviewer: Mmmhmm

Rebecca: And I don't know how to do that. Because I've made arrays of
characters and integers and stuff before /OK/ but never pointers, so.
(Interview 4 of 5, April 6, 2012)

Rebecca was seeing a temporal problem: how could she reference an array
that didn't exist yet?\footnote{As a crude analogy, imagine trying to
  write your own will. If you think forward in time, you might later
  have children to whom you'd like to bequeath your assets. But how can
  you enumerate those children in the will now, before you've had them?
  You need your will to refer to something that does not exist at the
  time you write the will. Extrapolating forward and rather ludicrously,
  we could imagine you are always capable of producing children. So, up
  until the day of your death, there is always the seeming risk that
  your will might need to refer to a beneficiary who does not yet exist.
  Recognizing this subtle problem as Rebecca did is an act I consider to
  be sense-making, even though she as-yet still did not have a solution.}

Rebecca started exploring ways to solve that problem of getting tracks
to point to the array of pointers.

Rebecca: But I'm not sure where it (tracks) would point to, because you
don't have that array made yet. So I \{3 sec pause\} so you'd have to, I
guess you could make the array, and then make it point to the array, but
\{3 sec pause\} Cuz you have to make an array of pointers.

Interviewer: Mmmhmm

Rebecca: And I don't know how to do that. Because I've made arrays of
characters and integers and stuff before /OK/ but never pointers, so.
(Interview 4 of 5, April 6, 2012)

Rebecca landed on a potential solution: make the array first, then make
tracks point to the array. But, with that solution came another problem:
how do you make an array of pointers?

\paragraph{Rebecca could propose candidate syntax and build pseudocode
around
it.}\label{rebecca-could-propose-candidate-syntax-and-build-pseudocode-around-it.}

Rebecca came up with an idea to create an array of pointers, but she
wasn't sure if it would work.

Rebecca: if I were to d---when I declared the array, if I were to---I
dunno if you even can, like, whenever we---we do ints, and then it would
be like star-p /Mmmhmm/ and then star-p would be the pointer /ok/ I
don't know if you can do like int star-p
\textbar{}\textbar{}bracket-bracket\textbar{}\textbar{}
\textbar{}\{makes square brackets with hands\}\textbar{} and that would
be an array or not.

Interviewer: {[}{[}OK

Rebecca: But I don't know{]}{]} if you can do that.

Notice what's happening here. Rebecca is taking patterns that she knows
work:

int age; // creates an integer variable called age

int ages{[}35{]}; // creates an array called ages,

// 35 integers long

int *p; // creates a pointer-to-an-integer;

// said pointer is called p

and considering a pattern that might work:

int *p{[}35{]}; // Rebecca is not exactly sure what this will do

Notice further that in this exchange, Rebecca wondered ``if you can do
that.'' Determining the referent of \emph{that} was, again,
consequential. One possibility was Rebecca meant ``int *p{[}35{]}''
might not be the proper syntax for creating an array of 35
pointers-to-ints. If so, Rebecca was struggling with her aforementioned
problem of not being able to translate her logic to code. Another
possibility was Rebecca meant ``I don't know if there exists \emph{any}
syntax in C to create an array of pointers.'' So, I followed up on this
latter interpretation. We both agreed it \emph{should} be possible to
create an array of pointers, it was indeed just a question of whether
that syntax accomplished it (Interview 4 of 5, April 6, 2012). So, that
left the former interpretation: it looked like Rebecca was unsure how to
``transfer her logic to code.''

We both agreed creating an array of pointers should be possible. So, I
asked Rebecca to suppose her candidate syntax worked. She readily
agreed.

Interviewer: OK. So then, um, what if we just pretended for a minute,
that that, like==

Rebecca: ==That that works==

Interviewer: ==That that worked==

Rebecca: ==OK==

Interviewer: ==OK. So then, um. \{begins writing\} So then you might
write like to make an array of \{stops writing\} actually what would we
call this? This is==

Rebecca: ==array of pointers, I guess==

Interviewer: ==OK. \{under breath\}
\textbar{}\textbar{}pointers\textbar{}\textbar{} \textbar{}\{writes
``pointers''\}\textbar{}. It'd be, well. So. The---the one we had was
like \textbar{}\textbar{}int, star p, um, it would be some
number\textbar{}\textbar{} \textbar{}\{writing\}\textbar{}, um, and I
guess that'd be it in order to just declare /Yes/ that right? OK.

In this exchange note every turn boundary is latched. Once I bid for
pretending, Rebecca accedes. When I'm unsure what to call the array,
Rebecca finishes my sentence. Rebecca rapidly approves the final product
before I even finish my last sentence. Figure 10 below is the syntax I
wrote in the exchange.

\includegraphics[width=5.12500in,height=1.09792in]{media/image17.emf}

If that syntax created the array-of-pointers, the next challenge for
storing track names was how to access elements in array-of-pointers.
Figure 11 presents the pseudo-code Rebecca ultimately wrote to scan in
and store track names.

\includegraphics[width=5.36667in,height=1.30000in]{media/image18.emf}

\protect\hypertarget{ux5fToc252445974}{}{}Figure ‑ -- Rebecca's
pseudo-code for scanning in and storing track names

The next section will show the rich talk and gestures that accompanied
Rebecca's generation of this pseudo-code. I argue activity in the form
of gesture and talk constituted and sustained Rebecca's in-the-moment
approach to programming.

\paragraph{Rebecca's talk and gestures constituted and sustained her
in-the-moment approach to
programming}\label{rebeccas-talk-and-gestures-constituted-and-sustained-her-in-the-moment-approach-to-programming}

Throughout this section I detail the speech and gestures that accompany
Rebecca's pseudocode production. Along both modalities I saw evidence of
a marked shift in Rebecca's orientation. It was a shift from being
hunched, silent, thinking over the keyboard to being animated, open, and
gestural toward me. It was also a shift from seeming diffident,
uncertain, and hedging in speech to seeming confident, assured, and able
to dispatch my questions. Taken in their totality across the episode,
these subtleties constitute evidence of an approach that strongly
differs from those Rebecca had when discussing her struggles in the
course (section 3.5.1).

First, Rebecca suggested she could write the pseudo-code as we talked.

\protect\hypertarget{ux5fToc252445963}{}{}Table -- Rebecca starts
talking out pseudo-code and asks for the pen to write it

\begin{longtable}[]{@{}ll@{}}
\toprule
What she said & What she did\tabularnewline
\midrule
\endhead
Uh, well, what I was think---like if I wanted to like,
\textbar{}\textbar{}save it\textbar{}\textbar{} \textbar{}\{pinches
right hand, thumb and index finger a few inches apart\}\textbar{} or
whatever, /Yeah/ &
\includegraphics[width=2.68889in,height=3.23333in]{media/image19.png}\tabularnewline
I could make a while loop, \textbar{}\textbar{}scan in
the\textbar{}\textbar{} \textbar{}\{drags hand across the
table\}\textbar{} ---scan in the data from the album, /OK/ \{interviewer
begins writing\} uh, which is, {[}{[}I can write it if you want &
\includegraphics[width=2.66667in,height=3.23333in]{media/image20.png}\tabularnewline
\bottomrule
\end{longtable}

Rebecca then took the pen and started describing her thinking while she
wrote.

\protect\hypertarget{ux5fToc252445964}{}{}Table -- Rebecca's
verbal/gestural overview of looping through the input track titles
(Interview 4 of 5, April 6, 2012)

\begin{longtable}[]{@{}lll@{}}
\toprule
Panel \# & What she said & What she did\tabularnewline
\midrule
\endhead
1 & And then my thinking at least, is you should be able to, um, say
that

``star p of i'' /mmhmm/ equals, uh, the title, and then you just do i++,
so then it'll \textbar{}\textbar{}move to the next
one\textbar{}\textbar{} \textbar{}\{makes looping gesture with left
hand\}\textbar{} /OK/ &
\includegraphics[width=2.03333in,height=2.33333in]{media/image21.png}\tabularnewline
2 & and you just keep \textbar{}\textbar{}saving each of the
pointers\textbar{}\textbar{} \textbar{}\{left hand makes horizontal
chops in the air, like rungs down a ladder\}\textbar{} &
\includegraphics[width=3.40000in,height=1.92222in]{media/image22.emf}\tabularnewline
3 & in the array \textbar{}\textbar{}to a title\textbar{}\textbar{}
\textbar{}\{right hand makes pinching motion, left-hand points to
it\}\textbar{} &
\includegraphics[width=2.18889in,height=2.24444in]{media/image23.png}\tabularnewline
4 & \textbar{}\textbar{}And\textbar{}\textbar{} \textbar{}\{right hand
makes a large loop counterclockwise (from Rebecca's
perspective)\}\textbar{} &
\includegraphics[width=3.13333in,height=1.54444in]{media/image24.emf}\tabularnewline
5 & \textbar{}\textbar{}you just increment by 1\textbar{}\textbar{}
\textbar{}\{right hand makes a cycloid to her right, in a plane parallel
to the back wall, like counting dots on a line\}\textbar{} until you
reach the end of file. &
\includegraphics[width=1.06667in,height=3.62222in]{media/image25.emf}\tabularnewline
\bottomrule
\end{longtable}

``Like, that would make sense to me,'' Rebecca said. I asked whether the
while loop would get a fresh line of text when it runs again.

\begin{longtable}[]{@{}lll@{}}
\toprule
Panel \# & What she said & What she did\tabularnewline
\midrule
\endhead
1 & Yeah, because{]}{]} uh, \textbar{}\textbar{}what the while loop
does\textbar{}\textbar{} \textbar{}\{pinches right thumb and
forefinger\}\textbar{} &
\includegraphics[width=1.74444in,height=1.95556in]{media/image26.png}\tabularnewline
2 & \textbar{}\textbar{}is it reads in the line\textbar{}\textbar{}
\textbar{}\{right hand, palm down, chest height, slides out to the
right\}\textbar{} &
\includegraphics[width=1.76667in,height=2.83333in]{media/image27.emf}\tabularnewline
3 & and then \textbar{}\textbar{}once it reaches\textbar{}\textbar{}
\textbar{}\{right hand rises, lowers in a chop\}\textbar{} &
\includegraphics[width=1.78889in,height=2.11111in]{media/image28.png}\tabularnewline
4 & the end of \textbar{}\textbar{}character it'll\textbar{}\textbar{}
\textbar{}\{right hand rises, swoops down in a crescent, index finger
extended, looping back up\}\textbar{} &
\includegraphics[width=2.74444in,height=1.36667in]{media/image29.emf}\tabularnewline
5 & go back down---it'll \textbar{}\textbar{}do the while
loop\textbar{}\textbar{} \textbar{}\{right hand moves back down toward
table\}\textbar{} \textbar{}\textbar{}and then\textbar{}\textbar{}
\textbar{}\{right hand moves back up above shoulder-height,
palm-down\}\textbar{} &
\includegraphics[width=2.63333in,height=1.55556in]{media/image30.emf}\tabularnewline
6 & \textbar{}\textbar{}go back\textbar{}\textbar{} \textbar{}\{right
hand creates a loop down and back up\}\textbar{} down to the next
thing---line, &
\includegraphics[width=3.38889in,height=1.31111in]{media/image31.emf}\tabularnewline
7 & it'll \textbar{}\textbar{}read in the line\textbar{}\textbar{}
\textbar{}\{right hand scans rightward at shoulder height\}\textbar{} &
\includegraphics[width=1.61111in,height=2.47778in]{media/image32.emf}\tabularnewline
8 & \textbar{}\textbar{}until it\textbar{}\textbar{} \textbar{}\{pinches
right thumb and forefinger, swipes right hand back to the left, palm
down\}\textbar{} reaches the end of file &
\includegraphics[width=3.38889in,height=1.45556in]{media/image33.emf}\tabularnewline
\bottomrule
\end{longtable}

In panel 4, notice that Rebecca's gesture of swooping back actually
\emph{precedes} her saying ``go back down.'' In other words, her body
articulates the idea of cycling back before she actually speaks it. A
similar argument could be made for panel 8, where Rebecca's gesture for
a process hitting the bottom (hand swooping and hitting an inflection
point) precedes her saying ``reaches the end of file''

I tried to sum up my understanding of Rebecca's description:

Interviewer: So as you step through this loop /mmhmm/ \{points to i++\}
i keeps going up by one==

Rebecca: ==Yes. (Interview 4 of 5, April 6, 2012)

As I thought about that, Rebecca continued explaining and gesturing:

\protect\hypertarget{ux5fToc252445965}{}{}Table -- Rebecca concludes her
visual and gestural explanation for scanning in track titles (Interview
4 of 5, April 6, 2012)

\begin{longtable}[]{@{}lll@{}}
\toprule
Panel \# & What she said & What she did\tabularnewline
\midrule
\endhead
1 & And the \textbar{}\textbar{}lines keep going
down\textbar{}\textbar{} \textbar{}\{left hand horizontally chops the
air, creating ``ladder rungs''\}\textbar{} & (Gesture partially obscured
by computer )\tabularnewline
2 & so, \textbar{}\textbar{}that way,\textbar{}\textbar{}
\textbar{}\{left hand raises up to her head, palm down and parallel to
the table, and it is replaced by right hand, index finger extended and
pointing to her left\}\textbar{} &
\includegraphics[width=2.10000in,height=2.60000in]{media/image34.emf}\tabularnewline
3 & \textbar{}\textbar{}the first line\textbar{}\textbar{}
\textbar{}\{right hand scans across to her right\}\textbar{} &
\includegraphics[width=1.57778in,height=3.50000in]{media/image35.emf}\tabularnewline
4 & is going to \textbar{}\textbar{}be\textbar{}\textbar{}
\textbar{}\{right hand rotates to become pinched thumb and forefinger,
palm out\}\textbar{} , \textbar{}\textbar{}uh the
p---\textbar{}\textbar{} \textbar{}\{right-hand wiggles\}\textbar{} &
\includegraphics[width=3.25556in,height=1.62222in]{media/image36.emf}\tabularnewline
5 & element zero /OK/ uh \textbar{}\textbar{}the
second\textbar{}\textbar{} \textbar{}\{hand swoops slightly up, then
curves down and locks in in a position below the prior one\}\textbar{}
one'll be element one &
\includegraphics[width=3.25556in,height=1.55556in]{media/image37.emf}\tabularnewline
\bottomrule
\end{longtable}

\subsection{Conclusion}\label{conclusion-1}

In pulling together concluding ideas, I revisit the points I established
at the end of section 3.1.

\subsubsection{Students' early-stage design activity reveals patterns
outside the explanatory scope of (mis)conceptions
accounts}\label{students-early-stage-design-activity-reveals-patterns-outside-the-explanatory-scope-of-misconceptions-accounts}

Lionel's strategy of working at a whiteboard, keeping himself at a
top-level of planning, and copying pseudo-code into a computer isn't
explained by appealing to ``concepts'' in computer science, which in
many research accounts are simply mappings of content in computer
science curricula. I demonstrate this gap in Appendix 4, where the
``conceptual features'' of Lionel's code reveal little (if anything)
about the resources involved in Lionel's design process.

Rebecca, in a similar fashion, has lots of productive knowledge for
thinking about arrays (section 3.5.2), but at the time of the interview
none of her code reflected that knowledge. Moreover, Rebecca was able to
sense-make (Danielak et al., in press) about an array-of-pointers
design, where her sense-making was again a complex activity that would
be poorly accounted for in CSEd frameworks that appeal to
misconceptions. In Rebecca's case, the situation is particularly
paradoxical if we consider conceptual knowledge about a topic to naively
be something students have or don't have.

Suppose Rebecca had the requisite conceptual knowledge for thinking
about an array of pointers. It's a puzzle, then, to explain solely via
conceptual knowledge why she was so frustrated with programming and felt
stuck when she began Interview 4. Why hadn't the knowledge she
possessed---relevant to solving the problem at hand---manifested already
some way? Why was she stuck?

The alternative is to assume that as of Interview 4 she didn't have the
conceptual knowledge for thinking about the problem at hand. But, that
doesn't make sense either. If Rebecca didn't know how to think about an
array of pointers, why was she able to do it so well during the
interview? Both possibilities---that either Rebecca had or did not have
conceptual knowledge about an array of pointers---lead to fairly
non-sensical conclusions under naïve have/don't have assumptions of
conceptual knowledge. Yet, again, the majority of conceptual knowledge
research in CSEd today is silent on the issue of degree when it comes to
conceptual knowledge: students either have a mental model that matches
canonical function or they don't.

\subsubsection{Rebecca had epistemological resources to support
expert-like practices, but she framed those practices
differently}\label{rebecca-had-epistemological-resources-to-support-expert-like-practices-but-she-framed-those-practices-differently}

Consider these summative statements made by Lionel and Rebecca on how
they use pseudo-code in their work.

\protect\hypertarget{ux5fToc252445966}{}{}Table -- Comparing Lionel's
and Rebecca's views toward pseudo-code

\begin{longtable}[]{@{}ll@{}}
\toprule
Lionel & Rebecca\tabularnewline
\midrule
\endhead
``on my computer I'd, you know I'd write out the pseudocode, I'd
ac---I'd literally write out the pseudocode, even though obviously it
wouldn't compile and actually work.'' (Interview 1 of 1, October 17,
2011) & ``So, I can't just type in ``if the white piece reaches in,''
I---the, the C language, I guess, putting it in those, their terms,
their terminology into programming language.'' (Interview 2 of 5,
February 17, 2012)\tabularnewline
\bottomrule
\end{longtable}

As illustrative examples, they capture a fundamental difference in
practice between Lionel and Rebecca. Lionel worked in pseudo-code on his
computer, even going so far as to copy it into his source-code files
despite the fact that it ``obviously wouldn't compile and actually
work.'' Rebecca also wrote pseudo-code, and she even wrote it into her
source code files. But, her local sense of the activity of writing
pseudo-code was different. It came to represent a discontinuity with the
final code she was trying to write because there was nonetheless a gap
between writing in English and writing C in ``their terms.''

These diverging stances take the same practice---in this case,
pseudo-coding---and situate it within a different kind of
epistemological coherence. For Lionel, the coherence associated with
pseudo-coding is productive and optimistic. Copying non-working
pseudo-code into the computer is, in a colloquial sense, all part of the
plan. It's a legitimate step toward creating a final working program
and, crucially, it reflects his own understanding of what's supposed to
happen in his code. For Rebecca, by contrast, writing pseudo-code is a
fallback. Pseudo-code is what she writes ``if I know what I want to put
underneath of it,'' by which she means if she's not sure how to flesh
out a loop or other control structure. Rather than being part of a
planful coherence, the practice of pseudo-coding for Rebecca gets
triggered as part of a stopgap coherence; a contingent measure at times
associated with a kind of diffidence.

Crucially, Rebecca has intellectual resources for pseudo-coding.
Moreover, her practice of writing pseudo-code to flesh out a loop is the
plausible beginning of ``method stubbing,'' the process by which one
might defer implementing the guts of a function and instead create a
simple stopgap. For example, the function below will always print 28
Fahrenheit, which is fairly useless as far as temperature-getting
functions go:

getOutsideTemperature \textless{}- function(sensor) \{

print(``28F'')

\}

But, the advantage of having \emph{something} in the function body, no
matter how trivial, is enormous. Now, because
\textbf{getOutsideTemperature} is defined and takes arguments, other
functions can safely call it. It also offers a diagnostic output---by
printing ``\textbf{28F}'' out---that we can rely on if we ever need to
make sure the function was called. At the time of creating
\textbf{getOutsideTemperature}, a programmer may have no idea how the
temperature will be obtained but will nonetheless need to be able to
write code that depends on it getting the temperature. Stubbing is, in
such cases, a highly productive thing for a programmer to be able to do.
But, stubbing is as much a strict practice as it is an epistemological
move, because it represents a decision that an entity-to-be-known can
be, for the moment, underspecified so it can be incorporated into a
system before its behavior is defined.

Rebecca is already capable of simple stubbing. It's part of an
epistemological coherence that's triggered when she doesn't know how to
flesh out a control structure. So, from a constructivist standpoint
(Smith et al., 1993), she has knowledge of something that can be further
refined into software engineering knowledge. Indeed, much of the
surrounding data from Rebecca suggests that she would have been served
well if an instructional intervention had taught her some simple ways to
stub out functions so she could call them even if she hadn't defined
their bodies yet. To my knowledge, she was never exposed to stubbing
formally, which is unfortunate considering how much it might have led to
improving her confidence in programming.

Lastly, consider Rebecca's episode in reasoning about an array of
pointers (section 3.5.2). As I show, Rebecca clearly had resources for
reasoning about the sensibility of an array-of-pointers design and for
predicting array-of-pointers syntax by extrapolating from patterns she
had seen before. And, when I explicitly suggested that we suppose a
candidate syntax works, Rebecca was able to fluidly articulate the
procedure she would build around that syntax (section 3.5.2.3). Rebecca
knew things, but her knowledge was deployed in such a highly contextual
way as to be sensitive to what kind of knowledge-activity she thought
she was supposed to be doing. It's thus sensible to model that set of
phenomena from an epistemological standpoint, rather than a strictly
conceptual one. Rebecca had certain kinds of knowledge, but part of my
in-interview intervention was to establish a frame where \emph{supposing
syntax worked} was a valid part of programming. Once we mutually
negotiated that frame there was a strong microcoherence of Rebecca
stably explaining her code. But, crucially, my intervention wasn't about
introducing conceptual content but rather establishing a kind of
knowledge game---supposing---that one could play as part of programming
design.

\subsubsection{Students displayed a diversity of approaches to
programming in the
moment}\label{students-displayed-a-diversity-of-approaches-to-programming-in-the-moment}

The diversity of programming approaches I saw students take is far
greater than just the space spanned by Lionel and Rebecca. Within and
across my data, I continually observed students saying and doing things
that not only distinguished them from each other, but constituted
phenomena I haven't seen described in prior CSEd research. I saw
students employ clever tests for debugging, create their own makeshift
debugger, and discuss at length whether certain programming constructs
represented natural ways of thinking. For illustrative purposes, I'll
discuss three such examples. While I don't analyze them at length, I
include them here as proof of principle of the kinds of phenomena we're
not currently capturing that nonetheless have a strong effect on
students' programming approaches. Moreover, these ``in the wild''
(Hutchins, 1995a) practices are further evidence of the kinds of
knowledge that could be further refined into student expertise (Smith et
al., 1993). So, not only do our research accounts miss these kinds
phenomena, by missing them they preclude research and practice from
building off them.

\paragraph{Isaac used a thoughtful debugging strategy that code
snapshots alone could never
capture}\label{isaac-used-a-thoughtful-debugging-strategy-that-code-snapshots-alone-could-never-capture}

When Isaac was debugging an error in his checkers game code, he made use
of a clever test to discover he had reversed array index subscripts. He
had become uncertain about which of the two subscripts corresponded to
what he thought of as the x-coordinate on the board and which controlled
the y-coordinate. So, he took a piece whose x-y position he knew and
instructed his program to print it as a ``7'' on the board. Then, he
told it to increment the first index of that known piece by 1, leave the
second index alone and print the resulting piece on the board as ``8''.
Finally, he told his program to increment the second index of the known
piece by 1, leave the first piece alone, and print the resulting piece
on the board as ``9''. By inspecting the visual output, Isaac could tell
which subscript controlled which position because he knew where 8 and 9
were relative to the unchanged 7. When he reflected on his work in the
interview, he explained that what was tricky was realizing that the
coordinate system for the checkers board didn't work the way standard
coordinate systems in math worked; in the checkers program the indices
were reversed.

Because Isaac had been set up as a code-snapshot participant, we have a
snapshot record of his work for that portion of the interview. But, the
only snapshot-visible changes are the reversal of the indices and the
accompanying test cases. The narrative surrounding Isaac's particular
debugging text, including his reflection that coordinate systems in
programming are ``trickier'' than those in math, is entirely invisible
to the snapshot record. If I hadn't interviewed him I never would have
captured it.

\paragraph{Dana created her own debugging
environment}\label{dana-created-her-own-debugging-environment}

Dana was working on her checkers program and needed to debug some errant
behavior. In class and in discussion students had been exposed to the
GNU Debugger (GDB), so in the interview I asked Dana if she'd like to
use it. She said she didn't feel comfortable using GDB. Instead, using
my MacBook, Dana decided to create three separate instances of a
terminal. In one instance, she had her checkers program source code
open. In the second, she had the source code for a set of test inputs
she was developing to try to pinpoint the error. The third window was
dedicated to compiling and running the code. Since the program was
designed to print an ASCII (text) representation of the board at each
turn, the dedicated program window was Dana's visual output for which
pieces were where at any state in the game.

The important finding from this episode is how Dana deliberately
structured her environment to support her activity. At the lowest level,
she could have tried to do everything in a single terminal window. But,
if she did that she'd have to constantly switch contexts. Because
students in the course used a terminal version of emacs, she had no easy
way to simultaneously view her code while compiling it. GDB would have
been in many respects the right tool for that problem: it would let her
set breakpoints, step through iterations of code, and inspect variable
values as she did so. But Dana chose not to use GDB.

Instead, Dana exploited the fact that Terminal.app on Mac OS X 10.7 can
have multiple instances of itself open at the same time on the same
screen. With three terminals running, she could persistently have her
source code visible and executing at the same time. With one glance
pattern she could move from the board's visual output to the test to the
program source and back again. Collectively, the field of displays
enabled a kind of programming that was substantially different---in
terms of its representational affordances---from using GDB or a single
terminal instance. Moreover, a snapshot-history-only research approach
would have never had access to how or why Dana (re)configured her
development environment.

\paragraph{Toby said recursion was the ``hardest programming way to
think''}\label{toby-said-recursion-was-the-hardest-programming-way-to-think}

When I gave Toby a code sample task in an interview, he immediately
commented on the fact that it used recursion. Recursion was, as he said,
``the hardest programming way to think.'' At one point his exact
phrasing, quoting the film \emph{Ice Age}, was that ``recursion is bad
juju.'' The task in question was adapted from an example in Abelson and
Sussman (1996) that approximates the square root of a number using an
elegant recursive approach.

Toby's first explanation for why recursion was bad was that it's a
completely unnatural way to think. He cited examples from cooking by
saying, in effect, ``Nobody ever stirs by saying stir once, and if it's
not stirred, stir once. They say stir 100 times.'' He cited simple
operations---including incrementing a number---where using recursion to
push and pop a stack seemed needlessly complex. One of his most damning
indictments of recursion was that his class introduced it as a solution
to generate numbers in the Fibbonacci sequence. Of course recursion
works for examples that are mathematically recursive, he reasoned, but
outside of those obviously contrived examples it was bad joo joo when
compared to understandable, sensible iteration.

The fuller account of Toby's resistance to recursion---and slight change
of heart after an instructional moment---is beyond the scope of what I
can present here. Suffice it to say that Toby's feelings toward
recursion---rooted in a sense of it being an unnatural way to
think---strongly directed his approach to programming. The approach was
so strong, in fact, that during our first interview he wasn't fully able
to produce working code examples of why recursion was absurd despite his
vehement conviction that it was. In other words, he maintained that
recursion was bad and inefficient even though every time he tried to
demonstrate its inferiority he made mistakes in his code and his
examples didn't work.

\subsubsection{Dynamic epistemological models can offer a lens for
reforming assessment and
instruction.}\label{dynamic-epistemological-models-can-offer-a-lens-for-reforming-assessment-and-instruction.}

Historically, a recognition of in-pieces cognitive dynamics in science
education (diSessa, 1993; Hammer, 1994) has led to more direct research
on how those frameworks inform instruction (Hammer \& Elby, 2003;
Hammer, 1996; Louca et al., 2004). What I describe here is research
still at the formative end of generating models to explain cognition.
And, because the focus of the research was explicitly on learning and
not instruction, I advise prudence in trying to draw certain kinds of
specific recommendations from the work I present.

Those restrictions being said, we can think carefully about what this
work means for instruction. For practical purposes, let's refine
terminology so we can talk more specifically and precisely about
components of teaching. Below are my working definitions for teaching
components:

\begin{itemize}
\item
  How should instructors change the way they act in the classroom
  (\emph{instruction})?
\item
  How should this research inform the set of intended learning outcomes
  instructors prepare (\emph{curriculum})?
\item
  How should we change the way we measure or otherwise go about
  ascertaining what students know (\emph{assessment})?
\end{itemize}

The clearest implications of my work are in instruction and assessment.

From an instruction perspective, it may help teachers to know that
students can enact different epistemological stances as part of their
programming practice. As stated, that finding is in principle not new
(cf., Gaspar \& Langevin, 2007). What is new, I think, is the
characterization that these stances evidence epistemological resources
that can be tapped for learning (cf., Hammer \& Elby, 2003). That
sense-making about program design choices or computational concepts is a
worthwhile activity, for example, is a message instructors can send to
students that resonates with larger findings in engineering education
research (Danielak et al., in press). Moreover, instructors can be more
sensitive to what kinds of knowledge-activity students think they're
doing, particularly when making epistemological moves like
\emph{supposing} might help students work through design problems.

My findings also speak to assessment. In the course I studied, there
were few (if any) assessment instruments that would have revealed the
kinds of knowledge I document Lionel and Rebecca having in this study.
At no point in my observational data of the course were students asked
to explain a design choice or reason about competing solutions to handle
a problem. At no point were they encouraged to deeply sense-make about
the conceptual content of the course, particularly the traditionally
tricky topics of pointers and pointer arithmetic. At no point in my data
did the instructor model the kind of sense-making Rebecca did in her
interviews, a fact which she noted at one point as distinguishing the
course from Basic Programming, its first-semester counterpart. Because
the course never explicitly asked students to sense-make, reason about
design choices, or explain designs, it never used information about what
knowledge students had about those things to inform its instruction. It
couldn't. Moreover, because the course never explicitly asked students
to sense-make, reason about design choices, or explain designs, I was
left to conclude such things were not prioritized learning outcomes of
the course. To be clear, all courses establish a kind of focus by
deciding what won't be covered, and I don't fault the instructor at all
for running a course where design knowledge wasn't an enacted learning
outcome. But, I'm also left to wonder: why wasn't it? Given that even
beginning students can think about software design, and given that
engineering as a discipline fundamentally involves design, shouldn't it
have been?

\section{Conclusion}\label{conclusion-2}

If there is one overarching finding from this dissertation, it's that
students clearly have resources for thinking about designing programs
and a diversity of approaches to programming in the moment. Below, I
explain the kind of diversity I saw in programming approaches. Then, I
conclude with a discussion of what my research might mean for
assessment.

\subsection{We should think carefully about what students' programming
design knowledge means for
assessment}\label{we-should-think-carefully-about-what-students-programming-design-knowledge-means-for-assessment}

My research helps us document and model the kinds of knowledge students
have. In so doing, it points out kinds of information that the course's
assessments were apt to miss:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  How students frame or otherwise approach the task of programming in
  the moment
\item
  The role of different kinds of prior experience in stabilizing (or
  potentially destabilizing) certain kinds of frames
\item
  In-the-moment practices---including talking out solutions, sketching
  out debugging strategies, and writing out pseudo-code---that display
  students' competence and sense-making
\item
  The code history that traces how designs evolve, including how
  students start projects and what parts of their designs become
  dead-ends
\end{enumerate}

Together, those four points are relevant for formative assessment (Black
\& Wiliam, 1998) in introductory programming. (1) and (2) point us
toward what students think they're doing when they're programming.
Analogous work from science education tells us that students' sense of
the kind of knowledge activity they're enacting matters for learning and
assessment (Russ et al., 2008). For example, knowing early on that
students are blindly copying code or randomly trying syntax gives
instructors a chance to intervene. But, my research suggests
intervention can't just be about stopping a bad behavior: knowing that
copy-paste behavior is happening is different from knowing \emph{why}
it's happening.

Formative assessment has to be about diagnosing causes, not just
identifying symptoms. Otherwise, we run the risk of ignoring or even
harming students' productive knowledge. Take the example of copy-paste
behavior, as documented by Gaspar and Langevin (Gaspar \& Langevin,
2007). One reason we may see copy-paste behavior, as Study 1
demonstrates, is that students might see situations as new instances of
already-solved problems. In professional practice, seeing old solved
problems in new situations can be productive. Such insights can, for
example, direct engineers to use a pre-built library of functions
instead of building their own. But, another related reason for
copy-paste may be efficiency. For Rebecca, copying and pasting code was
very fast; it required only a few quick keystrokes. In the short run,
Rebecca's choice to copy the code was a faster, less demanding, more
trustworthy route to go on than was abstracting the code to a function.
When we consider students trying to see common problems in new scenarios
and solve such problems efficiently, ``copy-paste'' becomes a symptom
rather than a root cause. Epistemological frameworks, then, can inform
assessments by offering explanations that aim at the root cause of
certain behaviors.

Points 3 and 4 complement knowledge analysis by drawing focus to
artifacts, practices, and history. As data from Lionel shows (Study 2),
a crucial part of his design process involves artifacts and activities
that were only distally knowable to the assessments he got in class.
Lionel's instructor had no direct access to:

\begin{itemize}
\item
  Lionel's whiteboard
\item
  the hours Lionel may have spent working out designs in chalk
\item
  how Lionel might have talked out design features
\item
  how Lionel's initial ``pseudo-code'' evolved into his final design.
\end{itemize}

In Rebecca's case, commented-out code in her final project submission
were perhaps the only clues at all that she tried abstracting some
flight day-checking procedures into functions (Study 1). Without
Rebecca's history, the instructor would have had no reliable way of
knowing:

\begin{itemize}
\item
  Rebecca began her flights database project by copying code from an
  earlier project
\item
  Rebecca's original solution for day-checking evolved from a seven-fold
  conditional structure, one for each day
\item
  Rebecca may have tried abstracting day-checking procedures, even
  creating a function called check\_days.
\end{itemize}

To sum up: traditional assessments can tell us what students finally
produce but not how they produce it, when they produce it, why they
produce it, or what the production process was. In the summative
assessment structure of the course, final student products were the only
products submitted to the instructor. My research shows that what's
happening in the interstices---before typing, between compiles, away
from the computer---can be captured and, in principle, analyzed and
acted upon by instructors.

\subsection{What We Might Change About Classroom
Practice}\label{what-we-might-change-about-classroom-practice}

What follows is my speculation about how we might specifically change
classroom practices in light of my research findings. I move from
recommendations I think are most strongly supported by my data to
recommendations that align with my findings but are more expansive, and
thus less strongly supported. As my recommendations broaden, I try to
offer not only an instructional recommendation but a concomitant
question for research.

\subsubsection{Instructors could look beyond content to understand
student
difficulties}\label{instructors-could-look-beyond-content-to-understand-student-difficulties}

I think first and foremost, instructors have to have a willingness to
recognize that the difficulties they think they're seeing in students
are difficulties they may be viewing through a lens of content. For
example, after seeing a student struggle an instructor might say ``the
student doesn't know assignment statements.'' But, those difficulties
almost certainly have a deeper explanation. I say ``almost certainly,''
where what I mean is that there are a number of patterns we've
identified as common novice errors, but most research stops before
asking why that's a common novice error. With rare exceptions (Fleury,
1991, 2000), the computing education community doesn't encourage asking
the question \emph{why might a student be doing this?} and \emph{what
might this tell me about the way a student is making sense of this?} By
contrast, my research suggests those orientations are ripe, low-hanging
fruit for instructors to take on. I think one of the first things as
instructor could ask is, \emph{why would a student be thinking that this
is the appropriate thing to be doing?} And that holds true whether the
thing in question is writing the statement this way, or interpreting the
code this way, or enacting this kind of programming activity this way.

Moreover, we know from research on metacognition (Schoenfeld, 1987,
1992) that thinking beyond content opens the palette of kinds of
interventions an instructor can make. Schoenfeld (1987), for example,
came to encourage metacognition in his classroom by doggedly asking
student groups what they were doing, why they were doing it, and how
they hoped it would lead them to a solution. Students ultimately came to
internalize such strategies and spent far less time floundering along
solution paths that weren't ultimately productive. Might the same idea
be true in introductory programming courses? To find out, we could begin
researching in earnest how often and in what ways instructors model
metacognition for their students. Currently, I would argue, we don't
know whether and how instructors do so.

\subsubsection{Instructors could use code history to inform
interventions}\label{instructors-could-use-code-history-to-inform-interventions}

When, for example, a student comes to office hours with a problem on a
project, an instructor needs to quickly come to grips with the state of
the students' project, the logic of their design, where they're stuck,
why they're stuck, and what might best help them. That's no small task.
But, given my analysis I have strong reason to believe that having code
history available to instructor could change both the nature of coming
to terms with a student's project and the conversation with a student
that results. If an instructor can see the evolution of a student's
code, at the very least they could see where the student started, how
the code was growing, and where the student was working most recently.
Even if the instructor had never before seen the code (or its history)
until that moment in office hours, having both available changes the
kind of view the instructor can get of the code and the specifics of the
intervention that might result. In Rebecca's case, for example, it might
have been an opportunity to explore why her check\_days abstracted
function was failing on her flights database project.

Having code-snapshot capabilities also changes the kind of research we
can do. First and foremost, a result of this dissertation was to create
a freely-available, lightweight, open-source framework for capturing and
visualizing students' code histories.\footnote{https://github.com/briandk/gitvisualizations}
So, the most basic kind of study would involve deploying that framework
from the instructional side of a course and exploring what happens, as,
for example, Hurd (2013) has done. One could ask questions of how
instructors build code snapshotting into their course, how such
information could or did inform assessment, and how such information
could or did change the nature of instructional interactions with
students. Was it for the better? If so, how do we know? If not, how do
we know?

\subsubsection{Instructors could establish a norm of asking why a design
choice makes
sense}\label{instructors-could-establish-a-norm-of-asking-why-a-design-choice-makes-sense}

The most far-reaching implication of the research I present here is that
instructors should establish a norm in their classes where anyone, at
any point, for any piece of code, is allowed to ask \emph{why does this
make sense as a design choice? Why is that an obvious choice to make?
How does that work?} Consider a parallel example from mathematics
education. In Don Saari's calculus class at UC Irvine, he ``invokes the
principle of what he calls ``WGAD''---``Who gives a damn?'' (Bain, 2004,
pp. 38--39). Bain explains:

At the beginning of his courses, he tells his students that they are
free to ask him the question on any day during the course, at any moment
in class. He will stop and explain to his students why the material
under consideration at that moment---however abstruse and minuscule a
piece of the big picture it may be---is important, and how it relates to
the larger questions and issues of the course. (Bain, 2004, p. 39)

What I'm suggesting is even broader than that, because it's not just a
question students can ask of professors; it's a question \emph{anyone}
can ask of \emph{anyone.} I see it as the programming and design
extension of a sociomathematical norm (Yackel \& Cobb, 1996), and I
think it could lead to collaborative sense-making. That is, I'm asking
for an accepted cultural practice that is also itself a design practice,
whereby an instructor can help create a safe, stable space for a
community to be reflective and critical about design choices. I also
think it's an idea that leads to others, such as letting, if not
requiring, students review one another's code. When students are forced
to reckon with someone else's code and understand their design
decisions, they're also forced to justify their own decisions. And
establishing ``why that design choice?'' or ``how does that make
sense?'' as a norm provides reciprocal opportunities for students, not
just instructors, to improve the code of others.

The research questions that come out of such an idea would include:

\begin{itemize}
\item
  How can an instructor satisfactorily establish norms about design in a
  classroom?
\item
  What does it look like when students collaboratively sense-make about
  a program's design choices?
\item
  What does it look like when students are asked to reflect on their own
  design choices?
\item
  How might introducing a code review component into a course change the
  way students approach design? How might it improve students'
  conceptual understanding? How might it improve the quality of the code
  students produce?
\end{itemize}

\subsection{Final Remarks}\label{final-remarks}

How students design programs matters for learning and instruction in
engineering. It matters because finished code reflects what students
know about design, whether or not instructors capture such information.
It matters because students have resources for learning about and
engaging in design; whether or not curriculum, instruction, and
assessment choose to tap into those resources. It matters because design
should be an intellectual thread that runs through all engineering
courses. That thread shouldn't stop when we introduce students to
programming.

\section{Appendix 1 -- Transcript
conventions}\label{appendix-1-transcript-conventions}

\begin{itemize}
\item
  Turns are not numbered, but they are blank-line-delimited
\item
  Short interjected speech that does not interrupt a speaker's turn is
  bounded /by slashes/
\item
  Matching double brackets show the {[}{[}onset and termination{]}{]} of
  overlapping talk across turns.
\item
  *emphasized speech* is bounded by asterisks
\item
  Parenthetical clarifications by the analyst appear in parentheses. (He
  considered but rejected square brackets.)
\item
  matching double equals signs mark turn boundaries== ==with minimal or
  no audible silence (also known as latching turns)
\item
  Where gestures don't overlap speech they are in-lined by curly braces
  when they happen \{folds arms, having made his point\}. All gestures
  enacted \emph{by} a speaker appear within that speaker's turn unless
  otherwise noted \{smiles, self-satisfied at having made this important
  clarification\} \{audience scoffs\}.
\item
  When gestures happen \emph{during} speech, the speech is presented
  first and bounded by double pipes. Gestures that happen simultaneously
  with such speech are bounded by curly braces, nested within double
  pipes, and immediately follow the speech they overlap. For example:
\end{itemize}

Throwing chainsaws \textbar{}\textbar{}*up*\textbar{}\textbar{}
\textbar{}\{throws chainsaw\}\textbar{} is easy. Just be careful when
they come \textbar{}\textbar{}*down*\textbar{}\textbar{}
\textbar{}\{catches chainsaw for a punctuated finish\}\textbar{}.

\section{Appendix 2 -- Visual conventions for
gestures}\label{appendix-2-visual-conventions-for-gestures}

A challenge of this dissertation was trying to use still-frames to
convey motion and animation. Where I could, I annotated pictures with
arrows to show trajectories of movement:

\includegraphics[width=5.62222in,height=0.87778in]{media/image38.emf}

\begin{longtable}[]{@{}ll@{}}
\toprule
\textbf{Description} & \textbf{Example}\tabularnewline
\midrule
\endhead
\textbf{Blue} arrows project forward to show action that will occur but
hasn't yet in the frame you're looking at. The origin is where a hand
\emph{is}; the terminus is where it \emph{will be} shortly. &
\includegraphics[width=1.67778in,height=1.58889in]{media/image39.png}\tabularnewline
\textbf{Pink} arrows show action that already occurred before the frame
you're looking at. The origin is where a hand \emph{was}; the terminus
is where it \emph{is}. &
\includegraphics[width=1.65556in,height=2.37778in]{media/image20.png}\tabularnewline
\bottomrule
\end{longtable}

\section{\texorpdfstring{Appendix 3 -- Transcript of Rebecca's
pseudo-code episode \emph{without} gesture
codes}{Appendix 3 -- Transcript of Rebecca's pseudo-code episode without gesture codes}}\label{appendix-3-transcript-of-rebeccas-pseudo-code-episode-without-gesture-codes}

Interviewer: OK. So then, um, what if we just pretended for a minute,
that that, like==

Rebecca: ==That that works=

Interviewer: ==That that worked==

Rebecca: ==OK==

Interviewer: ==OK. So then, um. So then you might write like to make an
array of \{stops writing\} actually what would we call this? This is==

Rebecca: ==array of pointers, I guess==

Interviewer: ==OK. Pointers. It'd be, well. So. The---the one we had was
like int, star p, um, it would be some number, um, and I guess that'd be
it in order to just declare /Yes/ that right? OK. And then you're
saying, how would you like access an element of it?

Rebecca: Uh, well, what I was think---like if I wanted to like, save it
or whatever, /Yeah/ I could make a while loop, scan in the---scan in the
data from the album, /OK/ uh, which is, I can write it {[}if you want

Interviewer: Sure, go ahead{]}{]}

Rebecca: Um, so, it was==

Interviewer: {[}{[}Oh, sorry, just have to{]}

Rebecca: {[}Oh{]}

Interviewer: It's /Oh, OK/ one of those annoying ticks, but it works
best cause of the camera if, ah, /Ah, OK/ if that's pointing away from
your hand

Rebecca: {[}OK{]}

Interviewer: {[}That's fine{]}

Rebecca: Um, there was album, and then, so there was, uh, percent d, and
percent s, and then give those names, just, I'll just call it number,
and title. /Mmmhmm/ And so, at least, my---until it does not equal EOF,
and then my thinking at least, is you should be able to, um, say that

Rebecca: ``star p of i'' /mmhmm/ equals, uh, the title, and then you
just do i++, so then it'll move to the next one /OK/ and you just keep
saving each of the pointers in the array to a title /Mmmhmm/ And you
just increment by 1 until you reach the end of file.

Interviewer: OK==

Rebecca: ==Like *that* would make sense to me.

Interviewer: So then, like, what would go in p of one would be the first
title we read in==

Rebecca: ==Yes,

Interviewer: uh, would, when the while loop runs again, does it get a
fresh line {[}{[}line from that

Rebecca: Yeah, because{]}{]} uh, what the while loop does is it reads in
the line and then once it reaches the end of character it'll go back
down---it'll do the while loop and then go back down to the next
thing---line, it'll read in the line until it reaches the end of file

Interviewer: So as you step through this loop /mmhmm/ i keeps going up
by one==

Rebecca: ==Yes.

Interviewer: Uh

Rebecca: And the lines keep going down so, that way, the first line is
going to be, uh the p---element zero /OK/ uh the second one'll be
element one (Interview 4 of 5, April 6, 2012)

\section{Appendix 4 -- Conceptual knowledge frameworks in computing
don't tell us much about how Lionel structured an in-interview
program}\label{appendix-4-conceptual-knowledge-frameworks-in-computing-dont-tell-us-much-about-how-lionel-structured-an-in-interview-program}

In this section, I explore Lionel's work in solving an in-interview
programming task. First, I reproduce the prompt Lionel was given. Next,
I show the final source code of his solution to the problem, written in
C. In the analysis that follows, I move outward to explore the context
of that code's production.

My choice to start with the code first is deliberate. I'm trying to
mimic what the instructor of a typical programming course might see: the
final submitted form of a student's code. This approach is admittedly a
bit backwards, since the data I show after I present Lionel's submission
is all about what happened \emph{before} the code reached its final
form. But, I think this approach useful because it invites us to explore
an artifact first, then raise questions about the conditions of its
production. In that sense, my presentation has the anthropological feel
of understanding a found object, which I think is a faithful way of
looking at what many instructors (not to mention researchers, and of
course active software developers) face in their day-to-day work.

\subsection{\texorpdfstring{\protect\hypertarget{ux5fToc247188555}{}{\protect\hypertarget{ux5fToc252445950}{}{}}Analyzing
Lionel's Solution From a Conceptual Knowledge
View}{Analyzing Lionel's Solution From a Conceptual Knowledge View}}\label{analyzing-lionels-solution-from-a-conceptual-knowledge-view}

To infer conceptual knowledge, I'll use Elliott Tew's (2010) conceptual
categories for procedural programming in first-semester computer science
(CS1) courses. My reasons for doing so are:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Elliott Tew (2010) sought concepts that were language-agnostic. That
  is, the concepts Elliott Tew (2010) identified are not idiomatic to
  just one language. Rather, they are general ideas implemented in
  nearly every major procedural language taught to undergraduates (Java,
  Python, C-based languages, Scheme).
\item
  Elliott Tew's (2010, pp. 23--27) selection process was extensive. She
  began with curricular documents, moved to canonical texts, then
  ultimately conducted a thematic analysis on emerging categories to
  produce her final list of concepts.
\item
  To date, Elliott Tew's (2010) identification of CS1 concepts is the
  only work I know of that has been carried forward to create,
  administer, and validate concept inventories in CS1 (Elliott Tew \&
  Guzdial, 2010, 2011)
\end{enumerate}

Figure 2 below presents an overview of the conceptual features I
identified in Lionel's code. Even in a short, relatively simple program
Lionel exhibits four of the ten concept categories Elliott Tew (2010)
identifies. Moreover, he uses each of them ``correctly,'' so to speak,
in that not only does his program compile, it also properly finds the
range of the numbers it's given. It's also worth noting that Lionel's
code displays consistent, helpful use of whitespace (indentation) as
well as informative comments. While these attributes aren't conceptual,
per se, they are the sorts of features an instructor might consider in
assigning a grade to this program.

\includegraphics[width=5.60000in,height=5.67778in]{media/image40.jpeg}

\protect\hypertarget{ux5fToc252445975}{}{}Figure ‑ -- Highlighting the
conceptual features of Lionel's code given Elliott Tew's (2010) outline
of CS 1 concepts

\section{Appendix 5 -- Neverly-Asked Questions
(NAQs)}\label{appendix-5-neverly-asked-questions-naqs}

\subsubsection{Neverly Asked Questions (NAQs) about my conceptual
analysis of Lionel's
code}\label{neverly-asked-questions-naqs-about-my-conceptual-analysis-of-lionels-code}

\emph{Didn't Elliott Tew (2010) define those concepts as a way of
creating a concept inventory?}

Yes. In her work, I believe each concept was represented by at least one
multiple-choice question (MCQ) with carefully-chosen distractors. In
each MCQ a student would see a snippet of pseudo-code and be asked to
choose from a list of responses. The structure was explicitly modeled
after Force Concept Inventory (FCI) work in physics education.

\emph{So, if her intent (and research) was about making a concept
\textbf{inventory}, aren't you misusing her ideas of concepts?}

What makes you say that?

\emph{Well, first off, she used categories of concept to create concrete
instances of code, which she tests students on}\ldots{}

That's my understanding, yep.

\emph{But you used a concrete instance of code --- Lionel's code --- and
inferred the existence of concepts in it. So, that's not the same thing
as what Elliott Tew (2010) did.}

I agree I'm not doing what she did, but I don't think that observation
necessarily undermines the usefulness of what I \emph{am} doing. If the
concepts she identifies really are general concepts, then it would seem
silly to think we couldn't find concrete instances of code that
exemplify them. I think I might rephrase your objection: ``how can I
trust that you properly identified concepts in Lionel's code?'' My
answer to that is, again, yes, I have no coding scheme from Elliott Tew
to apply. But, the code Lionel wrote is there and my candidates for
assigning concepts are there: my results are open to inspection and
challenge.

\emph{OK. Fine. But what ``conceptual knowledge analysis'' is there to
speak of? You just made a slide and identified which parts of his code
might correspond to which concepts.}

What I did was assign concepts to canonically-correct --- that is,
syntactically valid parts of Lionel's code that also work properly to
make the program produce the right results --- sections of Lionel's
code. I think your objection is about my inferential basis: ``why does
the presence of those code chunks --- and my tagging certain chunks as
aligning with Elliott Tew's (2010) concepts --- imply that Lionel has
those concepts in his head?''

\emph{I accept your amendment. So, what basis do you have for saying
``this concept is in Lionel's mind?''}

I feel like I'm on the same playing field as the concept inventory
folks, though in some ways I'm setting a modified standard for validity.
Elliott Tew (2010) says this about validity:

validity is the evidence that assures us that questions about a
particular concept are indeed measuring that concept. For instance, a
question about arrays should require a student to have knowledge about
arrays, but should not require knowledge about another concept, such as
recursion. In addition, it is important that the question cannot be
answered correctly without knowledge of arrays. (Elliott Tew, 2010, p.
10)

I don't know that I share her assumptions about validity. But, what
specifically might be at issue here is ``does the presence of a for-loop
in code imply the student has knowledge of definite loops?'' My answer
is, ``if not that, then what?''

\emph{But couldn't I argue that the for-loop could be there because he
copied and pasted it from the internet? That he wouldn't have had to
know anything about definite loops to do that?}

Yes, you could. My rejoinder to that is in section 3.3.3.

\subsubsection{Neverly-asked questions about Lionel's verbal pseudo-code
description}\label{neverly-asked-questions-about-lionels-verbal-pseudo-code-description}

Why do you so pedantically analyze such a short, small snippet of
conversation?

Because by Lionel's own assertion what was said constitutes his ``main
concept'' (line 12) for the program. That is, by his own account he has
just expressed the entire top-level design for the program. But, he's
done it all all without writing a single additional line of code.

So?

If anything, this episode suggests that Lionel can and does have ways of
expressing a process or procedure that don't rely on writing code. More
importantly, I think, is that he views it as a legitimate activity to
express ideas at a level \emph{above} the particularities of syntax and
implementation.

What do you mean ``views it as a legitimate activity?''

I think here I'm appealing to my interpretation of Goffman's (1974)
\emph{frames}: a participant's understanding of ``what is it'' that's
going on in a social situation. Lionel takes up my bid for him to
explain what he's thinking, but he does so first by explaining his
``general concept'' for the program.

Right. But, I mean, he's just doing what you \textbf{asked} him to do. I
don't understand how any of this relates to Goffman or frames. That just
seems like a needlessly complicated way to explain he did what you
asked.

OK, let's step back for a second. Suppose you've just started going to
group therapy. One of the things group therapy environments often stress
is that you should try to talk about how \emph{you feel}, which can be a
hard transition for people. Group therapy participants may be used to
saying ``you always do {[}X{]}'' when in a fight with someone. And, if
prompted by a group leader to discuss how they feel about troublesome
instance, a participant may adopt that kind of language. ``My partner
always yells at me if I'm out late,'' for example.

Group therapy can encourage inward reflection in a way that reshapes how
participants talk about themselves and their feelings. So, after some
time in group therapy our hypothetical participant may change the way
they orient toward questions. If asked ``how are you feeling about the
fight you and your partner had?'' our hypothetical participant may now
say ``I feel like I don't have the freedom to see the people I want to
see.'' The focus in speech shifts from the \emph{other}---in this case a
partner---to the \emph{self}. Moreover, the substance of the speech is
about feelings, rather than behavior.

I think these reshapings of speech are neither trivial nor incidental.
If we had transcripts of our hypothetical person's early sessions in
group therapy and compared them to those of later sessions, I think we
would not be surprised to see a marked change. When asked the same
question---in this case ``how are you feeling about {[}X{]},'' the
structure and character of our participant's response changes after
therapy.

My point here is that an everyday sense of how participants' speech
patterns change in therapy primes our intuition for thinking more
broadly about social phenomena. A cynical view of our participant would
still have to acknowledge that, observably, the structure of their
response to the question ``how does that make you feel?'' changes after
therapy. I think a sociolinguistic stance would be more generous: our
participant is \emph{orienting} differently toward the question.
Furthermore, I take the following as evidence of such an orientation:

\begin{itemize}
\item
  A change in sentence focus from the other to the self
\item
  Introspection into the source of a feeling
\item
  Acknowledgment of personal responsibility for feeling those feelings
\end{itemize}

I think we as analysts can acknowledge that after therapy, our
participant frames the event of being asked, ``how does that make you
feel?'' differently.

OK. So, you're saying that if we look at the change in how someone
answers the question ``how does that make you feel?'' we can find
evidence of the superstructure---a frame---that directs their sense of
appropriate ways to answer that question?

Yes.

So how does this get back to your claims about Lionel?

I think it's tempting to take this position that, ``it's
inevitable/totally expected that he would just verbalize the top-level
description of his procedure! I mean, what else could he possibly do?''

I mean, yeah. Kind of.

Right, but I think actually that assumption isn't in line with a classic
frame analysis approach. Even if the assumption is right, frame analysis
changes the explanation from ``it's inevitable'' to ``hmm, I guess
almost everyone I've ever seen brings or enacts the same structure of
expectations.''

So, my argument is, if you ask someone to describe to you how they're
thinking about their program, they're gonna do that. What else
\textbf{would} they do?

Right. And I don't think we disagree that they'll describe their
thinking. I think where we disagree is that everyone will describe in
the same way. I'm saying \emph{if} they describe their program to you,
\emph{that} they're able to express it, what they choose to verbalize,
what they choose to leave out, how their gestures accompany the
explanations, and their ultimate criteria for having satisfied the task
of describing for themselves are all driven underneath (or from above?)
by structures of expectations. And even if \emph{one thousand} out of
one thousand people do the same kind of thing Lionel does, that doesn't
undermine the idea that people have and bring structures of expectations
to a task. What I think it does is show that one thousand out of one
thousand people share the same sets of expectations about what it is
that's going on when they're asked to verbalize their thinking on a
programming protocol.

So, what's the point of this whole section on Lionel's verbal
description, then?

I think the point is that when I ask Lionel to verbalize his thinking,
he outlines through talk what the ``main concept'' of his program is. I
focus on \emph{that} he decides to describe it out in words, \emph{how}
he describes it, \emph{what} he leaves out\emph{, how} he repairs, and
how all of that reflects structures of expectations about the activity
and what he's being asked to do. Ultimately, the argument is that
Lionel's choice to take me up on that bid and how he does so
reflects---to me, the analyst---that he views \emph{verbal description
of a high-level program} as a legitimate activity and part of his
approach to programming.

\section{References}\label{references}

Abelson, H., \& Sussman, G. J. (1996). \emph{Structure and
Interpretation of Computer Programs} (2nd ed.). Cambridge, Mass: MIT
Press.

Adelson, B., \& Soloway, E. (1985). The Role of Domain Experience in
Software Design. \emph{IEEE Transactions on Software Engineering},
\emph{SE-11}(11), 1351 -- 1360. doi:10.1109/TSE.1985.231883

Archer, L. B. (1979). Whatever became of design methodology.
\emph{Design Studies}, \emph{1}(1), 17--18.

Bain, K. (2004). \emph{What the best college teachers do}. Cambridge,
Mass: Harvard University Press.

Baker, A., \& van der Hoek, A. (2010). Ideas, subjects, and cycles as
lenses for understanding the software design process. \emph{Design
Studies}, \emph{31}(6), 590--613. doi:10.1016/j.destud.2010.09.008

Ball, L. J., Onarheim, B., \& Christensen, B. T. (2010). Design
requirements, epistemic uncertainty and~solution development strategies
in software design. \emph{Design Studies}, \emph{31}(6), 567--589.
doi:10.1016/j.destud.2010.09.003

Bayman, P., \& Mayer, R. E. (1983). A diagnosis of beginning
programmers' misconceptions of BASIC programming statements.
\emph{Communications of the ACM}, \emph{26}(9), 677--679.
doi:http://doi.acm.org/10.1145/358172.358408

Black, P., \& Wiliam, D. (1998). Inside the Black Box: Raising Standards
Through Classroom Assessment. \emph{Phi Delta Kappan}, \emph{80}(2),
139--44.

Boaler, J. (1998). Open and closed mathematics: Student experiences and
understandings. \emph{Journal for Research in Mathematics Education},
\emph{29}(1), 41--62.

Boaler, J. (2000). Mathematics from another world: Traditional
communities and the alienation of learners. \emph{The Journal of
Mathematical Behavior}, \emph{18}(4), 379--397.
doi:10.1016/S0732-3123(00)00026-2

Boaler, J. (2002). The development of disciplinary relationships:
Knowledge, practice and identity in mathematics classrooms. \emph{For
the Learning of Mathematics}, \emph{22}(1), 42--47.

Boaler, J., \& Greeno, J. G. (2000). Identity, agency, and knowing in
mathematical worlds. In J. Boaler (Ed.), \emph{Multiple perspectives on
mathematics teaching and learning} (pp. 171--200). Westport, CT: Ablex
Pub.

Bonar, J., \& Soloway, E. (1983). Uncovering principles of novice
programming. \emph{Proceedings of the 10th ACM SIGACT-SIGPLAN Symposium
on Principles of Programming Languages}, 10--13.
doi:10.1145/567067.567069

Bonar, J., \& Soloway, E. (1985). Preprogramming Knowledge: A Major
Source of Misconceptions in Novice Programmers. \emph{Human-Computer
Interaction}, \emph{1}(2), 133.

Bucciarelli, L. L. (1994). \emph{Designing engineers}. Cambridge, Mass:
MIT Press.

Clancy, M. (2004). Misconceptions and Attitudes that Interfere with
Learning to Program. In S. Fincher \& M. Petre (Eds.), \emph{Computer
Science Education Research} (pp. 85--100). London, UK: RoutledgeFalmer.

Danielak, B. A., Gupta, A., \& Elby, A. (in press). The Marginalized
Identities of Sense-Makers: Reframing Engineering Student Retention.
\emph{Journal of Engineering Education}.

Danielsiek, H., Paul, W., \& Vahrenhold, J. (2012). Detecting and
Understanding Students' Misconceptions Related to Algorithms and Data
Structures. In \emph{Proceedings of the 43rd ACM Technical Symposium on
Computer Science Education} (pp. 21--26). New York, NY, USA: ACM.
doi:10.1145/2157136.2157148

diSessa, A. A. (1986). Models of Computation. In D. A. Norman \& S. W.
Draper (Eds.), \emph{User centered system design: new perspectives on
human-computer interaction} (pp. 201--218). Hillsdale, N.J: L. Erlbaum
Associates.

diSessa, A. A. (1993). Toward an Epistemology of Physics.
\emph{Cognition and Instruction}, \emph{10}(2/3), 105--225.

diSessa, A. A. (2002). Why ``Conceptual Ecology'' is a good idea. In M.
Limón \& L. Mason (Eds.), \emph{Reconsidering conceptual change: issues
in theory and practice} (pp. 29--60). Dordrecht ; Boston: Kluwer
Academic Publishers.

diSessa, A. A., \& Sherin, B. L. (1998). What changes in conceptual
change? \emph{International Journal of Science Education},
\emph{20}(10), 1155--1191. doi:10.1080/0950069980201002

Duckworth, E. R. (2006). \emph{``The having of wonderful ideas'' and
other essays on teaching and learning} (3rd ed.). New York: Teachers
College Press.

Elby, A., \& Hammer, D. (2010). Epistemological resources and framing: A
cognitive framework for helping teachers interpret and respond to their
students' epistemologies. In L. D. Bendixen \& F. C. Feucht (Eds.),
\emph{Personal epistemology in the classroom: theory, research, and
implications for practice} (pp. 409--434). Cambridge, UK ; New York:
Cambridge University Press.

Elliott Tew, A. (2010). \emph{Assessing fundamental introductory
computing concept knowledge in a language independent manner} (Ph.D.).
Georgia Institute of Technology, Ann Arbor. Retrieved from ProQuest
Dissertations \& Theses Full Text. (873212789)

Elliott Tew, A., \& Guzdial, M. (2010). Developing a validated
assessment of fundamental CS1 concepts. \emph{Proceedings of the 41st
ACM Technical Symposium on Computer Science Education}, 97--101.
doi:10.1145/1734263.1734297

Elliott Tew, A., \& Guzdial, M. (2011). The FCS1: a language independent
assessment of CS1 knowledge. In \emph{Proceedings of the 42nd ACM
technical symposium on Computer science education} (pp. 111--116). New
York, NY, USA: ACM. doi:10.1145/1953163.1953200

Erickson, F. (1986). Qualitative methods in research on teaching. In M.
C. Wittrock (Ed.), \emph{Handbook of research on teaching} (3rd ed., pp.
119--161). New York : London: Macmillan ; Collier Macmillan.

Eynde, P., \& Hannula, M. (2006). The Case Study of Frank.
\emph{Educational Studies in Mathematics}, \emph{63}(2), 123--129.
doi:10.1007/s10649-006-9030-8

Fleury, A. E. (1991). Parameter passing: the rules the students
construct. In \emph{Proceedings of the twenty-second SIGCSE technical
symposium on Computer science education} (pp. 283--286). New York, NY,
USA: ACM. doi:10.1145/107004.107066

Fleury, A. E. (1993). Student Beliefs about Pascal Programming.
\emph{Journal of Educational Computing Research}, \emph{9}(3), 355--371.
doi:10.2190/VECR-P8T6-GB10-MXJ5

Fleury, A. E. (2000). Programming in Java: student-constructed rules.
\emph{SIGCSE Bull.}, \emph{32}(1), 197--201. doi:10.1145/331795.331854

Gainsburg, J. (2006). The mathematical modeling of structural engineers.
\emph{Mathematical Thinking \& Learning}, \emph{8}(1), 3--36.
doi:10.1207/s15327833mtl0801\_2

Gal-Ezer, J., \& Zur, E. (2004). The efficiency of
algorithms-\/-misconceptions. \emph{Computers \& Education},
\emph{42}(3), 215--226.

Gaspar, A., \& Langevin, S. (2007). Restoring ``coding with intention''
in introductory programming courses. \emph{Proceedings of the 8th ACM
SIGITE Conference on Information Technology Education}, 91--98.
doi:10.1145/1324302.1324323

Ginsburg, H. P. (1997). \emph{Entering the child's mind: the clinical
interview in psychological research and practice}. Cambridge ; New York:
Cambridge University Press.

Ginsburg, H. P., \& Opper, S. (1988). \emph{Piaget's Theory of
Intellectual Development} (3rd ed.). Englewood Cliffs, N.J:
Prentice-Hall. Retrieved from http://lccn.loc.gov/87017353

Goffman, E. (1974). \emph{Frame Analysis: An Essay on the Organization
of Experience}. New York: Harper \& Row.

Goldin-Meadow, S. (2003). \emph{Hearing gesture: how our hands help us
think}. Cambridge, Mass: Belknap Press of Harvard University Press.

Goodwin, C. (2000). Action and embodiment within situated human
interaction. \emph{Journal of Pragmatics}, \emph{32}(10), 1489--1522.
doi:10.1016/S0378-2166(99)00096-X

Gupta, A., Hammer, D., \& Redish, E. F. (2010). The Case for Dynamic
Models of Learners' Ontologies in Physics. \emph{Journal of the Learning
Sciences}, \emph{19}(3), 285. doi:10.1080/10508406.2010.491751

Hall, R. (1999). Following mathematical practices in design-oriented
work. In C. Hoyles, C. Morgan, \& G. Woodhouse (Eds.), \emph{Rethinking
the Mathematics Curriculum} (pp. 29--47). London: Falmer Press.

Hall, R., \& Nemirovsky, R. (2012). Introduction to the Special Issue:
Modalities of Body Engagement in Mathematical Activity and Learning.
\emph{Journal of the Learning Sciences}, \emph{21}(2), 207--215.
doi:10.1080/10508406.2011.611447

Hall, R., \& Stevens, R. (1995). Making space: A comparison of
mathematical work in school and professional design practices. In S. L.
Star (Ed.), \emph{The cultures of computing} (pp. 118--145). Oxford, UK:
Blackwell Publisher.

Hall, R., Stevens, R., \& Torralba, T. (2002). Disrupting
representational infrastructure in conversations across disciplines.
\emph{Mind, Culture \& Activity}, \emph{9}(3), 179--210.

Hall, R., Wright, K., \& Wieckert, K. (2007). Interactive and Historical
Processes of Distributing Statistical Concepts Through Work
Organization. \emph{Mind, Culture \& Activity}, \emph{14}(1/2),
103--127. doi:10.1080/10749030701307770

Hammer, D. (1989). Two approaches to learning physics. \emph{The Physics
Teacher}, \emph{27}(9), 664--670. doi:10.1119/1.2342910

Hammer, D. (1994). Epistemological beliefs in introductory physics.
\emph{Cognition and Instruction}, \emph{12}(2), 151--183.
doi:10.2307/3233679

Hammer, D. (1996). Misconceptions or P-Prims: How May Alternative
Perspectives of Cognitive Structure Influence Instructional Perceptions
and Intentions? \emph{Journal of the Learning Sciences}, \emph{5}(2),
97--127. doi:10.1207/s15327809jls0502\_1

Hammer, D., \& Elby, A. (2002). On the form of a personal epistemology.
In B. K. Hofer \& P. R. Pintrich (Eds.), \emph{Personal epistemology:
The psychology of beliefs about knowledge and knowing} (pp. 169--190).
Mahwah, N.J: L. Erlbaum Associates.

Hammer, D., \& Elby, A. (2003). Tapping epistemological resources for
learning physics. \emph{The Journal of the Learning Sciences},
\emph{12}(1), 53--90. doi:10.2307/1466634

Hammer, D., Elby, A., Scherr, R. E., \& Redish, E. F. (2005). Resources,
framing, and transfer. In J. P. Mestre (Ed.), \emph{Transfer of learning
from a modern multidisciplinary perspective}. Greenwich, CT: IAP.

Hannula, M., Evans, J., Philippou, G., \& Zan, R. (2004). Affect in
Mathematics Education--Exploring Theoretical Frameworks. Research Forum.
\emph{International Group for the Psychology of Mathematics Education},
30.

Henderson, K. (1999). \emph{On line and on paper: visual
representations, visual culture, and computer graphics in design
engineering}. Cambridge, Mass: MIT Press.

Herman, G. L., Kaczmarczyk, L., Loui, M. C., \& Zilles, C. (2008). Proof
by incomplete enumeration and other logical misconceptions.
\emph{Proceeding of the Fourth International Workshop on Computing
Education Research}, 59--70. doi:10.1145/1404520.1404527

Hofer, B. K., \& Pintrich, P. R. (1997). The development of
epistemological theories: Beliefs about knowledge and knowing and their
relation to learning. \emph{Review of Educational Research},
\emph{67}(1), 88--140. doi:10.3102/00346543067001088

Holland, S., Griffiths, R., \& Woodman, M. (1997). Avoiding object
misconceptions. In \emph{Proceedings of the twenty-eighth SIGCSE
technical symposium on Computer science education} (pp. 131--134). New
York, NY, USA: ACM. doi:10.1145/268084.268132

Hurd, A. (2013, March). \emph{Assessment in an Introduction to
Programming Course}. Presented at the 16th Annual Course Technology
Conference, San Diego, CA, USA. Retrieved from
http://www.slideshare.net/CengageLearning/andrew-hurd-assessment-in-an-intro-to-programming-course

Hutchins, E. (1995a). \emph{Cognition in the Wild}. Cambridge, Mass: MIT
Press.

Hutchins, E. (1995b). How a cockpit remembers its speeds.
\emph{Cognitive Science}, \emph{19}(3), 265--288.
doi:10.1016/0364-0213(95)90020-9

Izsák, A. (2004). Students' Coordination of Knowledge When Learning to
Model Physical Situations. \emph{Cognition \& Instruction},
\emph{22}(1), 81--128.

Jackson, M. (2010). Representing structure in a software system design.
\emph{Design Studies}, \emph{31}(6), 545--566.
doi:10.1016/j.destud.2010.09.002

Jadud, M. C. (2006). Methods and tools for exploring novice compilation
behaviour. In \emph{Proceedings of the 2006 international workshop on
Computing education research - ICER '06} (p. 73). Canterbury, United
Kingdom. doi:10.1145/1151588.1151600

Jordan, B., \& Henderson, A. (1995). Interaction analysis: Foundations
and practice. \emph{Journal of the Learning Sciences}, \emph{4}(1), 39.
doi:10.1207/s15327809jls0401\_2

Joy, M., Sinclair, J., Sun, S., Sitthiworachart, J., \& López-González,
J. (2009). Categorising computer science education research.
\emph{Education and Information Technologies}, \emph{14}(2), 105--126.
doi:10.1007/s10639-008-9078-4

Kaczmarczyk, L. C., Petrick, E. R., East, J. P., \& Herman, G. L.
(2010). Identifying student misconceptions of programming.
\emph{Proceedings of the 41st ACM Technical Symposium on Computer
Science Education}, 107--111. doi:10.1145/1734263.1734299

Kaiser, D. (2005). \emph{Drawing theories apart: the dispersion of
Feynman diagrams in postwar physics}. Chicago: University of Chicago
Press.

Keller, E. F. (1983). \emph{A feeling for the organism: the life and
work of Barbara McClintock}. San Francisco: W.H. Freeman.

Kolikant, Y. B.-D., \& Mussai, M. (2008). ``So my program doesn't run!''
Definition, origins, and practical expressions of students'
(mis)conceptions of correctness. \emph{Computer Science Education},
\emph{18}(2), 135.

Kölling, M., \& Utting, I. (2012). Building an open, large-scale
research data repository of initial programming student behaviour. In
\emph{Proceedings of the 43rd ACM technical symposium on Computer
Science Education} (pp. 323--324). New York, NY, USA: ACM.
doi:10.1145/2157136.2157234

Latour, B. (1987). \emph{Science in action: how to follow scientists and
engineers through society}. Cambridge, Mass: Harvard University Press.

Latour, B. (1990). Drawing things together. In M. Lynch \& S. Woolgar
(Eds.), \emph{Representation in Scientific Practice} (1st MIT Press ed.,
pp. 19--68). Cambridge, Mass: MIT Press.

Lehrer, R., Schauble, L., Carpenter, S., \& Penner, D. (2000). The
interrrelated development of inscriptions and conceptual understanding.
In P. Cobb, E. Yackel, \& K. McClain (Eds.), \emph{Symbolizing and
Communicating in Mathematics Classrooms: Perspectives on Discourse,
Tools, and Instructional Design} (pp. 325--360). Mahwah, N.J: Lawrence
Erlbaum Associates.

Lising, L., \& Elby, A. (2005). The impact of epistemology on learning:
A case study from introductory physics. \emph{American Journal of
Physics}, \emph{73}(4), 372. doi:10.1119/1.1848115

Louca, L., Elby, A., Hammer, D., \& Kagey, T. (2004). Epistemological
Resources: Applying a New Epistemological Framework to Science
Instruction. \emph{Educational Psychologist}, \emph{39}(1), 57--68.

Malmi, L., Sheard, J., Simon, Bednarik, R., Helminen, J., Korhonen, A.,
\ldots{} Taherkhani, A. (2010). Characterizing research in computing
education: a preliminary analysis of the literature. In
\emph{Proceedings of the Sixth international workshop on Computing
education research} (pp. 3--12). New York, NY, USA: ACM.
doi:10.1145/1839594.1839597

Martin, R. C. (2009). \emph{Clean code: a handbook of agile software
craftsmanship}. Upper Saddle River, NJ: Prentice Hall.

Mayer, R. E. (1981). The Psychology of How Novices Learn Computer
Programming. \emph{ACM Computing Surveys (CSUR)}, \emph{13}, 121--141.
doi:10.1145/356835.356841

Minsky, M. L. (1986). \emph{The Society of Mind}. New York: Simon and
Schuster.

Nasir, N. S., \& Cooks, J. (2009). Becoming a Hurdler: How Learning
Settings Afford Identities. \emph{Anthropology \& Education Quarterly},
\emph{40}(1), 41--61. doi:10.1111/j.1548-1492.2009.01027.x

Nasir, N. S., \& Hand, V. (2008). From the court to the classroom:
Opportunities for engagement, learning, and identity in basketball and
classroom mathematics. \emph{Journal of the Learning Sciences},
\emph{17}(2), 143--179. doi:10.1080/10508400801986108

Nasir, N. S., \& Hand, V. M. (2006). Exploring Sociocultural
Perspectives on Race, Culture, and Learning. \emph{Review of Educational
Research}, \emph{76}(4), 449--475.

Nemirovsky, R., Rasmussen, C., Sweeney, G., \& Wawro, M. (2012). When
the Classroom Floor Becomes the Complex Plane: Addition and
Multiplication as Ways of Bodily Navigation. \emph{Journal of the
Learning Sciences}, \emph{21}(2), 287--323.
doi:10.1080/10508406.2011.611445

Ochs, E., Gonzales, P., \& Jacoby, S. (1996). ``When I come down I'm in
the domain state'': grammar and graphic representation in the
interpretive activity of physicists. In \emph{Interaction and Grammar}
(pp. 328--369). Cambridge: Cambridge University Press.

Papert, S. (1980). \emph{Mindstorms: Children, Computers, and Powerful
Ideas}. New York: Basic Books. Retrieved from
http://lccn.loc.gov/79005200

Parsons, J., \& Saunders, C. (2004). Cognitive heuristics in software
engineering: Applying and extending anchoring and adjustment to artifact
reuse. \emph{IEEE Transactions on Software Engineering}, \emph{30}(12),
873 -- 888. doi:10.1109/TSE.2004.94

Patitsas, E., Craig, M., \& Easterbrook, S. (2013). On the Countably
Many Misconceptions About \#Hashtables (Abstract Only). In
\emph{Proceeding of the 44th ACM Technical Symposium on Computer Science
Education} (pp. 739--739). New York, NY, USA: ACM.
doi:10.1145/2445196.2445443

Paul, W., \& Vahrenhold, J. (2013). Hunting High and Low: Instruments to
Detect Misconceptions Related to Algorithms and Data Structures. In
\emph{Proceeding of the 44th ACM Technical Symposium on Computer Science
Education} (pp. 29--34). New York, NY, USA: ACM.
doi:10.1145/2445196.2445212

Pea, R. D. (1986). Language-independent conceptual`` bugs'' in novice
programming. \emph{Journal of Educational Computing Research},
\emph{2}(1), 25--36.

Pea, R. D., Soloway, E., \& Spohrer, J. C. (1987). The Buggy Path to the
Development of Programming Expertise. \emph{Focus on Learning Problems
in Mathematics}, \emph{9}(1), 5--30.

Petre, M., van der Hoek, A., \& Baker, A. (2010). Editorial.
\emph{Design Studies}, \emph{31}(6), 533--544.
doi:10.1016/j.destud.2010.09.001

Rodrigo, M. M. T., \& Baker, R. S. J. d. (2009). Coarse-grained
detection of student frustration in an introductory programming course.
In \emph{Proceedings of the fifth international workshop on Computing
education research workshop} (pp. 75--80). New York, NY, USA: ACM.
doi:10.1145/1584322.1584332

Rodrigo, M. M. T., Tabanao, E., Lahoz, M. B. ., \& Jadud, M. C. (2009).
Analyzing Online Protocols to Characterize Novice Java Programmers.
\emph{Philippine Journal of Science}, \emph{138}(2), 177--190.

Rooksby, J. (2010). ``Just try to do it at the whiteboard'':
Researcher-participant interaction and issues of generalisation. In
\emph{Proceedings of the Studying Professional Software Design (SPSD)
Conference}. San Diego, CA, USA.

Rooksby, J., \& Ikeya, N. (2012). Collaboration in Formative Design:
Working Together at a Whiteboard. \emph{IEEE Software}, \emph{29}(1), 56
--60. doi:10.1109/MS.2011.123

Rosenberg, S., Hammer, D., \& Phelan, J. (2006). Multiple
Epistemological Coherences in an Eighth-Grade Discussion of the Rock
Cycle. \emph{Journal of the Learning Sciences}, \emph{15}(2), 261--292.
doi:10.1207/s15327809jls1502\_4

Russ, R. S., Coffey, J. E., Hammer, D., \& Hutchison, P. (2008). Making
classroom assessment more accountable to scientific reasoning: A case
for attending to mechanistic thinking. \emph{Science Education},
\emph{93}(5), 875--891. doi:10.1002/sce.20320

Saussure, F. de. (1986). \emph{Course in general linguistics}. LaSalle,
Ill: Open Court.

Scherr, R. E., \& Hammer, D. (2009). Student Behavior and
Epistemological Framing: Examples from Collaborative Active-Learning
Activities in Physics. \emph{Cognition \& Instruction}, \emph{27}(2),
147--174. doi:10.1080/07370000902797379

Schoenfeld, A. H. (1987). What's all the fuss about metacognition? In
\emph{Cognitive Science and Mathematics Education} (pp. 189--215).
Lawrence Erlbaum Associates.

Schoenfeld, A. H. (1988). When good teaching leads to bad results: The
disasters of ``well-taught'' mathematics courses. \emph{Educational
Psychologist}, \emph{23}(2), 145--166. doi:10.1207/s15326985ep2302\_5

Schoenfeld, A. H. (1991). On mathematics as sense-making: An informal
attack on the unfortunate divorce of formal and informal mathematics. In
\emph{Informal reasoning and education} (pp. 311--343).

Schoenfeld, A. H. (1992). Learning to Think Mathematically: Problem
Solving, Metacognition, and Sense-Making in Mathematics. In D. Grouws
(Ed.), \emph{Handbook for research on mathematics teaching and learning}
(pp. 334--370). New York: MacMillan.

Schwartz, D. L., Chase, C. C., \& Bransford, J. D. (2012). Resisting
Overzealous Transfer: Coordinating Previously Successful Routines With
Needs for New Learning. \emph{Educational Psychologist}, \emph{47}(3),
204--214. doi:10.1080/00461520.2012.696317

Seppälä, O., Malmi, L., \& Korhonen, A. (2006). Observations on student
misconceptions---A case study of the Build -- Heap Algorithm.
\emph{Computer Science Education}, \emph{16}(3), 241--255.
doi:10.1080/08993400600913523

Sherin, B. L. (2001). How students understand physics equations.
\emph{Cognition and Instruction}, \emph{19}(4), 479--541.

Smith, J. P., diSessa, A. A., \& Roschelle, J. (1993). Misconceptions
Reconceived: A Constructivist Analysis of Knowledge in Transition.
\emph{The Journal of the Learning Sciences}, \emph{3}(2), 115--163.

Soloway, E. (1986). Learning to program = learning to construct
mechanisms and explanations. \emph{Communications of the ACM},
\emph{29}, 850--858. doi:10.1145/6592.6594

Soloway, E., Bonar, J., \& Ehrlich, K. (1983). Cognitive Strategies and
Looping Constructs: An Empirical Study. \emph{Commun. ACM},
\emph{26}(11), 853--860. doi:10.1145/182.358436

Soloway, E., \& Spohrer, J. C. (Eds.). (1989). \emph{Studying the Novice
Programmer}. Hillsdale, N.J: L. Erlbaum Associates.

Spacco, J., Hovemeyer, D., Pugh, W., Hollingsworth, J., Padua-Perez, N.,
\& Emad, F. (2006). Experiences with Marmoset: Designing and Using an
Advanced Submission and Testing System for Programming Courses. In
\emph{ITiCSE '06: Proceedings of the 11th annual conference on
Innovation and technology in computer science education}. ACM Press.

Spacco, J., Pugh, W., Ayewah, N., \& Hovemeyer, D. (2006). The Marmoset
project: an automated snapshot, submission, and testing system. In
\emph{Companion to the 21st ACM SIGPLAN symposium on Object-oriented
programming systems, languages, and applications} (pp. 669--670). New
York, NY, USA: ACM. doi:10.1145/1176617.1176665

Spacco, J., Strecker, J., Hovemeyer, D., \& Pugh, W. (2005). Software
Repository Mining with Marmoset: An Automated Programming Project
Snapshot and Testing System. In \emph{Proceedings of the Mining Software
Repositories Workshop (MSR 2005)}. St. Louis, Missouri, USA.

Spohrer, J. C., \& Soloway, E. (1986). Alternatives to construct-based
programming misconceptions. \emph{ACM SIGCHI Bulletin}, \emph{17},
183--191. doi:10.1145/22339.22369

Stevens, R., \& Hall, R. (1998). Disciplined perception: Learning to see
in technoscience. In \emph{Talking Mathematics in School: Studies of
Teaching and Learning} (pp. 107--150). Cambridge, U.K: Cambridge
University Press.

Stieff, M. (2007). Mental rotation and diagrammatic reasoning in
science. \emph{Learning and Instruction}, \emph{17}(2), 219--234.
doi:10.1016/j.learninstruc.2007.01.012

Tabanao, E. S., Rodrigo, M. M. T., \& Jadud, M. C. (2011). Predicting
at-risk novice Java programmers through the analysis of online
protocols. In \emph{Proceedings of the seventh international workshop on
Computing education research} (pp. 85--92). New York, NY, USA: ACM.
doi:10.1145/2016911.2016930

Tannen, D. (Ed.). (1993). \emph{Framing in Discourse}. New York: Oxford
University Press.

Trakhtenbrot, M. (2013). Students Misconceptions in Analysis of
Algorithmic and Computational Complexity of Problems. In
\emph{Proceedings of the 18th ACM Conference on Innovation and
Technology in Computer Science Education} (pp. 353--354). New York, NY,
USA: ACM. doi:10.1145/2462476.2465604

Traweek, S. (1988). \emph{Beamtimes and lifetimes: the world of high
energy physicists}. Cambridge, Mass: Harvard University Press.

Turkle, S., \& Papert, S. (1991). Epistemological Pluralism and the
Reevaluation of the Concrete. In I. Harel, S. Papert, \& Massachusetts
Institute of Technology (Eds.), \emph{Constructionism: research reports
and essays, 1985-1990}. Norwood, N.J: Ablex Pub. Corp.

Valentine, D. W. (2004). CS educational research: a meta-analysis of
SIGCSE technical symposium proceedings. \emph{SIGCSE Bull.},
\emph{36}(1), 255--259. doi:10.1145/1028174.971391

Van de Sande, C. C., \& Greeno, J. G. (2012). Achieving Alignment of
Perspectival Framings in Problem-Solving Discourse. \emph{Journal of the
Learning Sciences}, \emph{21}(1), 1--44.
doi:10.1080/10508406.2011.639000

VanLehn, K. (1990). \emph{Mind Bugs: The Origins of Procedural
Misconceptions}. Cambridge, Mass: MIT Press.

Wagner, J. F. (2006). Transfer in Pieces. \emph{Cognition \&
Instruction}, \emph{24}(1), 1--71. doi:10.1207/s1532690xci2401\_1

Wortham, S. (2006). \emph{Learning Identity: The Joint Emergence of
Social Identification and Academic Learning}. Cambridge: Cambridge
University Press.

Yackel, E., \& Cobb, P. (1996). Sociomathematical Norms, Argumentation,
and Autonomy in Mathematics. \emph{Journal for Research in Mathematics
Education}, \emph{27}(4), 458--477. doi:10.2307/749877

Yin, R. K. (2009). \emph{Case study research: design and methods} (4th
ed.). Los Angeles, Calif: Sage Publications.